### 2016-03-03 article.Rnw
first commit

\documentclass[article, shortnames]{jss}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{booktabs}

\author{Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And 
        Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
\title{\pkg{webchem}: An \proglang{R} Package to retrieve Chemical Information from the Web}

\Plainauthor{Eduard Sz\"ocs, Ralf B. Sch\"afer}
\Plaintitle{webchem: an R Package to retrieve Chemical Information from the Web}
\Shorttitle{\pkg{webchem}: Chemical Information from the Web}

\Abstract{
A lot of chemical information is freely available online, including chemical identifiers and experimental data as well as predicted properties.
However, these data are scattered over various data sources and not easily accessible to researchers.
Manual searching and downloading of such is time-consuming and error-prone.

We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query data from currently 11 web-sources. These cover a broad spectrum of information.
The data are automatically imported into an R object and can directly be used in subsequent analyses.

\pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web-sources.
It facilitates data-cleaning, identifying and reporting substances and consequently reduces the time researchers spend on chemical data compilation and cleaning.
}
\Keywords{ecotoxicology, chemistry, data-cleaning, web-scraping}
\Plainkeywords{ecotoxicology, chemistry, data-cleaning, web-scraping}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Eduard Sz\"ocs\\
  Institute for Environmental Sciences\\
  Universit\"at Koblenz-Landau\\
  Fortstraße 7\\
  76829 Landau, Germany\\
  E-mail: \email{szoecs@uni-landau.de}\\
  URL: \url{https://edild.github.io}
}

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

\begin{document}

<<global_options, include=FALSE>>=
library(knitr)
knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
knitr::render_sweave()
knitr::opts_chunk$set(prompt = TRUE,
                      comment = NA,
                      highlight = FALSE,
                      dev.args = list(pointsize = 11),
                      # fig.width = 4.9,
                      # fig.height = 3.675,
                      fig.align = 'center',
                      out.format = 'sweave',
                      tidy = FALSE,
                      fig.path = '')
# 
hook.c <- function(x, options) {
  paste0('\\begin{CodeChunk}\n', x, '\\end{CodeChunk}')
}
hook.i <- function(x, options) {
  paste0(c('\\begin{CodeInput}', knitr:::hilight_source(x, 'latex', options), '\\end{CodeInput}', ''), collapse = '\n')
}
hook.s <- function(x, options) {
  paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
}
knitr::knit_hooks$set(chunk = hook.c,
               source = hook.i,
               output = hook.s,
               message = hook.s,
               warning = hook.s,
               plot = knitr::hook_plot_tex)

options(prompt = "R> ", continue = "+  ", width = 80, useFancyQuotes = FALSE)
@

\section[Introduction]{Introduction}
Before each statistical analysis, data cleaning is often required to ensure a good data quality.
Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on multiple data sources that need to be integrated.

When dealing with chemical datasets (e.g. environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent queries and appending of compound-related physico-chemical or toxicological information.
Several web source are freely available chemical names or link them to unique codes (see also section \emph{Data sources} below).
However, manual searching for each compound, often through a graphical web interface, is tedious, error-prone and not reproducible \citep{Peng_2009}.

To simplify, robustify and automate this task, i.e. to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
\proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports fully reproducibility of each step \citep{Marwick_2016}.

We describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.

\section[Implementation and design details]{Implementation and design details}
The \pkg{webchem} package is written entirely in \proglang{R} and available under a MIT license.
The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://goo.gl/OKbZaF}).
\pkg{webchem} is part of the rOpenSci project (\url{https://ropensci.org/}), which aims at fully reproducible data analysis.

\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via Travis-CI and Appveyor; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.

\pkg{webchem} builds on top of the following \proglang{R} packages:
\pkg{RCurl} \citep{lang_rcurl:_2015} and \pkg{httr} \citep{wickham_httr} for data transfer,
\pkg{stringr} \cite{wickham_stringr:_2015} for string handling,
\pkg{xml2} \citep{wickham_xml2} and \pkg{rvest} \citep{wickham_rvest} for parsing HTML and XML,
\pkg{jsonlite} \citep{ooms_jsonlite_2014} for parsing JSON,
\pkg{rcdk} \citep{guha_rcdk} for parsing SMILES.
For parsing molfiles we use a lightweight implementation of \citep{Grabner_Varmuza_Dehmer_2012}.

Some data sources provide application programming interfaces (API).
Web APIs define  functions that allow accessing services and data via http and return data in a specific way.
\pkg{webchem} uses the API of a data source provider, where this is provided.
For sources where an API is lacking, data is directly searched and extracted from the web pages, analogous to manual interaction with a website.

Only few design decisions have been made:
Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider (see next section) as source to retrieve compound informations.
Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
The prefix \code{get} is used to denote functions that query identifiers, e.g. \code{get\_csid} to retrieve the identifier used in ChemSpider.

\pkg{webchem} is friendly to the resources of data providers. 
Between each request there is a time-out of 0.3 to 2 seconds depending on data-source. 
Therefore, processing of larger datasets can take some time, but still represents a major improvement compared to manual lookup.
We provide a link the \emph{Terms of Use} of data-providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
Moreover, all functions return an URL of the source, which can be used for \mbox{(micro-)attribution}.

\section[Data sources]{Data sources}
The backbone of \pkg{webchem} are data sources providing their data and functionality to the public.
Currently, data can be retrieved from 11 sources.
These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and legal information (Figure~\ref{fig:fig1}, the supplement includes a detailed overview of all sources):

\begin{description}
  \item[NIH Chemical Identifier Resolver (CIR) \citep{cir}]{A web service that converts from and to various chemical identifiers.}
  \item[Chemical Translation Service (CTS) \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}]{A web service that converts from and to various chemical identifiers.}
  \item[ETOX \citep{etox}]{Information System Ecotoxicology and Environmental Quality Targets by the German Federal Environmental Agency. Provides basic identifiers, synonyms, ecotoxicological data and quality targets for different countries.}
  \item[PAN Pesticide Database \citep{pan}]{Information on pesticides - provides basic identifiers, ecotoxicological data and chemical properties.}
  \item[SRC Physprop \citep{physprop}]{Contains physical properties for over 41,000 chemicals.
  Physical properties collected from a wide variety of sources including experimental and modelled values.}
  \item[PubChem \citep{Kim_2016}]{PubChem is a public repository for information on chemical substances, providing identifiers, properties and synonyms.
  We use an interface to the PUG-REST web-service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
  \item[Wikidata \citep{wiki}]{Wikipedia contains information entries for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently \pkg{webchem} can query only chemical identifiers.}
  \item[Compendium of Pesticide Common Names \citep{wood}]{The compendium provides information on pesticide common names, identifiers and classification.}
  \item[ChemID\emph{plus} \citep{Tomasulo_2002}]{is a large web-based database provided by the National Library of Medicine. It provides identifiers, synonyms, toxicological data and chemical properties.}
  \item[ChemSpider \citep{pence_chemspider:_2010}]{ChemSpider is a free chemical structure database providing access to over 40 million structures. It provides identifiers, properties and can also be used to convert identifiers.}
  \item[OPSIN  \citep{Lowe_Corbett_Murray-Rust_Glen_2011}]{The Open Parser for Systematic IUPAC nomenclature is a chemical name interpreter and provides InChI and SMILES identifiers.}
\end{description}

The data sources themselves are interrelated and information available redundant. However, many also give information uniquely available and readers are encouraged to explore the possibilities.

\begin{figure}[ht]
  \centering
  \includegraphics{fig1.pdf}
  \caption{Overview of current data sources. Input and output possibilities currently implemented in the package.}
  \label{fig:fig1}
\end{figure}

\section[Use Cases]{Use Cases}
\subsection[Install webchem]{Installation}
\pkg{webchem} can be easily installed and loaded from CRAN:

<<install, eval=FALSE>>=
install.packages("webchem")
library(webchem)
@

<<load, echo=FALSE, eval=TRUE>>=
library(webchem)
@

The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.35094}.

\subsection[Sample datasets]{Sample datasets}
To demonstrate the capabilities of  \pkg{webchem} we use two small publicly available real world datasets.
The datasets are only for demonstration purposes, have been slightly preprocessed (not shown) and are available through the package.

(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the year 2013 in the river Jagst, Germany.
The data is publicly available and can be retrieved from \citep{lubw_2016}.
It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.

<<loadjagst>>=
# load jagst dataset
data(jagst)
# print first 6 lines of data
head(jagst)
@

This dataset identifies substances only by substance names. Values below LOQ are indicated by a qualifier column.

(ii) \code{lc50}: This data consists of median acute lethal concentration for the water flea daphnia magna in 48 h tests ($LC_{50, D.magna, 48h}$) of 124 insecticides.
The data has been retrieved from the EPA ECOTOX database \citep{epa_2016}.

<<load_lc50>>=
data(lc50)
head(lc50)
@

This data set identifies the substances only by CAS numbers.

\subsection[Query Identifiers]{Query Identifiers}
The \code{jagst} data set covers 34 substances that are identified by (German) names.
Merging and linking these to other tables is hampered by differences and ambiguity in compound names.

One possibility to resolve this, is to use different chemical identifiers allowing easy identification.
There are several identifiers available, e.g.  registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}. 
In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to match with other data.

As were are dealing with German substance names we start to query ETOX for CAS registry numbers.
A common work flow when dealing with web resources is to 1) query a unique identifier of the source, 2) use this identifier to retrieve additional information and 3) extract the parts that are needed from the \proglang{R}-object \citep{Chamberlain_Szocs_2013}.

First we search for ETOX internal ID numbers:

<<get_etoxid, message=FALSE>>=
# unique substance names
subs <- unique(jagst$substance)
# search ETOX IDs
ids <- get_etoxid(subs, match = 'best')
head(ids)
@

Only three substances could not be found in ETOX. 
Here we specify that only the \emph{'best'} match (in terms of the Levenshtein distance between query and results) is returned. 
A manual check confirms appropriate matches. 
Other options include: \emph{'all'} - returns all matches; \emph{'first'} - returns only the first match (not necessarily the best match); \emph{'ask'} - this enters a interactive mode, where the user is asked for a choice if multiple matches are found and \emph{'na'} which returns \code{NA} in case of multiple matches.

We use this data to retrieve basic information on the substances

<<etox_basic, message=FALSE>>=
etox_data <- etox_basic(ids$etoxid)
@

The available output / information content can be very voluminous.
Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
<<etox_cas, message=FALSE>>=
etox_cas <- cas(etox_data)
head(etox_cas)
@

There is divers data available and we cannot provide extractor function for each of those.
Therefore, if users need to extract other data, they have to write simple extractor functions (see following examples).

In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
<<token, echo=FALSE>>=
token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
@

<<others, message = FALSE, warning=FALSE>>=
# query SMILES from PubChem
cids <- get_cid(etox_cas)
pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
pc_smiles <- smiles(pc_data)

# query InChiKey from ChemSpider
csids <- get_csid(etox_cas, token = token)
cs_data <- cs_compinfo(csids, token = token)
cs_inchikey <- inchikey(cs_data)
@

<<table>>=
# combine in one data.frame
res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
                  cid = pc_data$CID, inchikey = cs_inchikey, 
                  csid = cs_data$csid, 
                  stringsAsFactors = FALSE)
@

Note, that in order to use the ChemSpider functions a personal authentification key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
We end up with a compound table containing many different identifiers, allowing easy identification and merging with other datasets (Table \ref{tab:comptable}).

\begin{table}[ht]
\centering
<<print_table, echo=FALSE, results = 'asis'>>=
library("xtable")
tmp <- res[1:6, ]
tmp$smiles <- ifelse(is.na(tmp$smiles), NA, paste0(substr(tmp$smiles, 1, 7), '...'))
tmp$inchikey <- ifelse(is.na(tmp$inchikey), NA, paste0(substr(tmp$inchikey, 1, 7), '...'))
tmp[is.na(tmp)] <- '-'
colnames(tmp) <- c("Name", "CAS", "SMILES\\textsuperscript{a}","CID", "InChIKey\\textsuperscript{a}", "CSID")
print(xtable(tmp),
  floating = FALSE,
  hline.after = NULL,
  include.rownames = FALSE,
  add.to.row = list(pos = list(-1,0, 6),
                    command = c('\\toprule\n','\\midrule\n','\\bottomrule\n')),
  sanitize.text.function = function(x){x}
  )
@
\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. - = not found. \textsuperscript{a} Only the first 7 characters are shown.}
\label{tab:comptable}
\end{table}

\subsection[Toxicity of different pesticide groups]{Toxicity of different pesticide groups}
Another question we might ask is \emph{How does Toxicity vary between insecticide groups?}
Answering this question requires tedious lookup of insecticide group for each of the 124 CAS numbers in the lc50 dataset.
The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried with \pkg{webchem}.

<<aw_query, message = FALSE>>=
# query the compendium using CAS-numbers
aw_data <- aw_query(lc50$cas, type = 'cas')
@

To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply it to the retrieved data

<<subactivty, message = FALSE>>=
# shows internal structure of the data-object
# str(aw_data[[1]])
# extract chemical group
igroup <- sapply(aw_data, function(y) y$subactivity[1])
igroup[1:3]
@

Figure \ref{fig:fig2} displays the result after some more data-cleaning.
Overall, it took only 5 \proglang{R} statements to retrieve the needed data, clean and plot it.

\begin{figure}[ht]
<<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
# cleanup
lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
              ifelse(grepl('neonicotinoid', igroup), 'Neo-\nnicotinoids', 
                ifelse(grepl('pyrethroid', igroup), 'Pyrethroids', 
                  ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
                         'other'))))
lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", "Organo-\nphosphates", "Neo-\nnicotinoids", "other"))

# plot
require(ggplot2)
p <- ggplot(lc50, aes(x = type, y = value)) +
  geom_boxplot(fill = 'grey75') + 
  geom_jitter(width = 0.5) +
  scale_y_log10() +
  labs(y = expression(LC[50~'D.magna, 48h']), x = '') +
  theme_bw() +
  theme(text = element_text(size = 16))
p
@
\caption{Toxicity of different pesticide groups. LC50 values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
\label{fig:fig2}
\end{figure}

\subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
Some data sources contain also data chemical properties which can be queried.
Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database in order to build a simple QSAR to predict toxicity.

<<qsar, message=FALSE, cache=TRUE, warning=FALSE>>=
# query PHYSPROP DATABASE
pp_data <- pp_query(lc50$cas)
@

The database contains predicted and experimental values.
Extracting log P from the data object is slightly more complicated,  
because i) for some compounds no data could be found and ii) the data-object has a more complicated structure (a data frame within a list).

<<qsar2, warnings = FALSE>>=
# look  at internal structure
# str(pp_data[[1]])
lc50$logp <- sapply(pp_data, function(y){
  if (length(y) == 1 && is.na(y))
    return(NA)
  y$prop$value[y$prop$variable == 'Log P (octanol-water)']
})
@

We opted for this more complex approach, because the information available is very divers and we cannot provide for each a extractor function.
Moreover, it gives users full control how to organize their data. 
Nevertheless, we provide in the documentation of each function how to extract more complicated parts of the data.
The resulting data and model is displayed in Figure~\ref{fig:fig3}.

\begin{figure}[ht]
<<plot_qsar, echo=FALSE, out.width='0.7\\textwidth', fig.height=6, warning=FALSE>>=
# model
mod <- lm(log10(value) ~ logp, data = lc50)
coefs <- round(coef(mod), 2) # coeficients
rmse <- round(summary(mod)$sigma, 2) # RMSE

# check data types
logp_type <- sapply(pp_data, function(y){
  if (length(y) == 1 && is.na(y))
    return(NA)
  y$prop$type[y$prop$variable == 'Log P (octanol-water)']
})
tabs <- table(logp_type)
sna <- sum(is.na(logp_type))

p <- ggplot(lc50, aes(x = logp, y = value)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = 'lm') +
  labs(x = 'log P', y = expression(LC[50])) +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 2, 4, 6), labels = format(10^c(0, 2, 4, 6), scientific = FALSE))
p
@
\caption{Simple QSAR. predicting logLC50 of pesticides by log P. 
Log P values have been queried from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
Blue line indicates the regression model ($log LC_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]} logP$, RMSE = \Sexpr{rmse}).}
\label{fig:fig3}
\end{figure}

\subsection[Utility functions]{Utility functions}
\pkg{webchem} provides also basic functions to check identifiers which can be used for data quality assessment.
The functions either use simple formatting rules or web resources.

<<check>>=
# simple formatting check
is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5')
# or using ChemSpider
is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
      type = 'chemspider')
# formatting check
is.cas('64-17-6')
@

\section[Discussion]{Discussion}
\subsection[Related software]{Related software}
Within the \proglang{R} ecosystem , there are only few similar projects:
\pkg{rpubchem} \citep{rpubchem_2014} and \pkg{ChemmineR} \citep{chemminer_2008}  both provide a interface to PubChem which is more extensive then the current one in \pkg{webchem}.
\pkg{WikidataR} \citep{wikidatar_2016} provides a interface to wikidata which could be used to retrieve chemical data from Wikipedia.
However, it does not provide predefined methods for chemistry like \pkg{webchem}.
Within the \proglang{Python} ecosystem (a popular programming language) there are the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} available.
\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these provides a unique programmatic interface.
The Chemical Translation Service \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also one of the sources that can be queried,  allows batch conversion of chemical identifiers.
However, it does not provide access to other data (experimental, modeled or legal data).

\subsection[Open Science]{Open Science}
More and more scientific data is becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
To be usable for other researchers chemical compounds should be properly identified, not only by chemical names but also with accompanying identifiers like InChIKey, SMILES and authority-assigned identifiers.
\pkg{webchem} provides an easy way to create such meta tables like Table \ref{tab:comptable}.

\subsection[Further development]{Further development}
We have outlined only a few use cases that likely will be use full for many researchers.
Given the huge amount of publicly available information there are of course many other possibilities.
\pkg{webchem} is currently under active development and there are a lot of data sources on the issue list not implemented yet.
GitHub makes contributing easy and we strongly encourage contribution to the package.
Moreover, comments, feedback and feature requests are highly welcome.

\section[Conclusions]{Conclusions}
Researchers need to have easy access to the global knowledge.
\pkg{webchem} saves time used gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.

\section*{Acknowledgments}
We thank all resource maintainers for their work making their data open to the public.
We thank Johannes Ranke and Daniel Münch for their contributions to the webchem package, as well as all users who provided feedback and feature requests.

\bibliography{references.bib}

\end{document}

### 2016-03-03 article.tex
first commit

--- 2016-03-03 article.Rnw
@@
-\documentclass[article, shortnames]{jss}
+\documentclass[article, shortnames]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\usepackage{Sweave}
+
 \usepackage[utf8]{inputenc} %unicode support
@@
 %% need no \usepackage{Sweave.sty}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-
-<<global_options, include=FALSE>>=
-library(knitr)
-knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
-knitr::render_sweave()
-knitr::opts_chunk$set(prompt = TRUE,
-                      comment = NA,
-                      highlight = FALSE,
-                      dev.args = list(pointsize = 11),
-                      # fig.width = 4.9,
-                      # fig.height = 3.675,
-                      fig.align = 'center',
-                      out.format = 'sweave',
-                      tidy = FALSE,
-                      fig.path = '')
-# 
-hook.c <- function(x, options) {
-  paste0('\\begin{CodeChunk}\n', x, '\\end{CodeChunk}')
-}
-hook.i <- function(x, options) {
-  paste0(c('\\begin{CodeInput}', knitr:::hilight_source(x, 'latex', options), '\\end{CodeInput}', ''), collapse = '\n')
-}
-hook.s <- function(x, options) {
-  paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
-}
-knitr::knit_hooks$set(chunk = hook.c,
-               source = hook.i,
-               output = hook.s,
-               message = hook.s,
-               warning = hook.s,
-               plot = knitr::hook_plot_tex)
-
-options(prompt = "R> ", continue = "+  ", width = 80, useFancyQuotes = FALSE)
-@
 
@@
 
-<<install, eval=FALSE>>=
-install.packages("webchem")
-library(webchem)
-@
-
-<<load, echo=FALSE, eval=TRUE>>=
-library(webchem)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> install.packages("webchem")
+R> library(webchem)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-<<loadjagst>>=
-# load jagst dataset
-data(jagst)
-# print first 6 lines of data
-head(jagst)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # load jagst dataset
+R> data(jagst)
+R> # print first 6 lines of data
+R> head(jagst)
+\end{CodeInput}
+\begin{CodeOutput}
+        date          substance value qual
+1 2013-01-04 2,4-Dimethylphenol 0.006    <
+2 2013-01-29 2,4-Dimethylphenol 0.006    <
+3 2013-02-26 2,4-Dimethylphenol 0.006    <
+4 2013-03-26 2,4-Dimethylphenol 0.006    <
+5 2013-04-23 2,4-Dimethylphenol 0.006    <
+6 2013-05-22 2,4-Dimethylphenol 0.006    <
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<load_lc50>>=
-data(lc50)
-head(lc50)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> data(lc50)
+R> head(lc50)
+\end{CodeInput}
+\begin{CodeOutput}
+       cas        value
+4  50-29-3    12.415277
+12 52-68-6     1.282980
+15 55-38-9    12.168138
+18 56-23-5 35000.000000
+21 56-38-2     1.539119
+36 57-74-9    98.400000
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<get_etoxid, message=FALSE>>=
-# unique substance names
-subs <- unique(jagst$substance)
-# search ETOX IDs
-ids <- get_etoxid(subs, match = 'best')
-head(ids)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # unique substance names
+R> subs <- unique(jagst$substance)
+R> # search ETOX IDs
+R> ids <- get_etoxid(subs, match = 'best')
+R> head(ids)
+\end{CodeInput}
+\begin{CodeOutput}
+  etoxid                           match distance                  query
+1   8668     2,4-Dimethylphenol ( 8668 )        0     2,4-Dimethylphenol
+2   8494 4-Chlor-2-methylphenol ( 8494 )        0 4-Chlor-2-methylphenol
+3   <NA>                            <NA>     <NA>     4-para-nonylphenol
+4   8397                Atrazin ( 8397 )        0                Atrazin
+5   7240                 Benzol ( 7240 )        0                 Benzol
+6   7331        Desethylatrazin ( 7331 )        0        Desethylatrazin
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<etox_basic, message=FALSE>>=
-etox_data <- etox_basic(ids$etoxid)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> etox_data <- etox_basic(ids$etoxid)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
-<<etox_cas, message=FALSE>>=
-etox_cas <- cas(etox_data)
-head(etox_cas)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> etox_cas <- cas(etox_data)
+R> head(etox_cas)
+\end{CodeInput}
+\begin{CodeOutput}
+       8668        8494        <NA>        8397        7240        7331 
+ "105-67-9" "1570-64-5"          NA "1912-24-9"   "71-43-2" "6190-65-4" 
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
-<<token, echo=FALSE>>=
-token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
-@
 
-<<others, message = FALSE, warning=FALSE>>=
-# query SMILES from PubChem
-cids <- get_cid(etox_cas)
-pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
-pc_smiles <- smiles(pc_data)
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # query SMILES from PubChem
+R> cids <- get_cid(etox_cas)
+R> pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+R> pc_smiles <- smiles(pc_data)
+R> 
+R> # query InChiKey from ChemSpider
+R> csids <- get_csid(etox_cas, token = token)
+R> cs_data <- cs_compinfo(csids, token = token)
+R> cs_inchikey <- inchikey(cs_data)
+\end{CodeInput}
+\end{CodeChunk}
 
-# query InChiKey from ChemSpider
-csids <- get_csid(etox_cas, token = token)
-cs_data <- cs_compinfo(csids, token = token)
-cs_inchikey <- inchikey(cs_data)
-@
-
-<<table>>=
-# combine in one data.frame
-res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-                  cid = pc_data$CID, inchikey = cs_inchikey, 
-                  csid = cs_data$csid, 
-                  stringsAsFactors = FALSE)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # combine in one data.frame
+R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
++                    cid = pc_data$CID, inchikey = cs_inchikey, 
++                    csid = cs_data$csid, 
++                    stringsAsFactors = FALSE)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 \centering
-<<print_table, echo=FALSE, results = 'asis'>>=
-library("xtable")
-tmp <- res[1:6, ]
-tmp$smiles <- ifelse(is.na(tmp$smiles), NA, paste0(substr(tmp$smiles, 1, 7), '...'))
-tmp$inchikey <- ifelse(is.na(tmp$inchikey), NA, paste0(substr(tmp$inchikey, 1, 7), '...'))
-tmp[is.na(tmp)] <- '-'
-colnames(tmp) <- c("Name", "CAS", "SMILES\\textsuperscript{a}","CID", "InChIKey\\textsuperscript{a}", "CSID")
-print(xtable(tmp),
-  floating = FALSE,
-  hline.after = NULL,
-  include.rownames = FALSE,
-  add.to.row = list(pos = list(-1,0, 6),
-                    command = c('\\toprule\n','\\midrule\n','\\bottomrule\n')),
-  sanitize.text.function = function(x){x}
-  )
-@
+\begin{CodeChunk}
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Thu Mar  3 13:29:34 2016
+\begin{tabular}{llllll}
+  \toprule
+Name & CAS & SMILES\textsuperscript{a} & CID & InChIKey\textsuperscript{a} & CSID \\ 
+  \midrule
+2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
+  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
+  4-para-nonylphenol & - & - & - & - & - \\ 
+  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
+  Benzol & 71-43-2 & C1=CC=C... & 241 & UHOVQNZ... & 236 \\ 
+  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
+   \bottomrule
+\end{tabular}
+\end{CodeChunk}
 \caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. - = not found. \textsuperscript{a} Only the first 7 characters are shown.}
@@
 
-<<aw_query, message = FALSE>>=
-# query the compendium using CAS-numbers
-aw_data <- aw_query(lc50$cas, type = 'cas')
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # query the compendium using CAS-numbers
+R> aw_data <- aw_query(lc50$cas, type = 'cas')
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-<<subactivty, message = FALSE>>=
-# shows internal structure of the data-object
-# str(aw_data[[1]])
-# extract chemical group
-igroup <- sapply(aw_data, function(y) y$subactivity[1])
-igroup[1:3]
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # shows internal structure of the data-object
+R> # str(aw_data[[1]])
+R> # extract chemical group
+R> igroup <- sapply(aw_data, function(y) y$subactivity[1])
+R> igroup[1:3]
+\end{CodeInput}
+\begin{CodeOutput}
+                                  50-29-3 
+            "organochlorine insecticides" 
+                                  52-68-6 
+               "phosphonate insecticides" 
+                                  55-38-9 
+"phenyl organothiophosphate insecticides" 
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 \begin{figure}[ht]
-<<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
-# cleanup
-lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
-              ifelse(grepl('neonicotinoid', igroup), 'Neo-\nnicotinoids', 
-                ifelse(grepl('pyrethroid', igroup), 'Pyrethroids', 
-                  ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
-                         'other'))))
-lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", "Organo-\nphosphates", "Neo-\nnicotinoids", "other"))
+\begin{CodeChunk}
 
-# plot
-require(ggplot2)
-p <- ggplot(lc50, aes(x = type, y = value)) +
-  geom_boxplot(fill = 'grey75') + 
-  geom_jitter(width = 0.5) +
-  scale_y_log10() +
-  labs(y = expression(LC[50~'D.magna, 48h']), x = '') +
-  theme_bw() +
-  theme(text = element_text(size = 16))
-p
-@
+{\centering \includegraphics[width=0.7\textwidth]{plot_lc50-1}
+
+}
+
+\end{CodeChunk}
 \caption{Toxicity of different pesticide groups. LC50 values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
@@
 
-<<qsar, message=FALSE, cache=TRUE, warning=FALSE>>=
-# query PHYSPROP DATABASE
-pp_data <- pp_query(lc50$cas)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # query PHYSPROP DATABASE
+R> pp_data <- pp_query(lc50$cas)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-<<qsar2, warnings = FALSE>>=
-# look  at internal structure
-# str(pp_data[[1]])
-lc50$logp <- sapply(pp_data, function(y){
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$value[y$prop$variable == 'Log P (octanol-water)']
-})
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # look  at internal structure
+R> # str(pp_data[[1]])
+R> lc50$logp <- sapply(pp_data, function(y){
++    if (length(y) == 1 && is.na(y))
++      return(NA)
++    y$prop$value[y$prop$variable == 'Log P (octanol-water)']
++  })
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 \begin{figure}[ht]
-<<plot_qsar, echo=FALSE, out.width='0.7\\textwidth', fig.height=6, warning=FALSE>>=
-# model
-mod <- lm(log10(value) ~ logp, data = lc50)
-coefs <- round(coef(mod), 2) # coeficients
-rmse <- round(summary(mod)$sigma, 2) # RMSE
+\begin{CodeChunk}
 
-# check data types
-logp_type <- sapply(pp_data, function(y){
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$type[y$prop$variable == 'Log P (octanol-water)']
-})
-tabs <- table(logp_type)
-sna <- sum(is.na(logp_type))
+{\centering \includegraphics[width=0.7\textwidth]{plot_qsar-1}
 
-p <- ggplot(lc50, aes(x = logp, y = value)) +
-  geom_point() +
-  scale_y_log10() +
-  geom_smooth(method = 'lm') +
-  labs(x = 'log P', y = expression(LC[50])) +
-  theme_bw() +
-  scale_x_continuous(breaks = c(0, 2, 4, 6), labels = format(10^c(0, 2, 4, 6), scientific = FALSE))
-p
-@
+}
+
+\end{CodeChunk}
 \caption{Simple QSAR. predicting logLC50 of pesticides by log P. 
-Log P values have been queried from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
-Blue line indicates the regression model ($log LC_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]} logP$, RMSE = \Sexpr{rmse}).}
+Log P values have been queried from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
+Blue line indicates the regression model ($log LC_{50} = 2.88\ensuremath{-0.37} logP$, RMSE = 1.45).}
 \label{fig:fig3}
@@
 
-<<check>>=
-# simple formatting check
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5')
-# or using ChemSpider
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
-      type = 'chemspider')
-# formatting check
-is.cas('64-17-6')
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # simple formatting check
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5')
+\end{CodeInput}
+\begin{CodeOutput}
+strings contains numbers.
+\end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\begin{CodeInput}
+R> # or using ChemSpider
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
++        type = 'chemspider')
+\end{CodeInput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\begin{CodeInput}
+R> # formatting check
+R> is.cas('64-17-6')
+\end{CodeInput}
+\begin{CodeOutput}
+Checksum is not correct! 5 vs. 6
+\end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\end{CodeChunk}
 

### 2016-03-04 article.Rnw
update
R script, more polishing

@@
 \Abstract{
-A lot of chemical information is freely available online, including chemical identifiers and experimental data as well as predicted properties.
+A lot of chemical information is freely available online, including identifiers, experimental and predicted chemical properties.
 However, these data are scattered over various data sources and not easily accessible to researchers.
@@
 
-We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query data from currently 11 web-sources. These cover a broad spectrum of information.
-The data are automatically imported into an R object and can directly be used in subsequent analyses.
+We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query chemical data from currently 11 web-sources. 
+These cover a broad spectrum of information.
+The data are automatically imported into an \proglang{R} object and can directly be used in subsequent analyses.
 
 \pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web-sources.
-It facilitates data-cleaning, identifying and reporting substances and consequently reduces the time researchers spend on chemical data compilation and cleaning.
+It facilitates data cleaning, identification and reporting of substances and retrieval of chemical data.
+Consequently it reduces the time researchers spend on chemical data compilation.
 }
-\Keywords{ecotoxicology, chemistry, data-cleaning, web-scraping}
-\Plainkeywords{ecotoxicology, chemistry, data-cleaning, web-scraping}
+\Keywords{ecotoxicology, chemistry, data-cleaning, web-scraping, ropensci}
+\Plainkeywords{ecotoxicology, chemistry, data-cleaning, web-scraping, ropensci}
 
@@
 
-\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via Travis-CI and Appveyor; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
+\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via Travis-CI and AppVeyor; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
 
@@
 Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
-They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider (see next section) as source to retrieve compound informations.
+They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound informations.
 Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
@@
 Currently, data can be retrieved from 11 sources.
-These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and legal information (Figure~\ref{fig:fig1}, the supplement includes a detailed overview of all sources):
+These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and legal information (Figure~\ref{fig:fig1}, a detailed overview of all sources is included as supplement):
 
@@
 library(webchem)
+library(ggplot2)
+library(xtable)
 @
 
-The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.35094}.
+The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
 
@@
 
-The available output / information content can be very voluminous.
+\pkg{webchem} always returns a named list (one entry for each substance) and the available output / information content can be very voluminous.
 Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
@@
 
-Note, that in order to use the ChemSpider functions a personal authentification key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
-We end up with a compound table containing many different identifiers, allowing easy identification and merging with other datasets (Table \ref{tab:comptable}).
+Note, that in order to use the ChemSpider functions a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
+We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g. the lc50 dataset based on CAS.
 
@@
 <<print_table, echo=FALSE, results = 'asis'>>=
-library("xtable")
 tmp <- res[1:6, ]
@@
 
-Figure \ref{fig:fig2} displays the result after some more data-cleaning.
-Overall, it took only 5 \proglang{R} statements to retrieve the needed data, clean and plot it.
+Figure \ref{fig:fig2} displays the result after some more data-cleaning (see supplement for full code).
+Overall, it took only 5 \proglang{R} statements to retrieve the needed data, clean and plot it (using \pkg{ggplot2} \citep{ggplot2}).
 
@@
 # plot
-require(ggplot2)
 p <- ggplot(lc50, aes(x = type, y = value)) +
@@
 Some data sources contain also data chemical properties which can be queried.
-Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database in order to build a simple QSAR to predict toxicity.
+Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database in order to build a simple Quantitative structure–activity relationship (QSAR) to predict toxicity.
 
@@
   geom_smooth(method = 'lm') +
-  labs(x = 'log P', y = expression(LC[50])) +
+  labs(x = 'P', y = expression(LC[50])) +
   theme_bw() +
@@
 \caption{Simple QSAR. predicting logLC50 of pesticides by log P. 
-Log P values have been queried from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
+Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
 Blue line indicates the regression model ($log LC_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]} logP$, RMSE = \Sexpr{rmse}).}
@@
 # simple formatting check
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5')
+is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
 # or using ChemSpider
@@
 
+\pkg{webchem} facilitates chemical data availability to researchers.
+However, data quality is also crucial for data-analysis \citep{Stieger_2014}. 
+Ensuring good quality requires additional effort and methods to be developed.
+
 \subsection[Further development]{Further development}
@@
 Researchers need to have easy access to the global knowledge.
-\pkg{webchem} saves time used gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.
+\pkg{webchem} can save \emph{"hundreds of working hours"} gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 We thank all resource maintainers for their work making their data open to the public.

### 2016-03-04 article.tex
update
R script, more polishing

@@
 
-The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.35094}.
+The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
 
@@
 % latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Thu Mar  3 13:29:34 2016
+% Fri Mar  4 08:35:28 2016
 \begin{tabular}{llllll}
@@
 \caption{Simple QSAR. predicting logLC50 of pesticides by log P. 
-Log P values have been queried from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
+Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
 Blue line indicates the regression model ($log LC_{50} = 2.88\ensuremath{-0.37} logP$, RMSE = 1.45).}
@@
 R> # simple formatting check
-R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5')
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
 \end{CodeInput}
 \begin{CodeOutput}
-strings contains numbers.
+Hyphens not at position 15 and 26.
 \end{CodeOutput}

### 2016-03-04 article.R

@@ -1,0 +1,139 @@
+# -------------------------------------------------------------------------
+# Supplemental R Code to reproduce the resutls of
+# Szöcs, Schäfer (2016). webchem: an R Package to retrieve Chemical Information from the Web
+# Date: 04.03.2016
+# ------------------------------------------------------------------------
+
+
+# Install webchem from CRAN -----------------------------------------------
+install.packages("webchem")
+
+
+# load packages -----------------------------------------------------------
+library(webchem)
+library(ggplot2)
+library(xtable)
+
+
+# load datasets -----------------------------------------------------------
+# load jagst dataset
+data(jagst)
+# print first 6 lines of data
+head(jagst)
+
+# lc50 dataset
+data(lc50)
+head(lc50)
+
+
+# Use Case 1: Query identifiers -------------------------------------------
+# unique substance names
+subs <- unique(jagst$substance)
+# search ETOX IDs, keeping only the best match
+ids <- get_etoxid(subs, match = 'best')
+head(ids)
+# use this id to query information from ETOX
+etox_data <- etox_basic(ids$etoxid)
+# extract CAS numbers from retrieved data
+etox_cas <- cas(etox_data)
+head(etox_cas)
+
+
+# query other identifiers from other resources
+# query SMILES from PubChem
+cids <- get_cid(etox_cas)
+pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+pc_smiles <- smiles(pc_data)
+# ChemSpider needs a security token
+token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
+# query InChiKey from ChemSpider
+csids <- get_csid(etox_cas, token = token)
+cs_data <- cs_compinfo(csids, token = token)
+cs_inchikey <- inchikey(cs_data)
+
+
+# combine in one data.frame
+res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+                  cid = pc_data$CID, inchikey = cs_inchikey, 
+                  csid = cs_data$csid, 
+                  stringsAsFactors = FALSE)
+head(res)
+
+
+
+# Use Case II: Toxicity of different pesticide groups ---------------------
+# query the pesticide compendium using CAS-numbers
[…]

### 2016-03-04 article.R
need xtable

@@ -12,7 +12,6 @@ install.packages("webchem")
 # load packages -----------------------------------------------------------
 library(webchem)
 library(ggplot2)
-library(xtable)
 
 
 # load datasets -----------------------------------------------------------

### 2016-03-08 article.Rnw
edits based on comments of rbs

@@
 However, these data are scattered over various data sources and not easily accessible to researchers.
-Manual searching and downloading of such is time-consuming and error-prone.
+Manual searching and downloading of such data is time-consuming and error-prone.
 
@@
 \pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web-sources.
-It facilitates data cleaning, identification and reporting of substances and retrieval of chemical data.
-Consequently it reduces the time researchers spend on chemical data compilation.
+In addition, it facilitates data cleaning, identification and reporting of substances.
+Consequently, it reduces the time researchers need to spend on chemical data compilation.
 }
@@
 Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
-In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on multiple data sources that need to be integrated.
+In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on multiple data sources that need to be joined.
 
 When dealing with chemical datasets (e.g.
-environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent queries and appending of compound-related physico-chemical or toxicological information.
-Several web source are freely available chemical names or link them to unique codes (see also section \emph{Data sources} below).
+environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
+Several web sources provide chemical names or link them to unique codes (see also section \emph{Data sources} below).
 However, manual searching for each compound, often through a graphical web interface, is tedious, error-prone and not reproducible \citep{Peng_2009}.
@@
 to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
-\proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports fully reproducibility of each step \citep{Marwick_2016}.
+\proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.
 
@@
 Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
-They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound informations.
+They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
 Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
-The prefix \code{get} is used to denote functions that query identifiers, e.g. \code{get\_csid} to retrieve the identifier used in ChemSpider.
+The prefix \code{get} is used to denote these functions, e.g. \code{get\_csid} to retrieve the identifier used in ChemSpider.
 
@@
 Therefore, processing of larger datasets can take some time, but still represents a major improvement compared to manual lookup.
-We provide a link the \emph{Terms of Use} of data-providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
+We provide a link to the \emph{Terms of Use} of data-providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
 Moreover, all functions return an URL of the source, which can be used for \mbox{(micro-)attribution}.
@@
 
-The data sources themselves are interrelated and information available redundant. However, many also give information uniquely available and readers are encouraged to explore the possibilities.
+Though the data sources exhibit some overlap in the provided information, each has been selected because it provides unique information and we encourage the interested reader to consult the related source for details.
+However, we provide a brief overview in the Supporting Information.
 
@@
 The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
+This document has been created using \pkg{webchem} version 0.1.
 
@@
 
-(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the year 2013 in the river Jagst, Germany.
+(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst of the year 2013, Germany.
 The data is publicly available and can be retrieved from \citep{lubw_2016}.
@@
 registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.
-In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to match with other data.
+In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to facilitate subsequent matching with other data.
 
@@
 A manual check confirms appropriate matches. 
-Other options include: \emph{'all'} - returns all matches; \emph{'first'} - returns only the first match (not necessarily the best match); \emph{'ask'} - this enters a interactive mode, where the user is asked for a choice if multiple matches are found and \emph{'na'} which returns \code{NA} in case of multiple matches.
+Other options include: \emph{'all'} - returns all matches; \emph{'first'} - returns only the first match (not necessarily the best match); \emph{'ask'} - this enters an interactive mode, where the user is asked for a choice if multiple matches are found and \emph{'na'} which returns \code{NA} in case of multiple matches.
 
-We use this data to retrieve basic information on the substances
+We use these data to retrieve basic information on the substances
 
@@
 
-There is divers data available and we cannot provide extractor function for each of those.
+A variety of data are available and we cannot provide extractor functions for each of those.
 Therefore, if users need to extract other data, they have to write simple extractor functions (see following examples).
@@
 <<token, echo=FALSE>>=
-token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
+# this is a webchem specific token. 
+# Please use it only for reproduction / testing.
+token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
 @
@@
 # query InChiKey from ChemSpider
+# token <- '<yourtokenhere>'
 csids <- get_csid(etox_cas, token = token)
@@
 
-Note, that in order to use the ChemSpider functions a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
+Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
 We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g. the lc50 dataset based on CAS.
@@
 \subsection[Toxicity of different pesticide groups]{Toxicity of different pesticide groups}
-Another question we might ask is \emph{How does Toxicity vary between insecticide groups?}
-Answering this question requires tedious lookup of insecticide group for each of the 124 CAS numbers in the lc50 dataset.
+Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
+Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the lc50 dataset.
 The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried with \pkg{webchem}.
@@
 
-To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply it to the retrieved data
+To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply this to the retrieved data
 
@@
 Figure \ref{fig:fig2} displays the result after some more data-cleaning (see supplement for full code).
-Overall, it took only 5 \proglang{R} statements to retrieve the needed data, clean and plot it (using \pkg{ggplot2} \citep{ggplot2}).
+Overall, it took only 5 \proglang{R} statements to retrieve, clean and plot the data using \pkg{ggplot2} \citep{ggplot2}.
 
@@
 \subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
-Some data sources contain also data chemical properties which can be queried.
-Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database in order to build a simple Quantitative structure–activity relationship (QSAR) to predict toxicity.
+Some data sources contain also data on chemical properties that can be queried.
+Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database to build a simple Quantitative structure–activity relationship (QSAR) to predict toxicity.
 
@@
 The database contains predicted and experimental values.
-Extracting log P from the data object is slightly more complicated,  
-because i) for some compounds no data could be found and ii) the data-object has a more complicated structure (a data frame within a list).
+Extracting $log~P_{oct/wat}$ from the data object is slightly more complicated,  
+because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).
 
@@
 
-We opted for this more complex approach, because the information available is very divers and we cannot provide for each a extractor function.
-Moreover, it gives users full control how to organize their data. 
+We opted for this more complex approach, because the information available is very divers and we cannot provide an extractor function for each purpose.
+Moreover, it gives users full control on how to organize their data. 
 Nevertheless, we provide in the documentation of each function how to extract more complicated parts of the data.
@@
 
-p <- ggplot(lc50, aes(x = logp, y = value)) +
+p <- ggplot(lc50, aes(x = logp, y = log10(value))) +
   geom_point() +
-  scale_y_log10() +
   geom_smooth(method = 'lm') +
-  labs(x = 'P', y = expression(LC[50])) +
-  theme_bw() +
-  scale_x_continuous(breaks = c(0, 2, 4, 6), labels = format(10^c(0, 2, 4, 6), scientific = FALSE))
+  labs(x = expression(log~P[oct/wat]), y = expression(log~LC[50])) +
+  theme_bw()
 p
 @
-\caption{Simple QSAR. predicting logLC50 of pesticides by log P. 
+\caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
 Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
@@
 \subsection[Utility functions]{Utility functions}
-\pkg{webchem} provides also basic functions to check identifiers which can be used for data quality assessment.
+\pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
 The functions either use simple formatting rules or web resources.
@@
 Within the \proglang{R} ecosystem , there are only few similar projects:
-\pkg{rpubchem} \citep{rpubchem_2014} and \pkg{ChemmineR} \citep{chemminer_2008}  both provide a interface to PubChem which is more extensive then the current one in \pkg{webchem}.
-\pkg{WikidataR} \citep{wikidatar_2016} provides a interface to wikidata which could be used to retrieve chemical data from Wikipedia.
-However, it does not provide predefined methods for chemistry like \pkg{webchem}.
-Within the \proglang{Python} ecosystem (a popular programming language) there are the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} available.
+\pkg{rpubchem} \citep{rpubchem_2014} and \pkg{ChemmineR} \citep{chemminer_2008} both provide an interface to PubChem that is more extensive than the current one in \pkg{webchem}.
+\pkg{WikidataR} \citep{wikidatar_2016} provides an interface to wikidata that could be used to retrieve chemical data from Wikipedia.
+However, it does not provide predefined methods for chemical data like \pkg{webchem}.
+Within the \proglang{Python} ecosystem (a popular programming language) the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
 \pkg{webchem} is not specialized and tries to integrate many data sources and for some of these provides a unique programmatic interface.
@@
 \subsection[Open Science]{Open Science}
-More and more scientific data is becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
+More and more scientific data are becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
 To be usable for other researchers chemical compounds should be properly identified, not only by chemical names but also with accompanying identifiers like InChIKey, SMILES and authority-assigned identifiers.
@@
 \subsection[Further development]{Further development}
-We have outlined only a few use cases that likely will be use full for many researchers.
-Given the huge amount of publicly available information there are of course many other possibilities.
-\pkg{webchem} is currently under active development and there are a lot of data sources on the issue list not implemented yet.
+We have outlined only a few use cases that likely will be useful for many researchers.
+Given the huge amount of publicly available information, many other possibilities can be envisioned.
+\pkg{webchem} is currently under active development and several other data sources have not been implemented yet but may be in the future.
 GitHub makes contributing easy and we strongly encourage contribution to the package.
@@
 \section[Conclusions]{Conclusions}
-Researchers need to have easy access to the global knowledge.
+Researchers need to have easy access to global knowledge on chemicals.
 \pkg{webchem} can save \emph{"hundreds of working hours"} gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.

### 2016-03-08 article.tex
edits based on comments of rbs

@@
 R> # query InChiKey from ChemSpider
+R> # token <- '<yourtokenhere>'
 R> csids <- get_csid(etox_cas, token = token)
@@
 \end{CodeChunk}
-\caption{Simple QSAR. predicting logLC50 of pesticides by log P. 
+\caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
 Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 

### 2016-03-08 article.R

@@ -44,7 +44,9 @@ cids <- get_cid(etox_cas)
 pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
 pc_smiles <- smiles(pc_data)
 # ChemSpider needs a security token
-token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
+# this is a webchem specific token. 
+# Please use it only for reproduction / testing.
+token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
 # query InChiKey from ChemSpider
 csids <- get_csid(etox_cas, token = token)
 cs_data <- cs_compinfo(csids, token = token)

### 2016-04-03 article.tex

@@
 \Abstract{
-A lot of chemical information is freely available online, including identifiers, experimental and predicted chemical properties.
+A wide range of chemical information is freely available online, including identifiers, experimental and predicted chemical properties.
 However, these data are scattered over various data sources and not easily accessible to researchers.
@@
 
-We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query chemical data from currently 11 web-sources. 
+We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query chemical data from currently 11 web sources. 
 These cover a broad spectrum of information.
@@
 
-\pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web-sources.
+\pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web sources.
 In addition, it facilitates data cleaning, identification and reporting of substances.
@@
 }
-\Keywords{ecotoxicology, chemistry, data-cleaning, web-scraping, ropensci}
-\Plainkeywords{ecotoxicology, chemistry, data-cleaning, web-scraping, ropensci}
+\Keywords{ecotoxicology, chemistry, data cleaning, web scraping, ropensci}
+\Plainkeywords{ecotoxicology, chemistry, data cleaning, web scraping, ropensci}
 
@@
 %% need no \usepackage{Sweave.sty}
+
+% citation alias for LUBW institution
+\defcitealias{lubw_2016}{LUBW (2016)}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 \section[Introduction]{Introduction}
-Before each statistical analysis, data cleaning is often required to ensure a good data quality.
-Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
+Before each statistical analysis, data cleaning is often required to ensure good data quality.
+Data cleaning is the process of detecting errors and inconsistencies in datasets \citep{Chapman_2005}.
 In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on multiple data sources that need to be joined.
 
-When dealing with chemical datasets (e.g.
-environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
+When dealing with chemical datasets (e.g.\ environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
 Several web sources provide chemical names or link them to unique codes (see also section \emph{Data sources} below).
@@
 
-To simplify, robustify and automate this task, i.e.
-to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
+To simplify, robustify and automate this task, i.e.\ to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
 \proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.
 
-We describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.
+In the following, we describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.
 
@@
 
-\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via Travis-CI and AppVeyor; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
+\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, (iii) version control (via git), (iv) an elaborate test suite covering more than 90\% of the relevant lines of code (currently $\sim 1500$ lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (v) continuous integration (via Travis-CI and AppVeyor; testing on Linux and Windows with current and development \proglang{R} versions), (vi) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vii) compliance with a style guide \citep{wickham_advanced_2015}.
 
@@
 \pkg{RCurl} \citep{lang_rcurl:_2015} and \pkg{httr} \citep{wickham_httr} for data transfer,
-\pkg{stringr} \cite{wickham_stringr:_2015} for string handling,
+\pkg{stringr} \citep{wickham_stringr:_2015} for string handling,
 \pkg{xml2} \citep{wickham_xml2} and \pkg{rvest} \citep{wickham_rvest} for parsing HTML and XML,
@@
 \pkg{rcdk} \citep{guha_rcdk} for parsing SMILES.
-For parsing molfiles we use a lightweight implementation of \citep{Grabner_Varmuza_Dehmer_2012}.
+For parsing molfiles we use a lightweight implementation from \citet{Grabner_Varmuza_Dehmer_2012}.
 
@@
 Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
-They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
+They follow the format of \code{source\_function}, e.g.\ \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
 Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
-The prefix \code{get} is used to denote these functions, e.g. \code{get\_csid} to retrieve the identifier used in ChemSpider.
+The prefix \code{get} is used to denote these functions, e.g.\ \code{get\_csid} to retrieve the identifier used in ChemSpider.
 
 \pkg{webchem} is friendly to the resources of data providers. 
-Between each request there is a time-out of 0.3 to 2 seconds depending on data-source. 
-Therefore, processing of larger datasets can take some time, but still represents a major improvement compared to manual lookup.
-We provide a link to the \emph{Terms of Use} of data-providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
+Between each request there is a time-out of 0.3 to 2 seconds depending on the datasource. 
+Therefore, processing larger datasets can take some time, but still represents a major improvement compared to manual lookup.
+We provide a link to the \emph{Terms of Use} of data providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
 Moreover, all functions return an URL of the source, which can be used for \mbox{(micro-)attribution}.
@@
   \item[PubChem \citep{Kim_2016}]{PubChem is a public repository for information on chemical substances, providing identifiers, properties and synonyms.
-  We use an interface to the PUG-REST web-service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
+  We use an interface to the PUG-REST web service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
   \item[Wikidata \citep{wiki}]{Wikipedia contains information entries for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}.
-Currently \pkg{webchem} can query only chemical identifiers.}
+Currently \pkg{webchem} can query chemical identifiers only.}
   \item[Compendium of Pesticide Common Names \citep{wood}]{The compendium provides information on pesticide common names, identifiers and classification.}
@@
 It provides identifiers, synonyms, toxicological data and chemical properties.}
-  \item[ChemSpider \citep{pence_chemspider:_2010}]{ChemSpider is a free chemical structure database providing access to over 40 million structures.
+  \item[ChemSpider \citep{pence_chemspider:_2010}]{is a free chemical structure database providing access to over 40 million structures.
 It provides identifiers, properties and can also be used to convert identifiers.}
@@
 
-The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
+The package is under active development. The latest development version is available from GitHub and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
 This document has been created using \pkg{webchem} version 0.1.
@@
 
-(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst of the year 2013, Germany.
-The data is publicly available and can be retrieved from \citep{lubw_2016}.
-It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.
+(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst, Germany, collected in 2013.
+The data is publicly available and can be retrieved from \citetalias{lubw_2016}.
+It comprises concentrations  (in $\mathrm{\mu g/L}$) of  34 substances  on 13 sampling occasions.
 
@@
 
-This dataset identifies substances only by substance names. Values below LOQ are indicated by a qualifier column.
+This dataset identifies substances only by substance names. Values below the limit of quantification (LOQ) are indicated by a qualifier column.
 
-(ii) \code{lc50}: This data consists of median acute lethal concentration for the water flea daphnia magna in 48 h tests ($LC_{50, D.magna, 48h}$) of 124 insecticides.
+(ii) \code{lc50}: This data consists of median acute lethal concentration for the water flea \textit{Daphnia magna} in 48 h tests ($LC_{50, D.magna, 48h}$) of 124 insecticides.
 The data has been retrieved from the EPA ECOTOX database \citep{epa_2016}.
@@
 
-This data set identifies the substances only by CAS numbers.
+This dataset identifies the substances only by CAS numbers.
 
 \subsection[Query Identifiers]{Query Identifiers}
-The \code{jagst} data set covers 34 substances that are identified by (German) names.
+The \code{jagst} dataset covers 34 substances that are identified by (German) names.
 Merging and linking these to other tables is hampered by differences and ambiguity in compound names.
 
-One possibility to resolve this, is to use different chemical identifiers allowing easy identification.
-There are several identifiers available, e.g.
-registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.
+One possibility to resolve this is to use different chemical identifiers allowing easy identification.
+There are several identifiers available, e.g.\ registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.
 In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to facilitate subsequent matching with other data.
@@
 As were are dealing with German substance names we start to query ETOX for CAS registry numbers.
-A common work flow when dealing with web resources is to 1) query a unique identifier of the source, 2) use this identifier to retrieve additional information and 3) extract the parts that are needed from the \proglang{R}-object \citep{Chamberlain_Szocs_2013}.
+A common work flow when dealing with web resources is to 1) query a unique identifier of the source, 2) use this identifier to retrieve additional information and 3) extract the parts that are needed from the \proglang{R} object \citep{Chamberlain_Szocs_2013}.
 
@@
 
-We use these data to retrieve basic information on the substances
+We use these data to retrieve basic information on the substances.
 
@@
 
-\pkg{webchem} always returns a named list (one entry for each substance) and the available output / information content can be very voluminous.
+\pkg{webchem} always returns a named list (one entry for each substance) and the available output \slash\ information content can be very voluminous.
 Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
@@
 
-A variety of data are available and we cannot provide extractor functions for each of those.
+A variety of data is available and we cannot provide extractor functions for each of those.
 Therefore, if users need to extract other data, they have to write simple extractor functions (see following examples).
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
+In the same manner we can now query other identifiers from other source using CAS numbers (Figure~\ref{fig:fig1}):
 
@@
 Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
-We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g. the lc50 dataset based on CAS.
+We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g.\ the \code{lc50} dataset based on CAS.
 
@@
 Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
-Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the lc50 dataset.
+Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the \code{lc50} dataset.
 The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried with \pkg{webchem}.
@@
 \begin{CodeInput}
-R> # query the compendium using CAS-numbers
+R> # query the compendium using CAS numbers
 R> aw_data <- aw_query(lc50$cas, type = 'cas')
@@
 
-To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply this to the retrieved data
+To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply this to the retrieved data:
 
@@
 \begin{CodeInput}
-R> # shows internal structure of the data-object
+R> # shows internal structure of the data object
 R> # str(aw_data[[1]])
@@
 
-Figure \ref{fig:fig2} displays the result after some more data-cleaning (see supplement for full code).
+Figure \ref{fig:fig2} displays the result after some more data cleaning (see supplement for full code).
 Overall, it took only 5 \proglang{R} statements to retrieve, clean and plot the data using \pkg{ggplot2} \citep{ggplot2}.
@@
 \end{CodeChunk}
-\caption{Toxicity of different pesticide groups. LC50 values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
+\caption{Toxicity of different pesticide groups. LC\textsubscript{50} values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
 \label{fig:fig2}
@@
 Some data sources contain also data on chemical properties that can be queried.
-Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database to build a simple Quantitative structure–activity relationship (QSAR) to predict toxicity.
+Here we query for the \code{lc50} data the $\mathrm{log}~P_{oct/wat}$ from the SRCs PHYSPROP database to build a simple quantitative structure–activity relationship (QSAR) to predict toxicity.
 
@@
 The database contains predicted and experimental values.
-Extracting $log~P_{oct/wat}$ from the data object is slightly more complicated,  
-because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).
+Extracting $\mathrm{log}~P_{oct/wat}$ from the data object is slightly more complicated,  
+because i) for some compounds no data could be found and ii) the data object has a more complex structure (a data frame within a list).
 
@@
 
-We opted for this more complex approach, because the information available is very divers and we cannot provide an extractor function for each purpose.
-Moreover, it gives users full control on how to organize their data. 
-Nevertheless, we provide in the documentation of each function how to extract more complicated parts of the data.
+We opted for this more complex approach because the information available is very diverse and we cannot provide an extractor function for each purpose.
+Moreover, it gives users full control over how to organize their data. 
+Nevertheless, we provide in the documentation of each function examples on how to extract more complicated parts of the data.
 The resulting data and model is displayed in Figure~\ref{fig:fig3}.
@@
 \end{CodeChunk}
-\caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
+\caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
 Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
-Blue line indicates the regression model ($log LC_{50} = 2.88\ensuremath{-0.37} logP$, RMSE = 1.45).}
+Blue line indicates the regression model ($\mathrm{log LC}_{50} = 2.88\ensuremath{-0.37} \mathrm{log} P$, RMSE = 1.45).}
 \label{fig:fig3}
@@
 \subsection[Utility functions]{Utility functions}
-\pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
+Furthermore, \pkg{webchem} provides basic functions to check identifiers that can be used for data quality assessment.
 The functions either use simple formatting rules or web resources.
@@
 \subsection[Related software]{Related software}
-Within the \proglang{R} ecosystem , there are only few similar projects:
+Within the \proglang{R} ecosystem, there are only few similar projects:
 \pkg{rpubchem} \citep{rpubchem_2014} and \pkg{ChemmineR} \citep{chemminer_2008} both provide an interface to PubChem that is more extensive than the current one in \pkg{webchem}.
@@
 Within the \proglang{Python} ecosystem (a popular programming language) the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
-\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these provides a unique programmatic interface.
+\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these it provides a unique programmatic interface.
 The Chemical Translation Service \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also one of the sources that can be queried,  allows batch conversion of chemical identifiers.
@@
 \subsection[Open Science]{Open Science}
-More and more scientific data are becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
+An increasing number of scientific data is becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
 To be usable for other researchers chemical compounds should be properly identified, not only by chemical names but also with accompanying identifiers like InChIKey, SMILES and authority-assigned identifiers.
-\pkg{webchem} provides an easy way to create such meta tables like Table \ref{tab:comptable}.
+\pkg{webchem} provides an easy way to create such meta tables as shown in Table \ref{tab:comptable}.
 
 \pkg{webchem} facilitates chemical data availability to researchers.
-However, data quality is also crucial for data-analysis \citep{Stieger_2014}. 
+However, data quality is also crucial for data analysis \citep{Stieger_2014}. 
 Ensuring good quality requires additional effort and methods to be developed.
@@
 \subsection[Further development]{Further development}
-We have outlined only a few use cases that likely will be useful for many researchers.
+We have outlined only a few use cases that will likely be useful for many researchers.
 Given the huge amount of publicly available information, many other possibilities can be envisioned.
@@
 Researchers need to have easy access to global knowledge on chemicals.
-\pkg{webchem} can save \emph{"hundreds of working hours"} gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.
+\pkg{webchem} can save \emph{"hundreds of working hours"} gathering this knowledge \citep{Munch_Galizia_2016} so that researchers can focus on other tasks.
 

### 2016-04-23 revision1/article.Rnw
started revisions1

--- 2016-03-08 article.Rnw
@@
 %% need no \usepackage{Sweave.sty}
+
+% citation alias for LUBW institution
+\defcitealias{lubw_2016}{LUBW (2016)}
 
@@
 The \pkg{webchem} package is written entirely in \proglang{R} and available under a MIT license.
-The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://goo.gl/OKbZaF}).
+The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://CRAN.R-project.org/package=webchem}).
 \pkg{webchem} is part of the rOpenSci project (\url{https://ropensci.org/}), which aims at fully reproducible data analysis.
@@
   \centering
-  \includegraphics{fig1.pdf}
+  \includegraphics{fig1}
   \caption{Overview of current data sources. Input and output possibilities currently implemented in the package.}
@@
 (i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst of the year 2013, Germany.
-The data is publicly available and can be retrieved from \citep{lubw_2016}.
+The data is publicly available and can be retrieved from \citetalias{lubw_2016}.
 It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.
@@
 
-\bibliography{references.bib}
+\bibliography{references}
 

### 2016-04-23 revision1/article.tex
started revisions1

--- 2016-04-23 revision1/article.Rnw
@@
 \defcitealias{lubw_2016}{LUBW (2016)}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-
-<<global_options, include=FALSE>>=
-library(knitr)
-knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
-knitr::render_sweave()
-knitr::opts_chunk$set(prompt = TRUE,
-                      comment = NA,
-                      highlight = FALSE,
-                      dev.args = list(pointsize = 11),
-                      # fig.width = 4.9,
-                      # fig.height = 3.675,
-                      fig.align = 'center',
-                      out.format = 'sweave',
-                      tidy = FALSE,
-                      fig.path = '')
-# 
-hook.c <- function(x, options) {
-  paste0('\\begin{CodeChunk}\n', x, '\\end{CodeChunk}')
-}
-hook.i <- function(x, options) {
-  paste0(c('\\begin{CodeInput}', knitr:::hilight_source(x, 'latex', options), '\\end{CodeInput}', ''), collapse = '\n')
-}
-hook.s <- function(x, options) {
-  paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
-}
-knitr::knit_hooks$set(chunk = hook.c,
-               source = hook.i,
-               output = hook.s,
-               message = hook.s,
-               warning = hook.s,
-               plot = knitr::hook_plot_tex)
-
-options(prompt = "R> ", continue = "+  ", width = 80, useFancyQuotes = FALSE)
-@
 
@@
 
-<<install, eval=FALSE>>=
-install.packages("webchem")
-library(webchem)
-@
-
-<<load, echo=FALSE, eval=TRUE>>=
-library(webchem)
-library(ggplot2)
-library(xtable)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> install.packages("webchem")
+R> library(webchem)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
-<<token, echo=FALSE>>=
-# this is a webchem specific token. 
-# Please use it only for reproduction / testing.
-token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
-@
 
-<<others, message = FALSE, warning=FALSE>>=
-# query SMILES from PubChem
-cids <- get_cid(etox_cas)
-pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
-pc_smiles <- smiles(pc_data)
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # query SMILES from PubChem
+R> cids <- get_cid(etox_cas)
+R> pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+R> pc_smiles <- smiles(pc_data)
+R> 
+R> # query InChiKey from ChemSpider
+R> # token <- '<yourtokenhere>'
+R> csids <- get_csid(etox_cas, token = token)
+R> cs_data <- cs_compinfo(csids, token = token)
+R> cs_inchikey <- inchikey(cs_data)
+\end{CodeInput}
+\end{CodeChunk}
 
-# query InChiKey from ChemSpider
-# token <- '<yourtokenhere>'
-csids <- get_csid(etox_cas, token = token)
-cs_data <- cs_compinfo(csids, token = token)
-cs_inchikey <- inchikey(cs_data)
-@
-
-<<table>>=
-# combine in one data.frame
-res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-                  cid = pc_data$CID, inchikey = cs_inchikey, 
-                  csid = cs_data$csid, 
-                  stringsAsFactors = FALSE)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # combine in one data.frame
+R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
++                    cid = pc_data$CID, inchikey = cs_inchikey, 
++                    csid = cs_data$csid, 
++                    stringsAsFactors = FALSE)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 \centering
-<<print_table, echo=FALSE, results = 'asis'>>=
-tmp <- res[1:6, ]
-tmp$smiles <- ifelse(is.na(tmp$smiles), NA, paste0(substr(tmp$smiles, 1, 7), '...'))
-tmp$inchikey <- ifelse(is.na(tmp$inchikey), NA, paste0(substr(tmp$inchikey, 1, 7), '...'))
-tmp[is.na(tmp)] <- '-'
-colnames(tmp) <- c("Name", "CAS", "SMILES\\textsuperscript{a}","CID", "InChIKey\\textsuperscript{a}", "CSID")
-print(xtable(tmp),
-  floating = FALSE,
-  hline.after = NULL,
-  include.rownames = FALSE,
-  add.to.row = list(pos = list(-1,0, 6),
-                    command = c('\\toprule\n','\\midrule\n','\\bottomrule\n')),
-  sanitize.text.function = function(x){x}
-  )
-@
+\begin{CodeChunk}
+% latex table generated in R 3.2.3 by xtable 1.8-2 package
+% Fri Mar  4 08:35:28 2016
+\begin{tabular}{llllll}
+  \toprule
+Name & CAS & SMILES\textsuperscript{a} & CID & InChIKey\textsuperscript{a} & CSID \\ 
+  \midrule
+2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
+  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
+  4-para-nonylphenol & - & - & - & - & - \\ 
+  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
+  Benzol & 71-43-2 & C1=CC=C... & 241 & UHOVQNZ... & 236 \\ 
+  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
+   \bottomrule
+\end{tabular}
+\end{CodeChunk}
 \caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. - = not found. \textsuperscript{a} Only the first 7 characters are shown.}
@@
 \begin{figure}[ht]
-<<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
-# cleanup
-lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
-              ifelse(grepl('neonicotinoid', igroup), 'Neo-\nnicotinoids', 
-                ifelse(grepl('pyrethroid', igroup), 'Pyrethroids', 
-                  ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
-                         'other'))))
-lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", "Organo-\nphosphates", "Neo-\nnicotinoids", "other"))
+\begin{CodeChunk}
 
-# plot
-p <- ggplot(lc50, aes(x = type, y = value)) +
-  geom_boxplot(fill = 'grey75') + 
-  geom_jitter(width = 0.5) +
-  scale_y_log10() +
-  labs(y = expression(LC[50~'D.magna, 48h']), x = '') +
-  theme_bw() +
-  theme(text = element_text(size = 16))
-p
-@
+{\centering \includegraphics[width=0.7\textwidth]{plot_lc50-1}
+
+}
+
+\end{CodeChunk}
 \caption{Toxicity of different pesticide groups. LC50 values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
@@
 \begin{figure}[ht]
-<<plot_qsar, echo=FALSE, out.width='0.7\\textwidth', fig.height=6, warning=FALSE>>=
-# model
-mod <- lm(log10(value) ~ logp, data = lc50)
-coefs <- round(coef(mod), 2) # coeficients
-rmse <- round(summary(mod)$sigma, 2) # RMSE
+\begin{CodeChunk}
 
-# check data types
-logp_type <- sapply(pp_data, function(y){
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$type[y$prop$variable == 'Log P (octanol-water)']
-})
-tabs <- table(logp_type)
-sna <- sum(is.na(logp_type))
+{\centering \includegraphics[width=0.7\textwidth]{plot_qsar-1}
 
-p <- ggplot(lc50, aes(x = logp, y = log10(value))) +
-  geom_point() +
-  geom_smooth(method = 'lm') +
-  labs(x = expression(log~P[oct/wat]), y = expression(log~LC[50])) +
-  theme_bw()
-p
-@
+}
+
+\end{CodeChunk}
 \caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
-Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
-Blue line indicates the regression model ($log LC_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]} logP$, RMSE = \Sexpr{rmse}).}
+Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
+Blue line indicates the regression model ($log LC_{50} = 2.88\ensuremath{-0.37} logP$, RMSE = 1.45).}
 \label{fig:fig3}
@@
 
-<<check>>=
-# simple formatting check
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
-# or using ChemSpider
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
-      type = 'chemspider')
-# formatting check
-is.cas('64-17-6')
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # simple formatting check
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
+\end{CodeInput}
+\begin{CodeOutput}
+Hyphens not at position 15 and 26.
+\end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\begin{CodeInput}
+R> # or using ChemSpider
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
++        type = 'chemspider')
+\end{CodeInput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\begin{CodeInput}
+R> # formatting check
+R> is.cas('64-17-6')
+\end{CodeInput}
+\begin{CodeOutput}
+Checksum is not correct! 5 vs. 6
+\end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\end{CodeChunk}
 

### 2016-04-23 submission/article.tex
started revisions1

--- 2016-04-23 revision1/article.tex
@@
 The \pkg{webchem} package is written entirely in \proglang{R} and available under a MIT license.
-The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://CRAN.R-project.org/package=webchem}).
+The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://goo.gl/OKbZaF}).
 \pkg{webchem} is part of the rOpenSci project (\url{https://ropensci.org/}), which aims at fully reproducible data analysis.
@@
   \centering
-  \includegraphics{fig1}
+  \includegraphics{fig1.pdf}
   \caption{Overview of current data sources. Input and output possibilities currently implemented in the package.}
@@
 
-\bibliography{references}
+\bibliography{references.bib}
 

### 2016-04-23 submission/article.R

@@ -1,0 +1,140 @@
+# -------------------------------------------------------------------------
+# Supplemental R Code to reproduce the resutls of
+# Szöcs, Schäfer (2016). webchem: an R Package to retrieve Chemical Information from the Web
+# Date: 04.03.2016
+# ------------------------------------------------------------------------
+
+
+# Install webchem from CRAN -----------------------------------------------
+install.packages("webchem")
+
+
+# load packages -----------------------------------------------------------
+library(webchem)
+library(ggplot2)
+
+
+# load datasets -----------------------------------------------------------
+# load jagst dataset
+data(jagst)
+# print first 6 lines of data
+head(jagst)
+
+# lc50 dataset
+data(lc50)
+head(lc50)
+
+
+# Use Case 1: Query identifiers -------------------------------------------
+# unique substance names
+subs <- unique(jagst$substance)
+# search ETOX IDs, keeping only the best match
+ids <- get_etoxid(subs, match = 'best')
+head(ids)
+# use this id to query information from ETOX
+etox_data <- etox_basic(ids$etoxid)
+# extract CAS numbers from retrieved data
+etox_cas <- cas(etox_data)
+head(etox_cas)
+
+
+# query other identifiers from other resources
+# query SMILES from PubChem
+cids <- get_cid(etox_cas)
+pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+pc_smiles <- smiles(pc_data)
+# ChemSpider needs a security token
+# this is a webchem specific token. 
+# Please use it only for reproduction / testing.
+token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
+# query InChiKey from ChemSpider
+csids <- get_csid(etox_cas, token = token)
+cs_data <- cs_compinfo(csids, token = token)
+cs_inchikey <- inchikey(cs_data)
+
+
+# combine in one data.frame
+res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+                  cid = pc_data$CID, inchikey = cs_inchikey, 
+                  csid = cs_data$csid, 
+                  stringsAsFactors = FALSE)
+head(res)
+
+
+
[…]

### 2016-04-23 revision1/article.Rnw
revised ms

@@
 \Plainauthor{Eduard Sz\"ocs, Ralf B. Sch\"afer}
-\Plaintitle{webchem: an R Package to retrieve Chemical Information from the Web}
+\Plaintitle{webchem: an R Package to Retrieve Chemical Information from the Web}
 \Shorttitle{\pkg{webchem}: Chemical Information from the Web}
@@
 
-options(prompt = "R> ", continue = "+  ", width = 80, useFancyQuotes = FALSE)
+options(prompt = "R> ", continue = "+  ", width = 76, useFancyQuotes = FALSE)
 @
@@
 
-When dealing with chemical datasets (e.g.
-environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
+When dealing with chemical datasets (e.g.\ environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
 Several web sources provide chemical names or link them to unique codes (see also section \emph{Data sources} below).
@@
 
-To simplify, robustify and automate this task, i.e.
-to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
+To simplify, robustify and automate this task, i.e.\ to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
 \proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.
@@
 Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
-They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
+They follow the format of \code{source\_function}, e.g.\ \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
 Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
-The prefix \code{get} is used to denote these functions, e.g. \code{get\_csid} to retrieve the identifier used in ChemSpider.
+The prefix \code{get} is used to denote these functions, e.g.\ \code{get\_csid} to retrieve the identifier used in ChemSpider.
 
@@
 
-\section[Use Cases]{Use Cases}
+\section[Use cases]{Use cases}
 \subsection[Install webchem]{Installation}
@@
 install.packages("webchem")
-library(webchem)
+library("webchem")
 @
@@
 It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.
+First we load the dataset and take a look at the first six rows:
 
 <<loadjagst>>=
-# load jagst dataset
-data(jagst)
-# print first 6 lines of data
+data("jagst")
 head(jagst)
@@
 <<load_lc50>>=
-data(lc50)
+data("lc50")
 head(lc50)
@@
 
-\subsection[Query Identifiers]{Query Identifiers}
+\subsection[Query identifiers]{Query identifiers}
 The \code{jagst} data set covers 34 substances that are identified by (German) names.
@@
 One possibility to resolve this, is to use different chemical identifiers allowing easy identification.
-There are several identifiers available, e.g.
-registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.
+There are several identifiers available, e.g.\  registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.
 In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to facilitate subsequent matching with other data.
@@
 
-First we search for ETOX internal ID numbers:
+First we search for ETOX internal ID numbers using the substance names:
 
-<<get_etoxid, message=FALSE>>=
-# unique substance names
+<<get_etoxid, message=FALSE, cache=TRUE>>=
 subs <- unique(jagst$substance)
-# search ETOX IDs
 ids <- get_etoxid(subs, match = 'best')
@@
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
+In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}), like PubChem
 <<token, echo=FALSE>>=
@@
 
-<<others, message = FALSE, warning=FALSE>>=
-# query SMILES from PubChem
+<<pubchem, message = FALSE, warning=FALSE>>=
 cids <- get_cid(etox_cas)
@@
 pc_smiles <- smiles(pc_data)
+@
 
-# query InChiKey from ChemSpider
-# token <- '<yourtokenhere>'
+or ChemSpider
+<<chemspider, message = FALSE, warning=FALSE>>=
 csids <- get_csid(etox_cas, token = token)
@@
 
+Finally, we combine the queried data into one data.frame
 <<table>>=
-# combine in one data.frame
 res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-                  cid = pc_data$CID, inchikey = cs_inchikey, 
-                  csid = cs_data$csid, 
+  cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
   stringsAsFactors = FALSE)
@@
 Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
-We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g. the lc50 dataset based on CAS.
+We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g.\ the lc50 dataset based on CAS.
 
@@
 tmp[is.na(tmp)] <- '-'
-colnames(tmp) <- c("Name", "CAS", "SMILES\\textsuperscript{a}","CID", "InChIKey\\textsuperscript{a}", "CSID")
+colnames(tmp) <- c("Name", "CAS", "SMILES","CID", "InChIKey", "CSID")
 print(xtable(tmp),
@@
 @
-\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. - = not found. \textsuperscript{a} Only the first 7 characters are shown.}
+\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
 \label{tab:comptable}
@@
 Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the lc50 dataset.
-The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried with \pkg{webchem}.
+The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried using CAS numbers with \pkg{webchem}:
 
 <<aw_query, message = FALSE>>=
-# query the compendium using CAS-numbers
 aw_data <- aw_query(lc50$cas, type = 'cas')
@@
 <<subactivty, message = FALSE>>=
-# shows internal structure of the data-object
-# str(aw_data[[1]])
-# extract chemical group
 igroup <- sapply(aw_data, function(y) y$subactivity[1])
@@
 <<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
-# cleanup
 lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
@@
 
-# plot
 p <- ggplot(lc50, aes(x = type, y = value)) +
@@
 <<qsar, message=FALSE, cache=TRUE, warning=FALSE>>=
-# query PHYSPROP DATABASE
 pp_data <- pp_query(lc50$cas)
@@
 <<qsar2, warnings = FALSE>>=
-# look  at internal structure
-# str(pp_data[[1]])
-lc50$logp <- sapply(pp_data, function(y){
+lc50$logp <- sapply(pp_data, function(y) {
   if (length(y) == 1 && is.na(y))
@@
 \pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
-The functions either use simple formatting rules or web resources.
+The functions either use simple formatting rules,
 
-<<check>>=
-# simple formatting check
+<<check_format>>=
 is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
-# or using ChemSpider
+is.cas('64-17-6')
+@
+
+or web resources like ChemSpider
+<<check_web>>=
 is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
   type = 'chemspider')
-# formatting check
-is.cas('64-17-6')
 @
@@
 However, it does not provide predefined methods for chemical data like \pkg{webchem}.
-Within the \proglang{Python} ecosystem (a popular programming language) the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
+Within the \proglang{Python} ecosystem the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
 \pkg{webchem} is not specialized and tries to integrate many data sources and for some of these provides a unique programmatic interface.

### 2016-04-23 revision1/article.tex
revised ms

@@
 R> install.packages("webchem")
-R> library(webchem)
+R> library("webchem")
 \end{CodeInput}
@@
 It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.
+First we load the dataset and take a look at the first six rows:
 
@@
 \begin{CodeInput}
-R> # load jagst dataset
-R> data(jagst)
-R> # print first 6 lines of data
+R> data("jagst")
 R> head(jagst)
@@
 \begin{CodeInput}
-R> data(lc50)
+R> data("lc50")
 R> head(lc50)
@@
 
-First we search for ETOX internal ID numbers:
+First we search for ETOX internal ID numbers using the substance names:
 
@@
 \begin{CodeInput}
-R> # unique substance names
 R> subs <- unique(jagst$substance)
-R> # search ETOX IDs
 R> ids <- get_etoxid(subs, match = 'best')
@@
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
+In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}), like PubChem
 
@@
 \begin{CodeInput}
-R> # query SMILES from PubChem
 R> cids <- get_cid(etox_cas)
@@
 R> pc_smiles <- smiles(pc_data)
-R> 
-R> # query InChiKey from ChemSpider
-R> # token <- '<yourtokenhere>'
+\end{CodeInput}
+\end{CodeChunk}
+
+or ChemSpider
+\begin{CodeChunk}
+\begin{CodeInput}
 R> csids <- get_csid(etox_cas, token = token)
@@
 
+Finally, we combine the queried data into one data.frame
 \begin{CodeChunk}
 \begin{CodeInput}
-R> # combine in one data.frame
 R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-+                    cid = pc_data$CID, inchikey = cs_inchikey, 
-+                    csid = cs_data$csid, 
++    cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
 +    stringsAsFactors = FALSE)
@@
 \begin{CodeChunk}
-% latex table generated in R 3.2.3 by xtable 1.8-2 package
-% Fri Mar  4 08:35:28 2016
+% latex table generated in R 3.2.5 by xtable 1.8-2 package
+% Sat Apr 23 22:38:04 2016
 \begin{tabular}{llllll}
   \toprule
-Name & CAS & SMILES\textsuperscript{a} & CID & InChIKey\textsuperscript{a} & CSID \\ 
+Name & CAS & SMILES & CID & InChIKey & CSID \\ 
   \midrule
-2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
-  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
+2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & - & - \\ 
+  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & - & - \\ 
   4-para-nonylphenol & - & - & - & - & - \\ 
-  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
-  Benzol & 71-43-2 & C1=CC=C... & 241 & UHOVQNZ... & 236 \\ 
-  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
+  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & - & - \\ 
+  Benzol & 71-43-2 & C1=CC=C... & 241 & - & - \\ 
+  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & - & - \\ 
    \bottomrule
@@
 \end{CodeChunk}
-\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. - = not found. \textsuperscript{a} Only the first 7 characters are shown.}
+\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
 \label{tab:comptable}
@@
 Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the lc50 dataset.
-The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried with \pkg{webchem}.
+The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried using CAS numbers with \pkg{webchem}:
 
@@
 \begin{CodeInput}
-R> # query the compendium using CAS-numbers
 R> aw_data <- aw_query(lc50$cas, type = 'cas')
@@
 \begin{CodeInput}
-R> # shows internal structure of the data-object
-R> # str(aw_data[[1]])
-R> # extract chemical group
 R> igroup <- sapply(aw_data, function(y) y$subactivity[1])
@@
 \begin{CodeInput}
-R> # query PHYSPROP DATABASE
 R> pp_data <- pp_query(lc50$cas)
@@
 \begin{CodeInput}
-R> # look  at internal structure
-R> # str(pp_data[[1]])
-R> lc50$logp <- sapply(pp_data, function(y){
+R> lc50$logp <- sapply(pp_data, function(y) {
 +    if (length(y) == 1 && is.na(y))
@@
 \pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
-The functions either use simple formatting rules or web resources.
+The functions either use simple formatting rules,
 
@@
 \begin{CodeInput}
-R> # simple formatting check
 R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
@@
 \begin{CodeInput}
-R> # or using ChemSpider
-R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
-+        type = 'chemspider')
-\end{CodeInput}
-\begin{CodeOutput}
-[1] FALSE
-\end{CodeOutput}
-\begin{CodeInput}
-R> # formatting check
 R> is.cas('64-17-6')
@@
 \end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\end{CodeChunk}
+
+or web resources like ChemSpider
+\begin{CodeChunk}
+\begin{CodeInput}
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
++    type = 'chemspider')
+\end{CodeInput}
 \begin{CodeOutput}

### 2016-04-23 revision1/response/reponse.tex
revised ms

% ======================================================= %
% Document: TEMPLATE FOR RESPONSES TO REVIEWERS
% Author: Andrea Ballatore
% Date: Jan 7, 2013
% Source: https://raw.githubusercontent.com/ucd-spatial/Datasets/master/tex_response_to_reviewers_template/responses_to_reviewers.tex
% Modified by Eduard Szöcs, 10.03.2015
% ======================================================= %
\documentclass[12pt]{article}

% packages
\usepackage{graphicx}
\usepackage{url}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{color}
\definecolor{mygray}{gray}{0.6}
\usepackage[utf8]{inputenc}
\usepackage[onehalfspacing]{setspace}
\usepackage[
	round,	%(defaultage in the main file and \input ) for round parentheses;
	colon,	% (default) to separate multiple citations with colons;
	authoryear,% (default) for author-year citations;
	sort,		% orders multiple citations into the sequence in which they 
]{natbib}					
\usepackage[%disable
	]{todonotes}

\usepackage{anysize}
\marginsize{2.5cm}{2.5cm}{1.5cm}{2.5cm}

% macros
% add a counter
\newcounter{cN}
\setcounter{cN}{0}

\newcommand{\comment}[1]{
	\vspace{2em} 
	\refstepcounter{cN} % incrment counter
	\noindent \hangindent=0em \textbf{\textcolor{Maroon}{\uline{Comment \thecN}:~}}\emph{"#1"}
	}

\newcommand{\response}[1]{
	\\[0.25em] 
	\hangindent=2.3em \textbf{\textcolor{NavyBlue}{\uline{Response}:~}}#1 
	}

\usepackage[normalem]{ulem}
\definecolor{darkred}{rgb}{1,.6,.6}
\DeclareRobustCommand\problemline{\bgroup\markoverwith{\textcolor{darkred}{\rule[-0.9ex]{4pt}{3pt}}}\ULon}
\DeclareRobustCommand{\problem}[1]{\problemline{#1}} % soul
\setcounter{secnumdepth}{-1}

\begin{document}
% ======================================================= %
\title{Responses to reviewers\\~\\Ms. No. ESPR-D-15-00741\\submitted to\\Environmental Science and Pollution Research}

\author{Eduard Szöcs and Ralf B. Schäfer}

\maketitle
% ======================================================= %
\noindent Dear editor Dr. Schulz  and reviewers,\\

We are thankful for reviewing our manuscript and the comments that helped to improve the paper. 
We revised the manuscript according to comments of the three reviewers and are re-submitting the manuscript for consideration for publication in Environmental Science and Pollution Research.

In the remainder of this document, we describe the changes that we have made to the paper for resubmission. 
To assist the assessment of our changes we have submitted two versions of the revised manuscript: one with highlighted changes and another without any highlighting. 
Note, that we did not highlight changes in citations and figures.

\vspace{2em}
\hfill Kind regards,

\hfill Eduard Szöcs and Ralf B. Schäfer
\newpage

% ======================================================= %
\section{Response to Reviewer 1}
\vspace{-2em}

%done
\comment{The purpose of the manuscript is highly commendable. Unfortunately, the
present manuscript has some major shortcomings: The simulation results,
which constitute the key contribution, are far from being as unequivocal as
the title may lead one to think. Also, it is disturbing to see a fairly technical
(and for ESPR non-standard) formulation of statistical models and yet the
authors seem to lack some understanding of the models in some cases.}
\response{We changed the subtitle (see also comment 2) and updated the notation of the models used (see also comments 6, 7, 8, 10, 25, and 43).
The simulation scenarios are tailored to experiments and data often encountered in ecotoxicology, which are characterised by small sample sizes and a factorial design (see also comment 13). 
Therefore, we submitted to ESPR, though, it might be interesting also to a greater audience of non-ecotoxicologists.}

%done
\comment{About the title: Please consider a more informative title such as ”Comparison of statistical approaches for analysis of non-normally distributed
response in ecotoxicology”. Also in view of the fact that the results are
not very clearly favouring the generalized linear models (GLMs).}
\response{We agree and changed the subtitle to \\
\emph{"A comparison of statistical approaches for analysis of count and proportion data in ecotoxicology"}. 
}

%done
\comment{p. 2, left column, lines 2–12: The issue of LOEC/NOEC versus regres-
sion is interesting but not really relevant to the aim of the manuscript.
Consider skipping this part. However, you’re right that for small sam-
ple sizes it is extremely important to choose the most efficient statistical
analysis and the better the model describes the data the more efficient.}
\response{We shortened this section. However, we kept a short introduction to LOEC/NOEC as inference on parameters is also part of our results.}

%done
\comment{p. 2, left column, lines 56–59: You aren’t mentioning the issue with
back-transformation So, in particular, $\beta x$ is not additive changes from
control on the original scale. To many ecotoxicologists such estimates
on a transformed scale aren’t directly interpretable as claimed by the
authors. A more detailed explanation or at least discussion on how
to back-transform is warranted (this is an issue regardless of whether
GLMs or models for transformed responses are used!).}
\response{We agree and include transformation bias in our discussion.}

%done
\comment{p. 2, right column, line 1: The real advantage of using a generalized
linear model for count data shows up in case of many small counts and
in the presence of ties. For large counts (e.g., corresponding to a large
$\mu_C$ ) there is practically no difference. However, this difference may not
show up in simulations when the sample size is kept between 3 and 9.}
\response{We agree that for large counts there is no practical difference between methods. However, our simulation design aimed to mimic data sets frequently encountered in ecotoxicology. See also response to comment 13.}

%done
\comment{p. 2, right column, lines 19–21: It is not the response variable that is
linked to the linear predictors, it is the mean of the response variable
(McCullagh \& Nelder, 1989). This is a basic fact of generalized linear
models as you also show in Eqn. (3). Please explain Eqn. (3) in more
detail (explain the parameters).}
\response{We agree and rephrased this complete section.}

%done
\comment{p. 2, right column, lines 25–27: There exists no quasi-Poisson distri-
bution!!! You use R terminology, but this is not in this case statistical
terminology. Please consult McCullagh \& Nelder (1989) to understand
that over-dispersion is dealt with by means of an ad hoc scaling of the
standard errors that takes place after having fitted the ordinary Poisson
GLM. So Eqn. (4) doesn’t make sense.}
\response{We agree and rephrased this complete section.}

%done
\comment{p. 2, right column, Eqn. (5): Have you explained the parameters in
the text?}
\response{We added the missing explanation for eqn. 5. Moreover, we unified the explanations of the models.}

%done
\comment{p. 3, left column, lines 26–42: What about over-dispersion for binomial
data? Shouldn’t that also be mentioned now that you consider models
for over-dispersion for count data.}
\response{We added that overdispersion might also occur in binomial data and pointed the reader to \citet{warton_arcsine_2011}.}

%done
\comment{p. 3, left column, line 33: Why not specify the mean next to the
variance in all equations? Or skip the variance where not needed (as
in this case).}
\response{We edited sections 2.1 and 2.2 and specified the mean next to the variance.}

%done
\comment{p. 3, right column, Fig. 1: Please be precise in the figure text: By using
the Poisson model the width of confidence intervals is underestimated
in the presence of overdispersion.}
\response{We agree and stated more precisely that the width of confidence intervals is underestimated.}

%done
\comment{p. 3, right column, line 28: How is multiple testing taken care of? This
may also affect the simulation results.}
\response{We adjusted for multiple testing using the method of \citet{holm_simple_1979}. We added this information to the manuscript. 
For the first submission we did not use Holm's method for the binomial data simulations. However, we unified this with the rerun of simulations (see comment 14).}

%done
\comment{p. 4, left column, line 2: The simulations will not be very informative
for such small sample sizes (3–9) as all methods will have low power.
Suggestion: Do also consider larger sample sizes: 12, 25, 50, 100. Not
completely unrealistic any more as experiments tend to get larger. And
then instead keep the $\mu_C$ small: 2, 4, 8, 16 to get more ties.}
\response{Our simulation design aimed to mimic data sets frequently encountered in ecotoxicology. 
Sample sizes between 3 and 9 are very common in regulatory ecotoxicology. 
For example, mesocosm experiments rarely exceed four replicates per treatment due to logistic constraints \citep{szocs_analysing_2015}. 
Similarly the \emph{Daphnia magna} standard test \citep{oecd_test_2004} is usually conducted with four to five replicates.
We believe that the current simulation design is relevant for ecotoxicologists and did not simulate larger sample sizes.
Moreover, we provide a fully reproducible source code at \url{https://github.com/EDiLD/usetheglm}, enabling other researchers to easily adapt our simulation design to specific needs.
}

%done
\comment{p. 4, left column, lines 10 and 28: In simulation studies it is quite
common to use 1000 simulated datasets for each scenario (it would
also reduce the sampling variability in the results).}
\response{We agree and rerun our simulations generating 1000 datasets for all scenarios. Because of the increased computational burden we run the computation in parallel on Amazon EC2 instance (with 15 GB RAM, 2.8 GHz and 8 cores) - taking more then 12 hours for the count data simulations.}

%done
\comment{p. 4, left column, line 36: Please do define the type I error properly.}
\response{We agree and rephrased to: \\ \emph{"[...] Type 1 error (detection of an effect when there is none) [...]"}}

%done
\comment{p. 4, right column, lines 26–29: The lack of convergence is not sur-
prising as the simulated datasets are simply too small to fit complex
models. Please re-consider the entire concept of the simulation (in
particular which scenarios to consider).}
\response{Our simulations covered scenarios often encountered by ecotoxicologists and therefore, it is like that practicing ecotoxicologists will also encounter such convergence problems. 
See also response to comment 13.}

%done
\comment{p. 4, right column, lines 31–33: One explanation for the reduced power
is certainly the multiple testing issue. Perhaps this point could be
addressed?}
\response{We added this point to the discussion.}

%done
\comment{p. 4, right column, lines 57–58: Did LM really maintain the nominal
significance level in all cases? This seems to be a subjective assessment
(as the level is above in some cases for N = 3).}
\response{We rerun our analyses using 1000 simulations (see also comment 14) and updated this section.}

%done
\comment{p. 6, right column, lines 43–46: Why didn’t you also use the transfor-
mation suggested by Brock et al. (2015)? Would be useful to see the
comparison.}
\response{As pointed out in the introduction, there is little information and advice which transformation to use. 
The $log(A~y + 1)$ is the most commonly used transformation in ecotoxicological mesocosm experiments.
We used only the $log(A~y + 1)$ transformation because our paper did not aim to compare different transformations and there are some inconsistencies in \citet{brock_minimum_2015}.
They state:  \\
\emph{"For the examples presented in this paper, we
followed Van den Brink et al. (2000) using the transformation
y(x)=ln(ax+1), where x is the measured abundance and the
factor ‘a’ is selected in such a way that the lowest non-zero
abundance of the data set is transformed to 1."} \\
Maybe this is just a typographical error, but in the publication of \citet{van_den_brink_impact_2000} it is stated: \\
\emph{"We decided that the
factor Ax in the ln(Ax +1) transformation should
make 2 by taking the lowest abundance value
higher than zero for x."} \\
Moreover, in the supporting material of \citet{brock_minimum_2015}, they used a $log(2~y + 1)$ transformation to which we refer to in our manuscript.
}

%done
\comment{p. 6, right column, line 49: Dunnett test is R terminology (yes, it
may be found in other recent publications, but it doesn’t make it more
correct). Please use a term like ”comparisons to the control”.}
\response{We agree and changed accordingly.}

%done
\comment{p. 7, right column, lines 7–10: A nice result that is not found in
many publication: non-parametric approaches lack power (due to less
assumptions being made). However, your results for the parametric
models are all quite similar and as it stands the paper does not present
a strong case for the use of GLMs; and a more balanced abstract would
be appropriate.}
\response{We agree and revised the abstract.}

%done
\comment{p. 7, Simulations: A suggestion: remove all diverging discussion of
LOEC/NOEC, GLMs for multivariate data, GLMMs, sample size cal-
culation and concentrate on discussing the simulation results. Or, alter-
natively, discuss in much more detail that GLMMs may actually offer a
difference approach for handling over-dispersion in binomial and count
data (instead of considering completely different designs).}
\response{We agree and pointed to GLMM as an alternative to deal with overdispersion in the methods section and discussion.}

% ======================================================= %	
\section{Response to Reviewer 2}
\vspace{-2em}

%done
\comment{One point to consider in a resubmission is that GLMs aren't for use for all non-normal data - some qualifications are needed of when to use GLMs.  Distributionally, GLMs replace the assumption of normality with the assumption that data come from the exponential family, which does not cover all situations.  More critically, GLMs replace the assumption of equal variance with an assumption that the mean-variance relationship follows a pre-specified pattern, and violation of this assumption has similar impacts to violations of the equal-variance assumption in a linear model would have.  Particular situations where GLMs are useful are binary data (known to be a quadratic mean-variance), count data (especially count data with lots of low counts, which typically cannot be transformed to normality effectively), and proportions constructed from counts.  These are the examples highlighted in the text, and are of clear relevance in ecotoxicology, but the argument does not extend much
more broadly than these examples - it is not a case of "your data aren't normal so you should use GLMs".  It is more accurate to say "if you have binary, counts or proportions from counts you should use GLMs".  I think this requires a modest adjustment to the phrasing of the abstract and introduction to better qualify the situations where GLMs are appropriate.  e.g. where O'Hara and Kotze and Warton \& Hui are mentioned, this should be qualified to refer to counts and proportions (of counts) respectively.  I  don't think the title needs changing, sufficient qualification elsewhere of the type of non-normality if interest should do the trick.}
\response{We agree and revised our manuscript accordingly. We changed the subtitle (see also comment 2) and the abstract (see also comment 21). The main text is now more specific to counts and proportions from counts.}

%done
\comment{In the first paragraph, "positive" is used where "non-negative" is needed (i.e. zero is a possible value), also change bonded to bounded}
\response{We fully agree and changed accordingly.}

%done
\comment{Page 2, second column - the parameterisation isn't equal to the Poisson model - the mean model is the same, I think that is what was meant.}
\response{We edited this section (see also comment 6) and corrected this.}

%done
\comment{Page 2, second column - also, strictly speaking, not all the methods mentioned here are GLMs.  Negative binomial regression ("NB2") is a GLM if the overdispersion parameter is fixed, otherwise it is a generalisation of GLM.  Similarly, quasi-Poisson isn't really a GLM, but a related method.}
\response{We agree and mention that quasi-Poisson is a related post hoc method. See also comment 7.}

%done
\comment{Section 2.2 - a description of what is meant by binomial data would be helpful at the start of this section.  Perhaps for completeness a description of what is meant by counts would help at the start of section 2.1}
\response{We added a brief description of a binomial random variable to the start of section 2.2.}

%done
\comment{Section 2.2.2 - it would be useful to also mention GLMMs with a random intercept term.  Often with this sort of data there are extra sources of variation from one sample to the next that are not explained by the binomial assumption, and this is the most natural way to account for it.}
\response{We added a reference to GLMM and \citep{warton_arcsine_2011} to this section.}

%done
\comment{Sections 2.1, 2.2: references to relevant articles or texts would be useful for GLMs, where the reader can find more details at a level appropriate for readership (e.g. a Zuur text?  Quinn \& Keough?).  Also point readers to supp 2 for code and a worked example.}
\response{We are thankful for pointing us to the missing reference to supplement 2. We added this to the methods section and pointed to introductory textbooks at the end of the manuscript. }

%done
\comment{page 3, line 58, the parametric bootstrap}
\response{We changed accordingly.}

%done
\comment{Page 3, second column, line 30: Wald tests should be used with caution, when a mean estimate is zero (or close to it) they can have strange behaviour (related to the issue of separable data - parameter estimates and ses diverging, test statistics going to zero).}
\response{We are thankful for this comment and added discussion of the Hauck-Donner effect.}

%done
\comment{Page 4, second column, lower means: it is noted that the means are lower when data are transformed.  This is known as transformation bias, and occurs because the transform linear model is actually estimating something different - it is no longer trying to estimate mean response.  It would be helpful to say this and maybe include a useful reference.  This also gets at the issue of interpretability - a disadvantage of the transform approach is a loss of interpretability, because we are no longer modelling mean response.}
\response{We added discussion of transformation bias, see also comment 4.}

% done
\comment{page 7 first column, line 7: bounds not bonds}
\response{Thanks for pointing to this typo.}

%done
\comment{Section 4.1 the case study discussion was a little confusing.  So is the point that you can get different results by linear models according to transformation, but that GLMs can resolve the problem of choice of transformation?  (by replacing it with a decision about mean-variance relationship?)  I think a stronger point is that the GLM is usually a better fit to count data, no transformation can make data Gaussian when it has lots of zeros and small counts, but GLM is designed for such data.  If diagnostic tools support this argument for your case study they could be included.}
\response{Yes, we find it much easier to choose appropriate distributions than from the virtually infinite possibilities of transformations. 
We agree, that GLMs give predictions that better fit to the data. 
But note, that \citet{warton_many_2005} found that "\emph{a Gaussian model based on transformed abundances fitted
data surprisingly well}" (see also comment 59). }

%done
\comment{page 7 column 1 line 27: it is not so much unreliable and biased - they are estimating a different parameter (in an unbiased fashion) - they estimate the mean of log(Ax+1) data, and a problem is that this has no natural interpretation in terms of the original data.  It is fair to make the point that it is biased as an estimate of the mean response, but it would be helpful to explain why and how the estimators have problematic interpretations.}
\response{We added discussion of transformation bias, see also comment 4.}

%done
\comment{page 7 column 1 line 50: a priori not a priory\\
page 7 column 2 line 37: data are (data being the plural of datum)\\
page 7 column 2 line 42: GLMs have (or GLM has)}
\response{We are thankful for pointing to these errors and changed accordingly.}

%done
\comment{page 7 column 2 line 44: the higher power in Fig 5 seemed to be through the GLMs being more conservative, rather than due to greater efficiency.  This issue could likely be fixed using a parametric bootstrap, as was done previously in figure 2.
A related point is that the linear model had the advantage that it seemed to be better at maintaining nominal significance levels at small sample sizes, which should be mentioned.  As the authors say however it did not have as good power (and also loses something in interpretability and biased estimation of means).  Hence while the GLM has a number of advantages, combining it with param bootstrap at small sample sizes might be worth considering to address the Type I error issue.}
\response{We agree and pointed to parametric bootstrap for situations of low sample sizes or estimates on the boundary of parameter space. See also comment 31}

\comment{Supplement 2 is a useful addition, and it might be worth giving it a little more emphasis in the text to ensure ecotoxicologists wanting to try out new analyses give it a look.  A couple of suggestions though:
- there is no parametric bootstrap code there at the moment.  This would be easy enough to do by using the mvabund package, calling manyglm (even on a univariate response) then anova(..., resamp="monte.carlo")  will do a parametric bootstrap.
- you can construct residual plots for GLMs on this package, using the plot function on a fitted manyglm object.}
\response{We are thankful for this comment. We refined Supplement 2 to incorporate all described models and methods and point the reader to it in the methods section.}

% ======================================================= %	
\section{Response to Reviewer 3}
\vspace{-2em}

%done
\comment{With its simulations in an ecotox. context, the paper is a fine contribution to make researchers aware of the possibilities and advantages of a generalized linear models (GLM). What is lacking is pointing to the disaster of using GLM in an improper way, notably without adjustment for overdispersion. The paper and abstract can still be made more useful by given not just the recommendation for GLM but also to give advice and warnings of proper usage. For example, (almost) never use loglinear/Poisson regression without adjustment for overdispersion, be aware that a nicer looking model like the negative binomial may give inflated type I error, a case that is overcome in this paper by bootstrapping. The same applies to binomial/logit regression unless the experiment really consist of independent observations for each 0/1 result (and is not just a count with a predetermined maximum). In this light it is strange that the logistic/binomial case is treated differently from
the loglinear/Poisson case! Should this not be repaired or at least given more warning. \\
(1) Also, if the authors stick to the binomial/logistic without overdispersion, the normal model could be made powerful in the simulations by treating the error variance as known, because it is a known constant for the arcsine transformation. \\
(2) Also, the simulation now use slight overdispersion whereas the example /typical data has large overdispersion. Add simulations with large overdispersion. The current ones can go to supplementary. \\
(3) Without the use of a GLM equivalent of the Williams test all the advantage of the use of GLM in terms of power are gone.  See the example. Discuss this ambiguity. You can perhaps use a bootstrap test based on (GLM?) monotonic regression or similar. I know some cues/leads in this direction.}
\response{We rerun our simulations and now include also the poisson GLM with overdispersed data. This clearly shows the effect of not accounting for overdispersion and is discussed now. \\
We stick to the binomial simulations without overdispersion, because we don't know if overdispersion is a common phenomena with ecotoxcicological data.
However, we point the reader to overdispersion and methods to account for it. \\
Our simulations mimicked the data of the case study and we clarified this (see also comment 48). \\
Williams type contrasts are also available in a GLM framework. The choice of which contrasts to use really depends on the assumptions one wants to make and which questions to answer. 
However, it does not affect our comparisons between methods.
We stress this in the revised MS.
}

%done
\comment{Page 1 Line 22 right column: a continuous proportion is fine for area cover and so on, but not for proportion of surviving animals where is just "k out of n" turn into a proportion. Discrete therefore. }
\response{Thank you for pointing to this error. Of course, we meant discrete proportions and changed accordingly.}

%done
\comment{Line 58 right add ref to (Warton 2005).}
\response{This reference fits very well and we added it.}

%done
\comment{Page 2 Line 40 : , for y>0 should for y>= 0 and  be on the next line without an indent after formula (if it is not new paragraph).}
\response{We clarified this equation and moved this part of eqn. 1 to the text: \\
\emph{"The factor $A$ was chosen in such way that $Ay$ equals 2 for the lowest non-zero abundance value (y)."}}

%done
\comment{Line 51 x\_i is undefined. The model appear to specify a simple linear regression or a control-one\_treatment model. Neither is used in the paper!!! Modify the notation therefore..}
\response{We clarified this whole section. Eqn. 2 is in fact a simple linear regression, but with dummy coded predictors / treatments.}

%done
\comment{Line 59 Mention backtransformation here, which works fine for the confidence intervals, and not the backtransformed is the median on the original scale. If you want to backtransform to the original mean, use exp(mean+ 0.5*sigma2). (Aitchison \& Brown 1969) page 8 (Jongman et al. 1995)page 19. }
\response{We agree and added a discussion of backtransformation bias. See also comment 4.}

%done
\comment{Page 3 line 17 n= 4.10 = 40 what does the mean here/ what is the purpose?}
\response{We are sorry for this error. It was a remnant of an earlier version of the manuscript - removed.}

%done
\comment{Line 41 : give more attention to the condition for the binomial: independent observations of each of the successes (k) out of the number of experiments (n).
And: thus give attention to overdispersion in this case as well, particularly as you found nasty things for the LR of  NB.}
\response{We now point the reader to overdispersed binomial data, though our simulation design does not cover this case.}

%done
\comment{Line 28 right. Ref for Dunnett contrasts.}
\response{We agree and added reference to the original description of Dunnett (1955).}

%done
\comment{Line 59 right. Only slight overdispersion, whereas the example data have large  overdispersion. Add large overdispersion.}
\response{We clarified this section. We now specify the overdisperion in terms of $\kappa$ in both, the case study and the simulation design. 
The simulated data mimicked the case study and such data is common for mesocosm studies in ecotoxicology.}

%done
\comment{Page 4  L 19 ( missing.}
\response{We fixed the wrong citation type.}

%done
\comment{Line 56 left to 2 right. Four p-values: what to believe? Your simulations (although with small overdispersion) show it. The LR of NB cannot be trusted. So the 0.016 is out. The remaining tests all show about the same p-value if you interpret statistics properly. Please note that the distinction between significant and non-signicant (here p = 0.061 and 0.042) in itself is statistically insignificant. (Gelman \& Stern 2006).}
\response{We fully agree with the reviewer and cautioned that p-values should not be overinterpreted by ecotoxicologists (citing Gelman \& Stern 2006). 
Parameter estimates (and their uncertainty) give much more information - on which we also emphasized in Figure 1.}

%done
\comment{Line 4 right : the backtransformed mean gives a median on the original scale, a value that is more interesting than the mean (generally for skew data),  and for skew right always smaller than the mean.}
\response{We agree and added a discussion of back-transformation bias. See also comments 4 and 44.}

%done
\comment{Section 3.2
Always start with the Type I error results because the power results are without meaning if the Type I error is inflated. This also applies the order of the figs: interchange the two rows of subfigs. }
\response{We agree and changed the order (first type I error, then power) in the text, as well as in the figures.}

%done
\comment{21 right: remove GLM\_nb in this sentence as GLM\_nb must be discarded due to inflated Type I error.}
\response{We agree and changed accordingly.}

%done
\comment{L32 right. ", but this." Although you should reorder the results, I mention that the "but this" should at least be "This could fortunately be …"[and we could not have known for sure (apart from the general warnings about LR) without the simulations] This inflated error for GLM\_nb must receive more attention. It is an interesting result of this study: you must to bootstrap when using NB.}
\response{We reordered and edited the respective results and discussion sections and emphasized this.}

%done
\comment{Page 5
Fig. 2 legend. Add that n =100 simulations. Add the interpretation that  GLM\_nb has inflated type I error, and that its line for power is just added for completeness and should not be interpreted as an estimate of the true power.}
\response{We unified the simulations and generated simulated 1000 data sets for each type of scenario (see comment 14). Moreover, we added \\
\emph{"Power levels for models with inflated type I error are shown for completeness."}\\
to figure captions 2 and 3.}

%done
\comment{Page 6
Fig.5 LM for 0.8 has a high Type I error. Does it deviate significantly from 0.05? [same question for the high Type I for GLM\_nb in fig 2]. Use confidence interval for binomial p = 0.05 n = 100.}
\response{We refined our simulations (see comment 14, )}

%done
\comment{Case study : 
You can check whether the difference between log(2y+1) and log(Ay+1) with A = 0.182 already causes the difference. (it probably does not). The important difference is the use of the Willems test! Without the use of a GLM equivalent of the Williams test all the advantage of the use of GLM in terms of power are gone! See the report of the Williams test below.}
\response{The reviewer is correct that the Williams test has higher power (if the assumptions are met). 
Williams type contrast are also available in a GLM framework and the choice of contrast to use really depends on the assumptions one wants to make and which questions to answer. 
This explains the difference between the results of Brock 2015 and us, but it does not affect our comparisons between methods.}

%done
\comment{Page 7 L35 Move ref to Warton to after data.}
\response{We changed accordingly.}

%done
\comment{L45 Add a reference to (ter Braak \& Šmilauer 2014) who advocate the use of the transformation approach in a multivariate (dimension reduction) context.}
\response{We agree and added reference to ter Braak \& Šmilauer 2014.}

%done
\comment{Line 1 right (like GLMs) -\textgreater (like GLMs and the Williams test)}
\response{We did not investigate the differences between multiple comparison procedures (MCP). 
We used Dunnett constrasts in all simulations, because only these allow individual comparisons between treatments and the control. 
For a comparison of MCP see \citet{jaki_statistical_2013}, who recommend a combination of both, Dunnett and Williams constrasts.
}

%done
\comment{Line 10-40 Again: discuss Type I error first. Then no mention of GLM\_nb in power.}
\response{We agree and changed accordingly. See also comment 52.}

%done
\comment{Line 18 right. Add after "non-normal data' the point the importance of proper accounting for under and overdispersion when using GLM. What is lacking is pointing to the disaster of using GLM in an improper way, notably without adjustment for overdispersion.}
\response{We rerun our simulations and also included the Poisson GLM. 
The results for Poisson GLM are now also displayed in Figures 2+3 and we discuss the consequences of failing to account for overdisperion.}

%done
\comment{Line 41 right size of 9 -\textgreater size of 6-9 (?)}
\response{We rerun our simulations and changed this to \\
\emph{for all simulated sample sizes}.}

%done
\comment{Line 44 right the transformation -\textgreater the LM}
\response{We changed accordingly.}

%% --------------------------------
\newpage
\bibliography{refs}
\bibliographystyle{spbasic}

% ======================================================= %
\end{document}
% ======================================================= %

### 2016-04-23 revision1/article.Rnw
revised ms

@@
 \usepackage{booktabs}
+\usepackage{fixltx2e}
 
@@
         Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
-\title{\pkg{webchem}: An \proglang{R} Package to retrieve Chemical Information from the Web}
+\title{\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web}
 
@@
 
-We describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.
+In the following, we describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.
 
@@
 \pkg{webchem} is friendly to the resources of data providers. 
-Between each request there is a time-out of 0.3 to 2 seconds depending on data-source. 
+Between each request there is a time-out of 0.3 to 2 seconds depending on the datasource. 
 Therefore, processing of larger datasets can take some time, but still represents a major improvement compared to manual lookup.
-We provide a link to the \emph{Terms of Use} of data-providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
+We provide a link to the \emph{Terms of Use} of data providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
 Moreover, all functions return an URL of the source, which can be used for \mbox{(micro-)attribution}.
@@
   \item[PubChem \citep{Kim_2016}]{PubChem is a public repository for information on chemical substances, providing identifiers, properties and synonyms.
-  We use an interface to the PUG-REST web-service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
+  We use an interface to the PUG-REST web service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
   \item[Wikidata \citep{wiki}]{Wikipedia contains information entries for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}.
-Currently \pkg{webchem} can query only chemical identifiers.}
+Currently \pkg{webchem} can query only chemical identifiers only.}
   \item[Compendium of Pesticide Common Names \citep{wood}]{The compendium provides information on pesticide common names, identifiers and classification.}
@@
 
-(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst of the year 2013, Germany.
+(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst, Germany, sampled in 2013.
 The data is publicly available and can be retrieved from \citetalias{lubw_2016}.
-It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.
+It comprises concentrations  (in $\mathrm{\mu g~/~L}$) of  34 substances  on 13 sampling occasions.
 First we load the dataset and take a look at the first six rows:
@@
 
-\pkg{webchem} always returns a named list (one entry for each substance) and the available output / information content can be very voluminous.
+\pkg{webchem} always returns a named list (one entry for each substance) and the available information content can be very voluminous.
 Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
@@
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}), like PubChem
+In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
 <<token, echo=FALSE>>=
@@
 Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
-We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g.\ the lc50 dataset based on CAS.
+We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g.\ the \code{lc50} dataset based on CAS.
 
@@
 Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
-Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the lc50 dataset.
+Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the \code{lc50} dataset.
 The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried using CAS numbers with \pkg{webchem}:
@@
 @
-\caption{Toxicity of different pesticide groups. LC50 values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
+\caption{Toxicity of different pesticide groups. LC\textsubscript{50} values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
 \label{fig:fig2}
@@
 Some data sources contain also data on chemical properties that can be queried.
-Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database to build a simple Quantitative structure–activity relationship (QSAR) to predict toxicity.
+Here we query for the \code{lc50} data the $\mathrm{log}~P_{oct/wat}$ from the SRC PHYSPROP database to build a simple quantitative structure–activity relationship (QSAR) to predict toxicity.
 
@@
 The database contains predicted and experimental values.
-Extracting $log~P_{oct/wat}$ from the data object is slightly more complicated,  
+Extracting $\mathrm{log}~P_{oct/wat}$ from the data object is slightly more complicated,  
 because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).
@@
 
-We opted for this more complex approach, because the information available is very divers and we cannot provide an extractor function for each purpose.
-Moreover, it gives users full control on how to organize their data. 
-Nevertheless, we provide in the documentation of each function how to extract more complicated parts of the data.
+We opted for this more complex approach, because the information available is very diverse and we cannot provide an extractor function for each purpose.
+Moreover, it gives users full control over how to organize their data. 
+Nevertheless, we provide in the documentation of each function examples on how to extract more complicated parts of the data.
 The resulting data and model is displayed in Figure~\ref{fig:fig3}.
@@
 @
-\caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
+\caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
 Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
-Blue line indicates the regression model ($log LC_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]} logP$, RMSE = \Sexpr{rmse}).}
+Blue line indicates the regression model ($\mathrm{log~LC}_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]}~\mathrm{log} P$, RMSE = \Sexpr{rmse}).}
 \label{fig:fig3}
@@
 \subsection[Utility functions]{Utility functions}
-\pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
+Furthermore, \pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
 The functions either use simple formatting rules,
@@
 Within the \proglang{Python} ecosystem the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
-\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these provides a unique programmatic interface.
+\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these it provides a unique programmatic interface.
 The Chemical Translation Service \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also one of the sources that can be queried, allows batch conversion of chemical identifiers.
@@
 We thank Johannes Ranke and Daniel Münch for their contributions to the webchem package, as well as all users who provided feedback and feature requests. 
+We are grateful to Zacharias Steinmetz for the valuable comments on the manuscript.
 

### 2016-04-23 revision1/article.tex
revised ms

@@
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}), like PubChem
+In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
 
@@
 \end{CodeChunk}
-\caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
+\caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
 Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
-Blue line indicates the regression model ($log LC_{50} = 2.88\ensuremath{-0.37} logP$, RMSE = 1.45).}
+Blue line indicates the regression model ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log} P$, RMSE = 1.45).}
 \label{fig:fig3}

### 2016-04-23 revision1/response/reponse.tex
revised ms

@@
 % ======================================================= %
-\title{Responses to reviewers\\~\\Ms. No. ESPR-D-15-00741\\submitted to\\Environmental Science and Pollution Research}
+\title{Response to editorial team\\~\\JSS 2581: Szocs, Schafer \\ webchem: An R Package to Retrieve Chemical Information from the Web}
 
@@
 % ======================================================= %
-\noindent Dear editor Dr. Schulz  and reviewers,\\
+\noindent Dear editorial team\\
 
-We are thankful for reviewing our manuscript and the comments that helped to improve the paper. 
-We revised the manuscript according to comments of the three reviewers and are re-submitting the manuscript for consideration for publication in Environmental Science and Pollution Research.
-
-In the remainder of this document, we describe the changes that we have made to the paper for resubmission. 
-To assist the assessment of our changes we have submitted two versions of the revised manuscript: one with highlighted changes and another without any highlighting. 
-Note, that we did not highlight changes in citations and figures.
+We are thankful for pre-screening our manuscript to meet the requirements of JSS.
+We made all required changes.  Moreover, we incorporated minor comments from a colleague.
+Please find below detailed description of the changes made.
 
@@
 % ======================================================= %
-\section{Response to Reviewer 1}
-\vspace{-2em}
 
 %done
-\comment{The purpose of the manuscript is highly commendable. Unfortunately, the
-present manuscript has some major shortcomings: The simulation results,
-which constitute the key contribution, are far from being as unequivocal as
-the title may lead one to think. Also, it is disturbing to see a fairly technical
-(and for ESPR non-standard) formulation of statistical models and yet the
-authors seem to lack some understanding of the models in some cases.}
-\response{We changed the subtitle (see also comment 2) and updated the notation of the models used (see also comments 6, 7, 8, 10, 25, and 43).
-The simulation scenarios are tailored to experiments and data often encountered in ecotoxicology, which are characterised by small sample sizes and a factorial design (see also comment 13). 
-Therefore, we submitted to ESPR, though, it might be interesting also to a greater audience of non-ecotoxicologists.}
-
-%done
-\comment{About the title: Please consider a more informative title such as ”Comparison of statistical approaches for analysis of non-normally distributed
-response in ecotoxicology”. Also in view of the fact that the results are
-not very clearly favouring the generalized linear models (GLMs).}
-\response{We agree and changed the subtitle to \\
-\emph{"A comparison of statistical approaches for analysis of count and proportion data in ecotoxicology"}. 
-}
-
-%done
-\comment{p. 2, left column, lines 2–12: The issue of LOEC/NOEC versus regres-
-sion is interesting but not really relevant to the aim of the manuscript.
-Consider skipping this part. However, you’re right that for small sam-
-ple sizes it is extremely important to choose the most efficient statistical
-analysis and the better the model describes the data the more efficient.}
-\response{We shortened this section. However, we kept a short introduction to LOEC/NOEC as inference on parameters is also part of our results.}
-
-%done
-\comment{p. 2, left column, lines 56–59: You aren’t mentioning the issue with
-back-transformation So, in particular, $\beta x$ is not additive changes from
-control on the original scale. To many ecotoxicologists such estimates
-on a transformed scale aren’t directly interpretable as claimed by the
-authors. A more detailed explanation or at least discussion on how
-to back-transform is warranted (this is an issue regardless of whether
-GLMs or models for transformed responses are used!).}
-\response{We agree and include transformation bias in our discussion.}
-
-%done
-\comment{p. 2, right column, line 1: The real advantage of using a generalized
-linear model for count data shows up in case of many small counts and
-in the presence of ties. For large counts (e.g., corresponding to a large
-$\mu_C$ ) there is practically no difference. However, this difference may not
-show up in simulations when the sample size is kept between 3 and 9.}
-\response{We agree that for large counts there is no practical difference between methods.
-However, our simulation design aimed to mimic data sets frequently encountered in ecotoxicology.
-See also response to comment 13.}
-
-%done
-\comment{p. 2, right column, lines 19–21: It is not the response variable that is
-linked to the linear predictors, it is the mean of the response variable
-(McCullagh \& Nelder, 1989). This is a basic fact of generalized linear
-models as you also show in Eqn. (3). Please explain Eqn. (3) in more
-detail (explain the parameters).}
-\response{We agree and rephrased this complete section.}
-
-%done
-\comment{p. 2, right column, lines 25–27: There exists no quasi-Poisson distri-
-bution!!! You use R terminology, but this is not in this case statistical
-terminology. Please consult McCullagh \& Nelder (1989) to understand
-that over-dispersion is dealt with by means of an ad hoc scaling of the
-standard errors that takes place after having fitted the ordinary Poisson
-GLM. So Eqn. (4) doesn’t make sense.}
-\response{We agree and rephrased this complete section.}
-
-%done
-\comment{p. 2, right column, Eqn. (5): Have you explained the parameters in
-the text?}
-\response{We added the missing explanation for eqn. 5. Moreover, we unified the explanations of the models.}
-
-%done
-\comment{p. 3, left column, lines 26–42: What about over-dispersion for binomial
-data? Shouldn’t that also be mentioned now that you consider models
-for over-dispersion for count data.}
-\response{We added that overdispersion might also occur in binomial data and pointed the reader to \citet{warton_arcsine_2011}.}
-
-%done
-\comment{p. 3, left column, line 33: Why not specify the mean next to the
-variance in all equations? Or skip the variance where not needed (as
-in this case).}
-\response{We edited sections 2.1 and 2.2 and specified the mean next to the variance.}
-
-%done
-\comment{p. 3, right column, Fig. 1: Please be precise in the figure text: By using
-the Poisson model the width of confidence intervals is underestimated
-in the presence of overdispersion.}
-\response{We agree and stated more precisely that the width of confidence intervals is underestimated.}
-
-%done
-\comment{p. 3, right column, line 28: How is multiple testing taken care of? This
-may also affect the simulation results.}
-\response{We adjusted for multiple testing using the method of \citet{holm_simple_1979}. We added this information to the manuscript. 
-For the first submission we did not use Holm's method for the binomial data simulations. However, we unified this with the rerun of simulations (see comment 14).}
-
-%done
-\comment{p. 4, left column, line 2: The simulations will not be very informative
-for such small sample sizes (3–9) as all methods will have low power.
-Suggestion: Do also consider larger sample sizes: 12, 25, 50, 100. Not
-completely unrealistic any more as experiments tend to get larger. And
-then instead keep the $\mu_C$ small: 2, 4, 8, 16 to get more ties.}
-\response{Our simulation design aimed to mimic data sets frequently encountered in ecotoxicology. 
-Sample sizes between 3 and 9 are very common in regulatory ecotoxicology. 
-For example, mesocosm experiments rarely exceed four replicates per treatment due to logistic constraints \citep{szocs_analysing_2015}. 
-Similarly the \emph{Daphnia magna} standard test \citep{oecd_test_2004} is usually conducted with four to five replicates.
-We believe that the current simulation design is relevant for ecotoxicologists and did not simulate larger sample sizes.
-Moreover, we provide a fully reproducible source code at \url{https://github.com/EDiLD/usetheglm}, enabling other researchers to easily adapt our simulation design to specific needs.
-}
-
-%done
-\comment{p. 4, left column, lines 10 and 28: In simulation studies it is quite
-common to use 1000 simulated datasets for each scenario (it would
-also reduce the sampling variability in the results).}
-\response{We agree and rerun our simulations generating 1000 datasets for all scenarios.
-Because of the increased computational burden we run the computation in parallel on Amazon EC2 instance (with 15 GB RAM, 2.8 GHz and 8 cores) - taking more then 12 hours for the count data simulations.}
-
-%done
-\comment{p. 4, left column, line 36: Please do define the type I error properly.}
-\response{We agree and rephrased to: \\ \emph{"[...] Type 1 error (detection of an effect when there is none) [...]"}}
-
-%done
-\comment{p. 4, right column, lines 26–29: The lack of convergence is not sur-
-prising as the simulated datasets are simply too small to fit complex
-models. Please re-consider the entire concept of the simulation (in
-particular which scenarios to consider).}
-\response{Our simulations covered scenarios often encountered by ecotoxicologists and therefore, it is like that practicing ecotoxicologists will also encounter such convergence problems. 
-See also response to comment 13.}
-
-%done
-\comment{p. 4, right column, lines 31–33: One explanation for the reduced power
-is certainly the multiple testing issue. Perhaps this point could be
-addressed?}
-\response{We added this point to the discussion.}
-
-%done
-\comment{p. 4, right column, lines 57–58: Did LM really maintain the nominal
-significance level in all cases? This seems to be a subjective assessment
-(as the level is above in some cases for N = 3).}
-\response{We rerun our analyses using 1000 simulations (see also comment 14) and updated this section.}
-
-%done
-\comment{p. 6, right column, lines 43–46: Why didn’t you also use the transfor-
-mation suggested by Brock et al. (2015)? Would be useful to see the
-comparison.}
-\response{As pointed out in the introduction, there is little information and advice which transformation to use. 
-The $log(A~y + 1)$ is the most commonly used transformation in ecotoxicological mesocosm experiments.
-We used only the $log(A~y + 1)$ transformation because our paper did not aim to compare different transformations and there are some inconsistencies in \citet{brock_minimum_2015}.
-They state:  \\
-\emph{"For the examples presented in this paper, we
-followed Van den Brink et al. (2000) using the transformation
-y(x)=ln(ax+1), where x is the measured abundance and the
-factor ‘a’ is selected in such a way that the lowest non-zero
-abundance of the data set is transformed to 1."} \\
-Maybe this is just a typographical error, but in the publication of \citet{van_den_brink_impact_2000} it is stated: \\
-\emph{"We decided that the
-factor Ax in the ln(Ax +1) transformation should
-make 2 by taking the lowest abundance value
-higher than zero for x."} \\
-Moreover, in the supporting material of \citet{brock_minimum_2015}, they used a $log(2~y + 1)$ transformation to which we refer to in our manuscript.
-}
-
-%done
-\comment{p. 6, right column, line 49: Dunnett test is R terminology (yes, it
-may be found in other recent publications, but it doesn’t make it more
-correct). Please use a term like ”comparisons to the control”.}
-\response{We agree and changed accordingly.}
-
-%done
-\comment{p. 7, right column, lines 7–10: A nice result that is not found in
-many publication: non-parametric approaches lack power (due to less
-assumptions being made). However, your results for the parametric
-models are all quite similar and as it stands the paper does not present
-a strong case for the use of GLMs; and a more balanced abstract would
-be appropriate.}
-\response{We agree and revised the abstract.}
-
-%done
-\comment{p. 7, Simulations: A suggestion: remove all diverging discussion of
-LOEC/NOEC, GLMs for multivariate data, GLMMs, sample size cal-
-culation and concentrate on discussing the simulation results. Or, alter-
-natively, discuss in much more detail that GLMMs may actually offer a
-difference approach for handling over-dispersion in binomial and count
-data (instead of considering completely different designs).}
-\response{We agree and pointed to GLMM as an alternative to deal with overdispersion in the methods section and discussion.}
-
-% ======================================================= %	
-\section{Response to Reviewer 2}
-\vspace{-2em}
-
-%done
-\comment{One point to consider in a resubmission is that GLMs aren't for use for all non-normal data - some qualifications are needed of when to use GLMs.
-Distributionally, GLMs replace the assumption of normality with the assumption that data come from the exponential family, which does not cover all situations.
-More critically, GLMs replace the assumption of equal variance with an assumption that the mean-variance relationship follows a pre-specified pattern, and violation of this assumption has similar impacts to violations of the equal-variance assumption in a linear model would have.
-Particular situations where GLMs are useful are binary data (known to be a quadratic mean-variance), count data (especially count data with lots of low counts, which typically cannot be transformed to normality effectively), and proportions constructed from counts.
-These are the examples highlighted in the text, and are of clear relevance in ecotoxicology, but the argument does not extend much
-more broadly than these examples - it is not a case of "your data aren't normal so you should use GLMs".
-It is more accurate to say "if you have binary, counts or proportions from counts you should use GLMs".
-I think this requires a modest adjustment to the phrasing of the abstract and introduction to better qualify the situations where GLMs are appropriate.
-e.g.
-where O'Hara and Kotze and Warton \& Hui are mentioned, this should be qualified to refer to counts and proportions (of counts) respectively.
-I  don't think the title needs changing, sufficient qualification elsewhere of the type of non-normality if interest should do the trick.}
-\response{We agree and revised our manuscript accordingly.
-We changed the subtitle (see also comment 2) and the abstract (see also comment 21).
-The main text is now more specific to counts and proportions from counts.}
-
-%done
-\comment{In the first paragraph, "positive" is used where "non-negative" is needed (i.e. zero is a possible value), also change bonded to bounded}
-\response{We fully agree and changed accordingly.}
-
-%done
-\comment{Page 2, second column - the parameterisation isn't equal to the Poisson model - the mean model is the same, I think that is what was meant.}
-\response{We edited this section (see also comment 6) and corrected this.}
-
-%done
-\comment{Page 2, second column - also, strictly speaking, not all the methods mentioned here are GLMs.
-Negative binomial regression ("NB2") is a GLM if the overdispersion parameter is fixed, otherwise it is a generalisation of GLM.
-Similarly, quasi-Poisson isn't really a GLM, but a related method.}
-\response{We agree and mention that quasi-Poisson is a related post hoc method. See also comment 7.}
-
-%done
-\comment{Section 2.2 - a description of what is meant by binomial data would be helpful at the start of this section.
-Perhaps for completeness a description of what is meant by counts would help at the start of section 2.1}
-\response{We added a brief description of a binomial random variable to the start of section 2.2.}
-
-%done
-\comment{Section 2.2.2 - it would be useful to also mention GLMMs with a random intercept term.
-Often with this sort of data there are extra sources of variation from one sample to the next that are not explained by the binomial assumption, and this is the most natural way to account for it.}
-\response{We added a reference to GLMM and \citep{warton_arcsine_2011} to this section.}
-
-%done
-\comment{Sections 2.1, 2.2: references to relevant articles or texts would be useful for GLMs, where the reader can find more details at a level appropriate for readership (e.g.
-a Zuur text?
-Quinn \& Keough?).
-Also point readers to supp 2 for code and a worked example.}
-\response{We are thankful for pointing us to the missing reference to supplement 2. We added this to the methods section and pointed to introductory textbooks at the end of the manuscript. }
-
-%done
-\comment{page 3, line 58, the parametric bootstrap}
+\comment{Rather than using a goo.gl link, the official stable CRAN URL should be used: https://CRAN.R-project.org/package=webchem}
 \response{We changed accordingly.}
@@
 %done
-\comment{Page 3, second column, line 30: Wald tests should be used with caution, when a mean estimate is zero (or close to it) they can have strange behaviour (related to the issue of separable data - parameter estimates and ses diverging, test statistics going to zero).}
-\response{We are thankful for this comment and added discussion of the Hauck-Donner effect.}
-
-%done
-\comment{Page 4, second column, lower means: it is noted that the means are lower when data are transformed.
-This is known as transformation bias, and occurs because the transform linear model is actually estimating something different - it is no longer trying to estimate mean response.
-It would be helpful to say this and maybe include a useful reference.
-This also gets at the issue of interpretability - a disadvantage of the transform approach is a loss of interpretability, because we are no longer modelling mean response.}
-\response{We added discussion of transformation bias, see also comment 4.}
-
-% done
-\comment{page 7 first column, line 7: bounds not bonds}
-\response{Thanks for pointing to this typo.}
-
-%done
-\comment{Section 4.1 the case study discussion was a little confusing.
-So is the point that you can get different results by linear models according to transformation, but that GLMs can resolve the problem of choice of transformation?
-(by replacing it with a decision about mean-variance relationship?)  I think a stronger point is that the GLM is usually a better fit to count data, no transformation can make data Gaussian when it has lots of zeros and small counts, but GLM is designed for such data.
-If diagnostic tools support this argument for your case study they could be included.}
-\response{Yes, we find it much easier to choose appropriate distributions than from the virtually infinite possibilities of transformations. 
-We agree, that GLMs give predictions that better fit to the data. 
-But note, that \citet{warton_many_2005} found that "\emph{a Gaussian model based on transformed abundances fitted
-data surprisingly well}" (see also comment 59). }
-
-%done
-\comment{page 7 column 1 line 27: it is not so much unreliable and biased - they are estimating a different parameter (in an unbiased fashion) - they estimate the mean of log(Ax+1) data, and a problem is that this has no natural interpretation in terms of the original data.
-It is fair to make the point that it is biased as an estimate of the mean response, but it would be helpful to explain why and how the estimators have problematic interpretations.}
-\response{We added discussion of transformation bias, see also comment 4.}
-
-%done
-\comment{page 7 column 1 line 50: a priori not a priory\\
-page 7 column 2 line 37: data are (data being the plural of datum)\\
-page 7 column 2 line 42: GLMs have (or GLM has)}
-\response{We are thankful for pointing to these errors and changed accordingly.}
-
-%done
-\comment{page 7 column 2 line 44: the higher power in Fig 5 seemed to be through the GLMs being more conservative, rather than due to greater efficiency.
-This issue could likely be fixed using a parametric bootstrap, as was done previously in figure 2.
-A related point is that the linear model had the advantage that it seemed to be better at maintaining nominal significance levels at small sample sizes, which should be mentioned.
-As the authors say however it did not have as good power (and also loses something in interpretability and biased estimation of means).
-Hence while the GLM has a number of advantages, combining it with param bootstrap at small sample sizes might be worth considering to address the Type I error issue.}
-\response{We agree and pointed to parametric bootstrap for situations of low sample sizes or estimates on the boundary of parameter space. See also comment 31}
-
-\comment{Supplement 2 is a useful addition, and it might be worth giving it a little more emphasis in the text to ensure ecotoxicologists wanting to try out new analyses give it a look.
-A couple of suggestions though:
-- there is no parametric bootstrap code there at the moment.
-This would be easy enough to do by using the mvabund package, calling manyglm (even on a univariate response) then anova(..., resamp="monte.carlo")  will do a parametric bootstrap.
-- you can construct residual plots for GLMs on this package, using the plot function on a fitted manyglm object.}
-\response{We are thankful for this comment. We refined Supplement 2 to incorporate all described models and methods and point the reader to it in the methods section.}
-
-% ======================================================= %	
-\section{Response to Reviewer 3}
-\vspace{-2em}
-
-%done
-\comment{With its simulations in an ecotox.
-context, the paper is a fine contribution to make researchers aware of the possibilities and advantages of a generalized linear models (GLM).
-What is lacking is pointing to the disaster of using GLM in an improper way, notably without adjustment for overdispersion.
-The paper and abstract can still be made more useful by given not just the recommendation for GLM but also to give advice and warnings of proper usage.
-For example, (almost) never use loglinear/Poisson regression without adjustment for overdispersion, be aware that a nicer looking model like the negative binomial may give inflated type I error, a case that is overcome in this paper by bootstrapping.
-The same applies to binomial/logit regression unless the experiment really consist of independent observations for each 0/1 result (and is not just a count with a predetermined maximum).
-In this light it is strange that the logistic/binomial case is treated differently from
-the loglinear/Poisson case! Should this not be repaired or at least given more warning. \\
-(1) Also, if the authors stick to the binomial/logistic without overdispersion, the normal model could be made powerful in the simulations by treating the error variance as known, because it is a known constant for the arcsine transformation.
-\\
-(2) Also, the simulation now use slight overdispersion whereas the example /typical data has large overdispersion.
-Add simulations with large overdispersion.
-The current ones can go to supplementary.
-\\
-(3) Without the use of a GLM equivalent of the Williams test all the advantage of the use of GLM in terms of power are gone.
-See the example.
-Discuss this ambiguity.
-You can perhaps use a bootstrap test based on (GLM?) monotonic regression or similar.
-I know some cues/leads in this direction.}
-\response{We rerun our simulations and now include also the poisson GLM with overdispersed data. This clearly shows the effect of not accounting for overdispersion and is discussed now. \\
-We stick to the binomial simulations without overdispersion, because we don't know if overdispersion is a common phenomena with ecotoxcicological data.
-However, we point the reader to overdispersion and methods to account for it. \\
-Our simulations mimicked the data of the case study and we clarified this (see also comment 48). \\
-Williams type contrasts are also available in a GLM framework. The choice of which contrasts to use really depends on the assumptions one wants to make and which questions to answer. 
-However, it does not affect our comparisons between methods.
-We stress this in the revised MS.
-}
-
-%done
-\comment{Page 1 Line 22 right column: a continuous proportion is fine for area cover and so on, but not for proportion of surviving animals where is just "k out of n" turn into a proportion.
-Discrete therefore.
-}
-\response{Thank you for pointing to this error. Of course, we meant discrete proportions and changed accordingly.}
-
-%done
-\comment{Line 58 right add ref to (Warton 2005).}
-\response{This reference fits very well and we added it.}
-
-%done
-\comment{Page 2 Line 40 : , for y>0 should for y>= 0 and  be on the next line without an indent after formula (if it is not new paragraph).}
-\response{We clarified this equation and moved this part of eqn. 1 to the text: \\
-\emph{"The factor $A$ was chosen in such way that $Ay$ equals 2 for the lowest non-zero abundance value (y)."}}
-
-%done
-\comment{Line 51 x\_i is undefined. The model appear to specify a simple linear regression or a control-one\_treatment model. Neither is used in the paper!!! Modify the notation therefore..}
-\response{We clarified this whole section. Eqn. 2 is in fact a simple linear regression, but with dummy coded predictors / treatments.}
-
-%done
-\comment{Line 59 Mention backtransformation here, which works fine for the confidence intervals, and not the backtransformed is the median on the original scale.
-If you want to backtransform to the original mean, use exp(mean+ 0.5*sigma2).
-(Aitchison \& Brown 1969) page 8 (Jongman et al.
-1995)page 19.
-}
-\response{We agree and added a discussion of backtransformation bias. See also comment 4.}
-
-%done
-\comment{Page 3 line 17 n= 4.10 = 40 what does the mean here/ what is the purpose?}
-\response{We are sorry for this error. It was a remnant of an earlier version of the manuscript - removed.}
-
-%done
-\comment{Line 41 : give more attention to the condition for the binomial: independent observations of each of the successes (k) out of the number of experiments (n).
-And: thus give attention to overdispersion in this case as well, particularly as you found nasty things for the LR of  NB.}
-\response{We now point the reader to overdispersed binomial data, though our simulation design does not cover this case.}
-
-%done
-\comment{Line 28 right. Ref for Dunnett contrasts.}
-\response{We agree and added reference to the original description of Dunnett (1955).}
-
-%done
-\comment{Line 59 right. Only slight overdispersion, whereas the example data have large  overdispersion. Add large overdispersion.}
-\response{We clarified this section. We now specify the overdisperion in terms of $\kappa$ in both, the case study and the simulation design. 
-The simulated data mimicked the case study and such data is common for mesocosm studies in ecotoxicology.}
-
-%done
-\comment{Page 4  L 19 ( missing.}
-\response{We fixed the wrong citation type.}
-
-%done
-\comment{Line 56 left to 2 right.
-Four p-values: what to believe?
-Your simulations (although with small overdispersion) show it.
-The LR of NB cannot be trusted.
-So the 0.016 is out.
-The remaining tests all show about the same p-value if you interpret statistics properly.
-Please note that the distinction between significant and non-signicant (here p = 0.061 and 0.042) in itself is statistically insignificant.
-(Gelman \& Stern 2006).}
-\response{We fully agree with the reviewer and cautioned that p-values should not be overinterpreted by ecotoxicologists (citing Gelman \& Stern 2006). 
-Parameter estimates (and their uncertainty) give much more information - on which we also emphasized in Figure 1.}
-
-%done
-\comment{Line 4 right : the backtransformed mean gives a median on the original scale, a value that is more interesting than the mean (generally for skew data),  and for skew right always smaller than the mean.}
-\response{We agree and added a discussion of back-transformation bias. See also comments 4 and 44.}
-
-%done
-\comment{Section 3.2
-Always start with the Type I error results because the power results are without meaning if the Type I error is inflated.
-This also applies the order of the figs: interchange the two rows of subfigs.
-}
-\response{We agree and changed the order (first type I error, then power) in the text, as well as in the figures.}
-
-%done
-\comment{21 right: remove GLM\_nb in this sentence as GLM\_nb must be discarded due to inflated Type I error.}
-\response{We agree and changed accordingly.}
-
-%done
-\comment{L32 right.
-", but this." Although you should reorder the results, I mention that the "but this" should at least be "This could fortunately be …"[and we could not have known for sure (apart from the general warnings about LR) without the simulations] This inflated error for GLM\_nb must receive more attention.
-It is an interesting result of this study: you must to bootstrap when using NB.}
-\response{We reordered and edited the respective results and discussion sections and emphasized this.}
-
-%done
-\comment{Page 5
-Fig.
-2 legend.
-Add that n =100 simulations.
-Add the interpretation that  GLM\_nb has inflated type I error, and that its line for power is just added for completeness and should not be interpreted as an estimate of the true power.}
-\response{We unified the simulations and generated simulated 1000 data sets for each type of scenario (see comment 14). Moreover, we added \\
-\emph{"Power levels for models with inflated type I error are shown for completeness."}\\
-to figure captions 2 and 3.}
-
-%done
-\comment{Page 6
-Fig.5 LM for 0.8 has a high Type I error. Does it deviate significantly from 0.05? [same question for the high Type I for GLM\_nb in fig 2]. Use confidence interval for binomial p = 0.05 n = 100.}
-\response{We refined our simulations (see comment 14, )}
-
-%done
-\comment{Case study : 
-You can check whether the difference between log(2y+1) and log(Ay+1) with A = 0.182 already causes the difference.
-(it probably does not).
-The important difference is the use of the Willems test!
-Without the use of a GLM equivalent of the Williams test all the advantage of the use of GLM in terms of power are gone!
-See the report of the Williams test below.}
-\response{The reviewer is correct that the Williams test has higher power (if the assumptions are met). 
-Williams type contrast are also available in a GLM framework and the choice of contrast to use really depends on the assumptions one wants to make and which questions to answer. 
-This explains the difference between the results of Brock 2015 and us, but it does not affect our comparisons between methods.}
-
-%done
-\comment{Page 7 L35 Move ref to Warton to after data.}
+\comment{Title in title style (“retrieve” should be capitalized)}
 \response{We changed accordingly.}
@@
 %done
-\comment{L45 Add a reference to (ter Braak \& Šmilauer 2014) who advocate the use of the transformation approach in a multivariate (dimension reduction) context.}
-\response{We agree and added reference to ter Braak \& Šmilauer 2014.}
+\comment{section, subsection, etc. should be in sentence style (see http://www.jstatsoft.org/about/submissions), e.g., change “4. Use Cases” to “4. Use cases”}
+\response{All sections etc. are not in sentence style.}
 
 %done
-\comment{Line 1 right (like GLMs) -\textgreater (like GLMs and the Williams test)}
-\response{We did not investigate the differences between multiple comparison procedures (MCP). 
-We used Dunnett constrasts in all simulations, because only these allow individual comparisons between treatments and the control. 
-For a comparison of MCP see \citet{jaki_statistical_2013}, who recommend a combination of both, Dunnett and Williams constrasts.
-}
+\comment{The code presented in the manuscript should not contain comments within the verbatim code. Instead the comments should be made in the normal LaTeX text.}
+\response{We removed all comments from the code blocks and clarified the text.}
 
 %done
-\comment{Line 10-40 Again: discuss Type I error first. Then no mention of GLM\_nb in power.}
-\response{We agree and changed accordingly. See also comment 52.}
+\comment{For the code layout in R publications, the input should use the text width (up to 76 or 77 characters) and be indented by two spaces, e.g.,}
+\response{We fixed the width to 76 characters and removed excessive indentation.}
 
 %done
-\comment{Line 18 right.
-Add after "non-normal data' the point the importance of proper accounting for under and overdispersion when using GLM.
-What is lacking is pointing to the disaster of using GLM in an improper way, notably without adjustment for overdispersion.}
-\response{We rerun our simulations and also included the Poisson GLM. 
-The results for Poisson GLM are now also displayed in Figures 2+3 and we discuss the consequences of failing to account for overdisperion.}
-
-%done
-\comment{Line 41 right size of 9 -\textgreater size of 6-9 (?)}
-\response{We rerun our simulations and changed this to \\
-\emph{for all simulated sample sizes}.}
-
-%done
-\comment{Line 44 right the transformation -\textgreater the LM}
+\comment{For R-related manuscripts: The first argument of data() and library() should always be quoted, e.g., library("foo"). [' library(webchem)', ' data(jagst)', ' data(lc50)']}
 \response{We changed accordingly.}
 
+\comment{ If using "e.g." and "i.e." add a comma after the period to keep LaTeX from interpreting them as the end of a sentence, i.e.: "e.g., " and "i.e., ".}
+\response{We added a backslash ("i.e.\textbackslash " or "e.g.\textbackslash"), which also prevents this behaviour.}
+
+\comment{There should not be further footnote-style annotations in tables; these should all be placed in the caption.}
+\response{We removed the footnote-style annotation from table 1.}
+
+\comment{As a reminder, please make sure that: proglang, pkg and code have been used for highlighting throughout the paper (including titles and references), except where explicitly escaped.}
+\response{We checked thoroughly the manuscript.}
+
+\comment{Springer-Verlag (not: Springer)}
+\response{We changed accordingly.}
+
+\comment{Please make sure that all software packages are cited properly.}
+\response{We checked thoroughly and updated the citations. }
+
+\comment{Please make sure that the files needed to replicate all code/examples within the manuscript are included in a standalone replication script.}
+\response{These are given in the accompanying file "article.R".}
+
 %% --------------------------------
-\newpage
-\bibliography{refs}
-\bibliographystyle{spbasic}
+% \newpage
+% \bibliography{refs}
+% \bibliographystyle{spbasic}
 

### 2016-04-24 revision1/article.R
revised R script

@@ -1,7 +1,7 @@
 # -------------------------------------------------------------------------
 # Supplemental R Code to reproduce the resutls of
-# Szöcs, Schäfer (2016). webchem: an R Package to retrieve Chemical Information from the Web
-# Date: 04.03.2016
+# Szöcs, Schäfer (2016). webchem: an R Package to Retrieve Chemical Information from the Web
+# Date: 24.04.2016
 # ------------------------------------------------------------------------
 
 
@@ -10,27 +10,29 @@ install.packages("webchem")
 
 
 # load packages -----------------------------------------------------------
-library(webchem)
-library(ggplot2)
+library("webchem")
+library("ggplot2")
 
 
 # load datasets -----------------------------------------------------------
 # load jagst dataset
-data(jagst)
+data("jagst")
 # print first 6 lines of data
 head(jagst)
 
 # lc50 dataset
-data(lc50)
+data("lc50")
 head(lc50)
 
 
 # Use Case 1: Query identifiers -------------------------------------------
 # unique substance names
 subs <- unique(jagst$substance)
+
 # search ETOX IDs, keeping only the best match
 ids <- get_etoxid(subs, match = 'best')
 head(ids)
+
 # use this id to query information from ETOX
 etox_data <- etox_basic(ids$etoxid)
 # extract CAS numbers from retrieved data
@@ -39,10 +41,12 @@ head(etox_cas)
 
 
 # query other identifiers from other resources
+
 # query SMILES from PubChem
 cids <- get_cid(etox_cas)
 pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
 pc_smiles <- smiles(pc_data)
+
 # ChemSpider needs a security token
 # this is a webchem specific token. 
 # Please use it only for reproduction / testing.
@@ -52,8 +56,7 @@ csids <- get_csid(etox_cas, token = token)
 cs_data <- cs_compinfo(csids, token = token)
 cs_inchikey <- inchikey(cs_data)
 
-
-# combine in one data.frame
+# combine into single data.frame
 res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
                   cid = pc_data$CID, inchikey = cs_inchikey, 
                   csid = cs_data$csid, 
@@ -78,7 +81,9 @@ lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
[…]

### 2016-09-20 build_article.R
do not ignore gitignore

@@ -1,0 +1,4 @@
+require(knitr)
+require(tools)
+knit('revision1/article.Rnw')
+texi2pdf('revision1/article.tex')

### 2016-09-20 revision2/article.Rnw
revised from reviewer comments

--- 2016-04-23 revision1/article.Rnw
@@
 Manual searching and downloading of such data is time-consuming and error-prone.  
-
 We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query chemical data from currently 11 web sources. 
@@
 The data are automatically imported into an \proglang{R} object and can directly be used in subsequent analyses.
-
 \pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web sources.
@@
 \section[Introduction]{Introduction}
-Before each statistical analysis, data cleaning is often required to ensure a good data quality.
+Before each statistical analysis, data cleaning is often required to ensure good data quality.
 Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
@@
 
-When dealing with chemical datasets (e.g.\ environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
+When dealing with chemical data sets (e.g.\ environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
 Several web sources provide chemical names or link them to unique codes (see also section \emph{Data sources} below).
@@
 To simplify, robustify and automate this task, i.e.\ to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
-\proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.
+\proglang{R} is one of the most widely used software environments for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.
 
@@
 The \pkg{webchem} package is written entirely in \proglang{R} and available under a MIT license.
-The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://CRAN.R-project.org/package=webchem}).
-\pkg{webchem} is part of the rOpenSci project (\url{https://ropensci.org/}), which aims at fully reproducible data analysis.
+The development repository is hosted on \citet{github} and a stable version is released on the official \proglang{R} repository \citep{cran}.
+\pkg{webchem} is part of the rOpenSci project \citep{boettiger2015building}, which aims at fully reproducible data analysis.
 
-\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via Travis-CI and AppVeyor; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
+\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via \citet{travis-ci} and \citet{appveyor}; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
 
@@
 \pkg{webchem} is friendly to the resources of data providers. 
-Between each request there is a time-out of 0.3 to 2 seconds depending on the datasource. 
-Therefore, processing of larger datasets can take some time, but still represents a major improvement compared to manual lookup.
+Between each request there is a time-out of 0.3 to 2 seconds depending on the data source. 
+Therefore, processing of larger data sets can take some time, but still represents a major improvement compared to manual look up.
 We provide a link to the \emph{Terms of Use} of data providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
@@
   \item[SRC Physprop \citep{physprop}]{Contains physical properties for over 41,000 chemicals.
-  Physical properties collected from a wide variety of sources including experimental and modelled values.}
+  Physical properties collected from a wide variety of sources including experimental and modeled values.}
   \item[PubChem \citep{Kim_2016}]{PubChem is a public repository for information on chemical substances, providing identifiers, properties and synonyms.
@@
 
-The package is under active development. The latest development version is available from GitHub and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
+The package is under active development. The latest development version is available from GitHub and also permanently available at \citet{zenodo}.
 This document has been created using \pkg{webchem} version 0.1.
 
-\subsection[Sample datasets]{Sample datasets}
-To demonstrate the capabilities of  \pkg{webchem} we use two small publicly available real world datasets.
-The datasets are only for demonstration purposes, have been slightly preprocessed (not shown) and are available through the package.
+\subsection[Sample data sets]{Sample data sets}
+To demonstrate the capabilities of  \pkg{webchem} we use two small publicly available real world data sets.
+The data sets are only for demonstration purposes, have been slightly preprocessed (not shown) and are available through the package.
 
-(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst, Germany, sampled in 2013.
+(i) \code{jagst}: This data set comprises environmental monitoring data of organic substances in the river Jagst, Germany, sampled in 2013.
 The data is publicly available and can be retrieved from \citetalias{lubw_2016}.
 It comprises concentrations  (in $\mathrm{\mu g~/~L}$) of  34 substances  on 13 sampling occasions.
-First we load the dataset and take a look at the first six rows:
+First we load the data set and take a look at the first six rows:
 
@@
 
-This dataset identifies substances only by substance names. Values below the limit of quantification (LOQ) are indicated by a qualifier column.
+This data set identifies substances only by substance names. Values below the limit of quantification (LOQ) are indicated by a qualifier column.
 
@@
 
-As were are dealing with German substance names we start to query ETOX for CAS registry numbers.
+As we are are dealing with German substance names we start to query ETOX for CAS registry numbers.
 A common work flow when dealing with web resources is to 1) query a unique identifier of the source, 2) use this identifier to retrieve additional information and 3) extract the parts that are needed from the \proglang{R} object \citep{Chamberlain_Szocs_2013}.
@@
 Only three substances could not be found in ETOX. 
-Here we specify that only the \emph{'best'} match (in terms of the Levenshtein distance between query and results) is returned. 
+Here we specify that only the \emph{`best'} match (in terms of the Levenshtein distance between query and results) is returned. 
 A manual check confirms appropriate matches. 
-Other options include: \emph{'all'} - returns all matches; \emph{'first'} - returns only the first match (not necessarily the best match); \emph{'ask'} - this enters an interactive mode, where the user is asked for a choice if multiple matches are found and \emph{'na'} which returns \code{NA} in case of multiple matches.
+Other options include: \emph{`all'} - returns all matches; \emph{`first'} - returns only the first match (not necessarily the best match); \emph{`ask'} - this enters an interactive mode, where the user is asked for a choice if multiple matches are found and \emph{`na'} which returns \code{NA} in case of multiple matches.
 
@@
 # Please use it only for reproduction / testing.
-token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
+token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
 @
@@
 or ChemSpider
-<<chemspider, message = FALSE, warning=FALSE>>=
+<<chemspider, message=FALSE, warning=FALSE>>=
 csids <- get_csid(etox_cas, token = token)
@@
 
-Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
-We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g.\ the \code{lc50} dataset based on CAS.
+Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider web page. 
+We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other data sets, e.g.\ the \code{lc50} data set based on CAS.
 
@@
 @
-\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
+\caption{Identifiers for the jagst data sets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
 \label{tab:comptable}
@@
 Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
-Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the \code{lc50} dataset.
+Answering this question would require tedious look up of insecticide groups for each of the 124 CAS numbers in the \code{lc50} data set.
 The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried using CAS numbers with \pkg{webchem}:
@@
 
-To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply this to the retrieved data:
+To extract the chemical group from the retrieved data set, we need to write a simple extractor function and apply this to the retrieved data:
 
@@
 
+\subsection[Legal information]{Legal information}
+Legal information is particularly of interest if concentrations exceed national thresholds.
+In the European Union (EU) the Water Framework Directive (WFD) defines Environmental Quality Standards (EQS).
+Similarly, in the U.S. and Canadian EPA and the WHO defines Quality Standards.
+Information on these standards can be queried with \pkg{webchem} from the PAN Pesticide Database (using \code{pan\_query()}) and from ETOX (using \code{etox\_targets()}).
+
+In this example we search for the minimum EQS for the EU for the compounds in the \code{jagst} data set, join these with measureds concentrations and check if there have been any exceedances.
+
+We re-use the above queried ETOX-IDs to get further information from ETOX, namely the MAC-EQS:
+<<search_targests, message=FALSE, warning=FALSE>>=
+eqs <- etox_targets(ids$etoxid)
+ids$mac <- sapply(eqs, function(y){
+  if (length(y) == 1 && is.na(y)) {
+    return(NA) 
+  } else {
+    res <- y$res
+    min(res[res$Country_or_Region == 'EEC / EU' & 
+              res$Designation == 'MAC-EQS', 'Value_Target_LR'])
+  }
+})
+@
+
+Again, the returned information is humongous and we encourage the users to study the returned objects and description of the data source.
+Here the column \code{Designation} defines the type of EQS and \code{Value_Target_LR} contains the value.
+Unfortunately, we found MAC-EQS values only for 5 substances:
+
+<<mac>>=
+(mac <- with(ids, ids[!is.na(mac) & is.finite(mac), 
+                      c('etoxid', 'query', 'mac')]))
+@
+
+The \code{get_etoxid()} function used to search ETOX-IDs returns also the original substance name (\code{query}),
+so that we can easily join the table with MAC values with the measurements table :
+<<>>=
+jagst_eqs <- merge(jagst, mac, by.x = 'substance', by.y = 'query')
+head(jagst_eqs)
+@
+
+Finally, we can compare the measured value with the MAC, which reveals that there have been no exceedances of these 5 compounds.
+
 \subsection[Utility functions]{Utility functions}
@@
 Within the \proglang{R} ecosystem, there are only few similar projects:
-\pkg{rpubchem} \citep{rpubchem_2014} and \pkg{ChemmineR} \citep{chemminer_2008} both provide an interface to PubChem that is more extensive than the current one in \pkg{webchem}.
+\pkg{rpubchem} \citep{rpubchem_2014} provides an interface PubChem.
+Similarly, \pkg{ChemmineR} \citep{chemminer_2008}, a mature chemo-informatics package, also provides an interface to Pubchem. 
+\pkg{webchem} does not provide any chemo-informatic functionality, but integrates access to many data sources.
 \pkg{WikidataR} \citep{wikidatar_2016} provides an interface to wikidata that could be used to retrieve chemical data from Wikipedia.
@@
 To be usable for other researchers chemical compounds should be properly identified, not only by chemical names but also with accompanying identifiers like InChIKey, SMILES and authority-assigned identifiers.
-\pkg{webchem} provides an easy way to create such meta tables as shown in Table \ref{tab:comptable}.
-
-\pkg{webchem} facilitates chemical data availability to researchers.
-However, data quality is also crucial for data analysis \citep{Stieger_2014}. 
-Ensuring good quality requires additional effort and methods to be developed.
+\pkg{webchem} provides an easy way to create such meta tables as shown in Table \ref{tab:comptable} and facilitates chemical data availability to researchers.
+However, good quality of data is crucial for every analysis \citep{Stieger_2014} and additional effort and methods are needed to validate data quality.
 
@@
 We thank Johannes Ranke and Daniel Münch for their contributions to the webchem package, as well as all users who provided feedback and feature requests. 
-We are grateful to Zacharias Steinmetz for the valuable comments on the manuscript.
+We are grateful to Zacharias Steinmetz for valuable comments on the manuscript.
 

### 2016-09-20 revision2/article.tex
revised from reviewer comments

--- 2016-09-20 revision2/article.Rnw
@@
 \defcitealias{lubw_2016}{LUBW (2016)}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-
-<<global_options, include=FALSE>>=
-library(knitr)
-knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
-knitr::render_sweave()
-knitr::opts_chunk$set(prompt = TRUE,
-                      comment = NA,
-                      highlight = FALSE,
-                      dev.args = list(pointsize = 11),
-                      # fig.width = 4.9,
-                      # fig.height = 3.675,
-                      fig.align = 'center',
-                      out.format = 'sweave',
-                      tidy = FALSE,
-                      fig.path = '')
-# 
-hook.c <- function(x, options) {
-  paste0('\\begin{CodeChunk}\n', x, '\\end{CodeChunk}')
-}
-hook.i <- function(x, options) {
-  paste0(c('\\begin{CodeInput}', knitr:::hilight_source(x, 'latex', options), '\\end{CodeInput}', ''), collapse = '\n')
-}
-hook.s <- function(x, options) {
-  paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
-}
-knitr::knit_hooks$set(chunk = hook.c,
-               source = hook.i,
-               output = hook.s,
-               message = hook.s,
-               warning = hook.s,
-               plot = knitr::hook_plot_tex)
-
-options(prompt = "R> ", continue = "+  ", width = 76, useFancyQuotes = FALSE)
-@
 
@@
 
-<<install, eval=FALSE>>=
-install.packages("webchem")
-library("webchem")
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> install.packages("webchem")
+R> library("webchem")
+\end{CodeInput}
+\end{CodeChunk}
 
-<<load, echo=FALSE, eval=TRUE>>=
-library(webchem)
-library(ggplot2)
-library(xtable)
-@
+\begin{CodeChunk}
+\end{CodeChunk}
 
@@
 
-<<loadjagst>>=
-data("jagst")
-head(jagst)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> data("jagst")
+R> head(jagst)
+\end{CodeInput}
+\begin{CodeOutput}
+        date          substance value qual
+1 2013-01-04 2,4-Dimethylphenol 0.006    <
+2 2013-01-29 2,4-Dimethylphenol 0.006    <
+3 2013-02-26 2,4-Dimethylphenol 0.006    <
+4 2013-03-26 2,4-Dimethylphenol 0.006    <
+5 2013-04-23 2,4-Dimethylphenol 0.006    <
+6 2013-05-22 2,4-Dimethylphenol 0.006    <
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<load_lc50>>=
-data("lc50")
-head(lc50)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> data("lc50")
+R> head(lc50)
+\end{CodeInput}
+\begin{CodeOutput}
+       cas        value
+4  50-29-3    12.415277
+12 52-68-6     1.282980
+15 55-38-9    12.168138
+18 56-23-5 35000.000000
+21 56-38-2     1.539119
+36 57-74-9    98.400000
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<get_etoxid, message=FALSE, cache=TRUE>>=
-subs <- unique(jagst$substance)
-ids <- get_etoxid(subs, match = 'best')
-head(ids)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> subs <- unique(jagst$substance)
+R> ids <- get_etoxid(subs, match = 'best')
+R> head(ids)
+\end{CodeInput}
+\begin{CodeOutput}
+  etoxid                           match distance                  query
+1   8668     2,4-Dimethylphenol ( 8668 )        0     2,4-Dimethylphenol
+2   8494 4-Chlor-2-methylphenol ( 8494 )        0 4-Chlor-2-methylphenol
+3   <NA>                            <NA>     <NA>     4-para-nonylphenol
+4   8397                Atrazin ( 8397 )        0                Atrazin
+5   7240                 Benzol ( 7240 )        0                 Benzol
+6   7331        Desethylatrazin ( 7331 )        0        Desethylatrazin
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
-<<token, echo=FALSE>>=
-# this is a webchem specific token. 
-# Please use it only for reproduction / testing.
-token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
-@
+\begin{CodeChunk}
+\end{CodeChunk}
 
-<<pubchem, message = FALSE, warning=FALSE>>=
-cids <- get_cid(etox_cas)
-pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
-pc_smiles <- smiles(pc_data)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> cids <- get_cid(etox_cas)
+R> pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+R> pc_smiles <- smiles(pc_data)
+\end{CodeInput}
+\end{CodeChunk}
 
 or ChemSpider
-<<chemspider, message=FALSE, warning=FALSE>>=
-csids <- get_csid(etox_cas, token = token)
-cs_data <- cs_compinfo(csids, token = token)
-cs_inchikey <- inchikey(cs_data)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> csids <- get_csid(etox_cas, token = token)
+R> cs_data <- cs_compinfo(csids, token = token)
+R> cs_inchikey <- inchikey(cs_data)
+\end{CodeInput}
+\end{CodeChunk}
 
 Finally, we combine the queried data into one data.frame
-<<table>>=
-res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-  cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
-  stringsAsFactors = FALSE)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
++    cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
++    stringsAsFactors = FALSE)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 \centering
-<<print_table, echo=FALSE, results = 'asis'>>=
-tmp <- res[1:6, ]
-tmp$smiles <- ifelse(is.na(tmp$smiles), NA, paste0(substr(tmp$smiles, 1, 7), '...'))
-tmp$inchikey <- ifelse(is.na(tmp$inchikey), NA, paste0(substr(tmp$inchikey, 1, 7), '...'))
-tmp[is.na(tmp)] <- '-'
-colnames(tmp) <- c("Name", "CAS", "SMILES","CID", "InChIKey", "CSID")
-print(xtable(tmp),
-  floating = FALSE,
-  hline.after = NULL,
-  include.rownames = FALSE,
-  add.to.row = list(pos = list(-1,0, 6),
-                    command = c('\\toprule\n','\\midrule\n','\\bottomrule\n')),
-  sanitize.text.function = function(x){x}
-  )
-@
+\begin{CodeChunk}
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Tue Sep 20 16:11:04 2016
+\begin{tabular}{llllll}
+  \toprule
+Name & CAS & SMILES & CID & InChIKey & CSID \\ 
+  \midrule
+2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & - & - \\ 
+  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & - & - \\ 
+  4-para-nonylphenol & - & - & - & - & - \\ 
+  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & - & - \\ 
+  Benzol & 71-43-2 & C1=CC=C... & 241 & - & - \\ 
+  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & - & - \\ 
+   \bottomrule
+\end{tabular}
+\end{CodeChunk}
 \caption{Identifiers for the jagst data sets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
@@
 
-<<aw_query, message = FALSE>>=
-aw_data <- aw_query(lc50$cas, type = 'cas')
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> aw_data <- aw_query(lc50$cas, type = 'cas')
+\end{CodeInput}
+\begin{CodeOutput}
+Warning: closing unused connection 35 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=67-66-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 34 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=75-25-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 33 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=108-88-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 32 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=127-18-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 31 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=5915-41-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 30 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=886-50-0&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 29 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=122-34-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 28 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=60207-90-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 27 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=139-40-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 26 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=23103-98-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 25 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1763-23-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 24 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=335-67-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 23 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=307-24-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 22 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=375-22-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 21 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=40487-42-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 20 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=66246-88-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 19 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=51218-45-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 18 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=67129-08-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 17 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=57837-19-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 16 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=94-74-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 15 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=34123-59-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 14 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=28159-98-0&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 13 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=83164-33-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 12 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=134-62-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 11 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=75-09-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 10 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=30125-63-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 9 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=6190-65-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 8 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=71-43-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 7 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1912-24-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 6 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1570-64-5&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\begin{CodeOutput}
+Warning: closing unused connection 5 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=105-67-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<subactivty, message = FALSE>>=
-igroup <- sapply(aw_data, function(y) y$subactivity[1])
-igroup[1:3]
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> igroup <- sapply(aw_data, function(y) y$subactivity[1])
+R> igroup[1:3]
+\end{CodeInput}
+\begin{CodeOutput}
+                                  50-29-3 
+            "organochlorine insecticides" 
+                                  52-68-6 
+               "phosphonate insecticides" 
+                                  55-38-9 
+"phenyl organothiophosphate insecticides" 
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 \begin{figure}[ht]
-<<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
-lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
-              ifelse(grepl('neonicotinoid', igroup), 'Neo-\nnicotinoids', 
-                ifelse(grepl('pyrethroid', igroup), 'Pyrethroids', 
-                  ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
-                         'other'))))
-lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", "Organo-\nphosphates", "Neo-\nnicotinoids", "other"))
+\begin{CodeChunk}
 
-p <- ggplot(lc50, aes(x = type, y = value)) +
-  geom_boxplot(fill = 'grey75') + 
-  geom_jitter(width = 0.5) +
-  scale_y_log10() +
-  labs(y = expression(LC[50~'D.magna, 48h']), x = '') +
-  theme_bw() +
-  theme(text = element_text(size = 16))
-p
-@
+{\centering \includegraphics[width=0.7\textwidth]{plot_lc50-1}
+
+}
+
+\end{CodeChunk}
 \caption{Toxicity of different pesticide groups. LC\textsubscript{50} values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
@@
 
-<<qsar, message=FALSE, cache=TRUE, warning=FALSE>>=
-pp_data <- pp_query(lc50$cas)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> pp_data <- pp_query(lc50$cas)
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-<<qsar2, warnings = FALSE>>=
-lc50$logp <- sapply(pp_data, function(y) {
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$value[y$prop$variable == 'Log P (octanol-water)']
-})
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> lc50$logp <- sapply(pp_data, function(y) {
++    if (length(y) == 1 && is.na(y))
++      return(NA)
++    y$prop$value[y$prop$variable == 'Log P (octanol-water)']
++  })
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 \begin{figure}[ht]
-<<plot_qsar, echo=FALSE, out.width='0.7\\textwidth', fig.height=6, warning=FALSE>>=
-# model
-mod <- lm(log10(value) ~ logp, data = lc50)
-coefs <- round(coef(mod), 2) # coeficients
-rmse <- round(summary(mod)$sigma, 2) # RMSE
+\begin{CodeChunk}
 
-# check data types
-logp_type <- sapply(pp_data, function(y){
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$type[y$prop$variable == 'Log P (octanol-water)']
-})
-tabs <- table(logp_type)
-sna <- sum(is.na(logp_type))
+{\centering \includegraphics[width=0.7\textwidth]{plot_qsar-1}
 
-p <- ggplot(lc50, aes(x = logp, y = log10(value))) +
-  geom_point() +
-  geom_smooth(method = 'lm') +
-  labs(x = expression(log~P[oct/wat]), y = expression(log~LC[50])) +
-  theme_bw()
-p
-@
+}
+
+\end{CodeChunk}
 \caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
-Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
-Blue line indicates the regression model ($\mathrm{log~LC}_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]}~\mathrm{log} P$, RMSE = \Sexpr{rmse}).}
+Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
+Blue line indicates the regression model ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log} P$, RMSE = 1.45).}
 \label{fig:fig3}
@@
 We re-use the above queried ETOX-IDs to get further information from ETOX, namely the MAC-EQS:
-<<search_targests, message=FALSE, warning=FALSE>>=
-eqs <- etox_targets(ids$etoxid)
-ids$mac <- sapply(eqs, function(y){
-  if (length(y) == 1 && is.na(y)) {
-    return(NA) 
-  } else {
-    res <- y$res
-    min(res[res$Country_or_Region == 'EEC / EU' & 
-              res$Designation == 'MAC-EQS', 'Value_Target_LR'])
-  }
-})
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> eqs <- etox_targets(ids$etoxid)
+R> ids$mac <- sapply(eqs, function(y){
++    if (length(y) == 1 && is.na(y)) {
++      return(NA) 
++    } else {
++      res <- y$res
++      min(res[res$Country_or_Region == 'EEC / EU' & 
++                res$Designation == 'MAC-EQS', 'Value_Target_LR'])
++    }
++  })
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-<<mac>>=
-(mac <- with(ids, ids[!is.na(mac) & is.finite(mac), 
-                      c('etoxid', 'query', 'mac')]))
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> (mac <- with(ids, ids[!is.na(mac) & is.finite(mac), 
++                        c('etoxid', 'query', 'mac')]))
+\end{CodeInput}
+\begin{CodeOutput}
+   etoxid       query    mac
+4    8397     Atrazin  2.000
+5    7240      Benzol 50.000
+11   8836     Irgarol  0.016
+12   7442 Isoproturon  1.000
+29   8756   Terbutryn  0.034
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 so that we can easily join the table with MAC values with the measurements table :
-<<>>=
-jagst_eqs <- merge(jagst, mac, by.x = 'substance', by.y = 'query')
-head(jagst_eqs)
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> jagst_eqs <- merge(jagst, mac, by.x = 'substance', by.y = 'query')
+R> head(jagst_eqs)
+\end{CodeInput}
+\begin{CodeOutput}
+  substance       date  value qual etoxid mac
+1   Atrazin 2013-09-10 0.0068    =   8397   2
+2   Atrazin 2013-10-08 0.0072    =   8397   2
+3   Atrazin 2013-03-26 0.0040    =   8397   2
+4   Atrazin 2013-04-23 0.0048    =   8397   2
+5   Atrazin 2013-11-05 0.0036    =   8397   2
+6   Atrazin 2013-07-16 0.0052    =   8397   2
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<check_format>>=
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
-is.cas('64-17-6')
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
+\end{CodeInput}
+\begin{CodeOutput}
+Hyphens not at position 15 and 26.
+\end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\begin{CodeInput}
+R> is.cas('64-17-6')
+\end{CodeInput}
+\begin{CodeOutput}
+Checksum is not correct! 5 vs. 6
+\end{CodeOutput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\end{CodeChunk}
 
 or web resources like ChemSpider
-<<check_web>>=
-is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
-  type = 'chemspider')
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
++    type = 'chemspider')
+\end{CodeInput}
+\begin{CodeOutput}
+[1] FALSE
+\end{CodeOutput}
+\end{CodeChunk}
 

### 2016-09-20 revision2/response/reponse.tex
revised from reviewer comments

% ======================================================= %
% Document: TEMPLATE FOR RESPONSES TO REVIEWERS
% Author: Andrea Ballatore
% Date: Jan 7, 2013
% Source: https://raw.githubusercontent.com/ucd-spatial/Datasets/master/tex_response_to_reviewers_template/responses_to_reviewers.tex
% Modified by Eduard Szöcs, 10.03.2015
% ======================================================= %
\documentclass[12pt]{article}

% packages
\usepackage{graphicx}
\usepackage{url}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{color}
\definecolor{mygray}{gray}{0.6}
\usepackage[utf8]{inputenc}
\usepackage[onehalfspacing]{setspace}
\usepackage[
	round,	%(defaultage in the main file and \input ) for round parentheses;
	colon,	% (default) to separate multiple citations with colons;
	authoryear,% (default) for author-year citations;
	sort,		% orders multiple citations into the sequence in which they 
]{natbib}					
\usepackage[%disable
	]{todonotes}

\usepackage{anysize}
\marginsize{2.5cm}{2.5cm}{1.5cm}{2.5cm}

% macros
% add a counter
\newcounter{cN}
\setcounter{cN}{0}

\newcommand{\comment}[1]{
	\vspace{2em} 
	\refstepcounter{cN} % incrment counter
	\noindent \hangindent=0em \textbf{\textcolor{Maroon}{\uline{Comment \thecN}:~}}\emph{"#1"}
	}

\newcommand{\response}[1]{
	\\[0.25em] 
	\hangindent=2.3em \textbf{\textcolor{NavyBlue}{\uline{Response}:~}}#1 
	}

\usepackage[normalem]{ulem}
\definecolor{darkred}{rgb}{1,.6,.6}
\DeclareRobustCommand\problemline{\bgroup\markoverwith{\textcolor{darkred}{\rule[-0.9ex]{4pt}{3pt}}}\ULon}
\DeclareRobustCommand{\problem}[1]{\problemline{#1}} % soul
\setcounter{secnumdepth}{-1}

\begin{document}
% ======================================================= %
\title{Response to the editorial team\\~\\JSS 2581: Szocs, Schafer \\ webchem: An R Package to Retrieve Chemical Information from the Web}

\author{Eduard Szöcs and Ralf B. Schäfer}

\maketitle
% ======================================================= %
\noindent Dear editorial team and reviewers,\\

\noindent
We thank the reviewer for checking the functionality of our package and also checking the appropriateness of the documentation and providing valuable comments on our manuscript.
We incorporated all the suggested changes and proof read the manuscript.

\noindent
Please find below detailed description of the changes made and responses to specific comments.

\vspace{2em}
\hfill Kind regards,

\hfill Eduard Szöcs and Ralf B. Schäfer
\newpage

% ======================================================= %

\section{Reviewer A}

We are thankful, for providing comments, checking the functionality of our package and also checking the appropriateness of the documentation.

%done
\comment{The article should be proof read to correct grammatical mistakes. page 1, “ensure a good data quality” should read “ensure good data quality”. A few other similar errors were scattered elsewhere}
\response{We proof read the manuscript and removed remaining spelling errors.}

%done
\comment{On page 2 the authors note that the query methods implement a time-delay. Is this user configurable? It doesn’t appear to be and this might be a parameter to consider providing to the user (acknowledging the fact that a user could mis-use it!)}
\response{The time delay is currently hard-coded and adjusted to the capabilities of the data sources.
Some providers block IP addresses if to many request arrive.
Therefore, we do not intend to make this user configurable, as misuse might affect whole institutions.
However, we might adapt the delays if data sources increase their capabilities. 
}

%done
\comment{Some of the man pages I looked at indicated that the query might fail if the API is unavailable. It might be useful to have a helper method to test the availability of the remote resource before the query is run.}
\response{CRAN runs automated tests and all examples in the documentation on different operating systems. Sometimes theses might fail, because an API is not available. From a related project (taxize package to handle taxonomic data in R) we made the experience that APIs are often down only for a short period (updates, maintenance etc). To avoid triggering a failed build on CRAN, we skip running the examples on CRAN and provide the reason why we omitted the examples.
We are thankful for the suggesting and will add ping\_*() functions in a future release to check if a data-source is up and running. 
}

%% --------------------------------
\section{Reviewer B}

\comment{The paper strikes this reviewer as a little "light" compared to most papers
in JSS. This is not to say that the information presented is not useful, but
that it is not very elaborate. I recommend that the authors consider either
extending one or more of the examples, or adding an additional example or two.
}
\response{We added an example how to query legal information in compounds, how to reuse already queried identifiers to query more information and how to join the retrieved information with the original data.}

%done
\comment{Please clarify what the differences are between ChemmineR and this package.
Is the main unique strength of this package that it allows use of more data
sources? If so, please make this more clear. }
\response{The main difference to all mentioned packages is that webchem integrates access to many data sources. We rephrased this paragraph and it now reads: \emph{"Within the R ecosystem, there are only few similar projects: rpubchem (Guha 2014) provides an interface PubChem. 
Similarly, ChemmineR (Cao et al. 2008), a mature chemoinformatics package, also provides an interface to Pubchem. 
webchem does not provide any chemoin-formatic functionality, but integrates access to many data sources.[...]"}
}

%done
\comment{However, data quality is also crucial for data analysis (Stieger et
al. 2014). Ensuring good quality requires additional effort and
methods to be developed.  - Yes, garbage in, garbage out. But just post-processing cannot make bad data
good. I would rather you said "validating the quality of
data" or something similar, and rephrased this paragraph. }
\response{We agree and rephrased this paragraph and merge it with the previous paragraph. It now reads \emph{"However, good quality of data is crucial for every analysis (Stieger et al. 2014) and additional effort and methods are needed to validate data quality."}}

%done
\comment{Recommend to put the abstract into a single paragraph, (same text is fine).}
\response{We agree and changed accordingly.}

%done
\comment{R is one of the most widely used software for data cleaning,-
R is one of the most widely used software enviroments for data cleaning,}
\response{We agree and changed accordingly.}

%done
\comment{For all URLS (to CRAN, Github, etc): Do not insert the URL in the text.
Instead, create a reference in the bibliography, and add a citation in the
text.}
\response{We replaced all URLs in the text with citations.}

%done
\comment{Add references for Travis-CI and AppVeyor.}
\response{We added references to both.}

%done
\comment{’best’ match - `best’ match (fix the quote here and in other places)}
\response{We are thankful for catching this typographical error and changed through the manuscript.}

%done
\comment{In sentence 1 of Section 5.1: change "both" to "and both"}
\response{We changed this paragraph. See also comment number 5.}

%done
\comment{Go through all the references and ensure that titles are in title
style, e.g., “ChemmineR: a compound mining framework for R.” --
“ChemmineR: A Compound Mining Framework for R.”}
\response{We checked all references and changed all titles to title style.}

%% --------------------------------
% \newpage
% \bibliography{refs}
% \bibliographystyle{spbasic}

% ======================================================= %
\end{document}
% ======================================================= %

### 2016-09-20 build_article.R

@@ -1,4 +1,4 @@
 require(knitr)
 require(tools)
-knit('revision1/article.Rnw')
-texi2pdf('revision1/article.tex')
+knit('article.Rnw')
+texi2pdf('article.tex')

### 2016-09-20 revision2/article.R

@@ -1,0 +1,147 @@
+# -------------------------------------------------------------------------
+# Supplemental R Code to reproduce the resutls of
+# Szöcs, Schäfer (2016). webchem: an R Package to Retrieve Chemical Information from the Web
+# Date: 24.04.2016
+# ------------------------------------------------------------------------
+
+
+# Install webchem from CRAN -----------------------------------------------
+install.packages("webchem")
+
+
+# load packages -----------------------------------------------------------
+library("webchem")
+library("ggplot2")
+
+
+# load datasets -----------------------------------------------------------
+# load jagst dataset
+data("jagst")
+# print first 6 lines of data
+head(jagst)
+
+# lc50 dataset
+data("lc50")
+head(lc50)
+
+
+# Use Case 1: Query identifiers -------------------------------------------
+# unique substance names
+subs <- unique(jagst$substance)
+
+# search ETOX IDs, keeping only the best match
+ids <- get_etoxid(subs, match = 'best')
+head(ids)
+
+# use this id to query information from ETOX
+etox_data <- etox_basic(ids$etoxid)
+# extract CAS numbers from retrieved data
+etox_cas <- cas(etox_data)
+head(etox_cas)
+
+
+# query other identifiers from other resources
+
+# query SMILES from PubChem
+cids <- get_cid(etox_cas)
+pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+pc_smiles <- smiles(pc_data)
+
+# ChemSpider needs a security token
+# this is a webchem specific token. 
+# Please use it only for reproduction / testing.
+token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
+# query InChiKey from ChemSpider
+csids <- get_csid(etox_cas, token = token)
+cs_data <- cs_compinfo(csids, token = token)
+cs_inchikey <- inchikey(cs_data)
+
+# combine into single data.frame
+res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+                  cid = pc_data$CID, inchikey = cs_inchikey, 
+                  csid = cs_data$csid, 
+                  stringsAsFactors = FALSE)
+head(res)
+
+
+
+# Use Case II: Toxicity of different pesticide groups ---------------------
[…]

### 2016-09-20 revision2/article.tex
fixed problems

@@
 % latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Tue Sep 20 16:11:04 2016
+% Tue Sep 20 16:29:57 2016
 \begin{tabular}{llllll}
@@
   \midrule
-2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & - & - \\ 
-  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & - & - \\ 
+2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
+  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
   4-para-nonylphenol & - & - & - & - & - \\ 
-  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & - & - \\ 
-  Benzol & 71-43-2 & C1=CC=C... & 241 & - & - \\ 
-  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & - & - \\ 
+  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
+  Benzol & 71-43-2 & C1=CC=C... & 241 & UHOVQNZ... & 236 \\ 
+  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
    \bottomrule
@@
 \end{CodeInput}
-\begin{CodeOutput}
-Warning: closing unused connection 35 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=67-66-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 34 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=75-25-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 33 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=108-88-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 32 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=127-18-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 31 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=5915-41-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 30 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=886-50-0&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 29 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=122-34-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 28 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=60207-90-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 27 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=139-40-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 26 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=23103-98-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 25 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1763-23-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 24 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=335-67-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 23 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=307-24-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 22 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=375-22-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 21 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=40487-42-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 20 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=66246-88-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 19 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=51218-45-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 18 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=67129-08-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 17 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=57837-19-1&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 16 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=94-74-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 15 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=34123-59-6&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 14 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=28159-98-0&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 13 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=83164-33-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 12 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=134-62-3&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 11 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=75-09-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 10 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=30125-63-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 9 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=6190-65-4&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 8 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=71-43-2&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 7 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1912-24-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 6 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=1570-64-5&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
-\begin{CodeOutput}
-Warning: closing unused connection 5 (http://www.chemspider.com/Search.asmx/SimpleSearch?query=105-67-9&token=37bf5e57-9091-42f5-9274-650a64398aaf)
-\end{CodeOutput}
 \end{CodeChunk}

### 2016-09-21 revision2/response/reponse.tex
submitted new rev

@@
 \noindent
-We thank the reviewer for checking the functionality of our package and also checking the appropriateness of the documentation and providing valuable comments on our manuscript.
-We incorporated all the suggested changes and proof read the manuscript.
+We thank the reviewers for checking the functionality of our package and also checking the appropriateness of the documentation and providing valuable comments on our manuscript.
+We incorporated all the suggested changes, proof read the manuscript and adapted to comply with JSS style.
 
@@
 Please find below detailed description of the changes made and responses to specific comments. 
+Additionally, we provide a file with highlighted changes (\texttt{diff.pdf}).
+Changes to code snippets and bibliography were not highlighted.
 
@@
 A few other similar errors were scattered elsewhere}
-\response{We proof read the manuscript and removed remaining spelling errors.}
+\response{We proof read the manuscript and removed remaining spelling and grammatical errors.}
 
@@
 \response{The time delay is currently hard-coded and adjusted to the capabilities of the data sources.
-Some providers block IP addresses if to many request arrive.
+Some providers block IP addresses if too many requests arrive.
 Therefore, we do not intend to make this user configurable, as misuse might affect whole institutions.
@@
 \response{CRAN runs automated tests and all examples in the documentation on different operating systems.
-Sometimes theses might fail, because an API is not available.
+Sometimes these might fail, because an API is not available.
 From a related project (taxize package to handle taxonomic data in R) we made the experience that APIs are often down only for a short period (updates, maintenance etc).
@@
 }
-\response{We added an example how to query legal information in compounds, how to reuse already queried identifiers to query more information and how to join the retrieved information with the original data.}
+\response{We added an example how to query regulatory information in compounds, how to reuse already queried identifiers to query more information and how to join the retrieved information with the original data.}
 
@@
 \response{The main difference to all mentioned packages is that webchem integrates access to many data sources.
-We rephrased this paragraph and it now reads: \emph{"Within the R ecosystem, there are only few similar projects: rpubchem (Guha 2014) provides an interface PubChem.
-Similarly, ChemmineR (Cao et al. 2008), a mature chemoinformatics package, also provides an interface to Pubchem. 
-webchem does not provide any chemoin-formatic functionality, but integrates access to many data sources.[...]"}
+We rephrased this paragraph and it now reads: \emph{"Within the R ecosystem, there are only a few similar projects: rpubchem (Guha 2014) provides an interface to PubChem.
+Similarly, ChemmineR (Cao et al. 2008), a mature chemo-informatics package, provides an interface to Pubchem. 
+webchem does not provide any chemo-informatic functionality, but integrates access to many data sources.[...]"}
 }
@@
 data" or something similar, and rephrased this paragraph. }
-\response{We agree and rephrased this paragraph and merge it with the previous paragraph.
+\response{We agree and rephrased this paragraph and merged it with the previous paragraph.
 It now reads \emph{"However, good quality of data is crucial for every analysis (Stieger et al.

### 2016-09-21 revision2/article.R

@@ -1,7 +1,8 @@
 # -------------------------------------------------------------------------
-# Supplemental R Code to reproduce the resutls of
-# Szöcs, Schäfer (2016). webchem: an R Package to Retrieve Chemical Information from the Web
-# Date: 24.04.2016
+# Supplemental R Code to reproduce the results of
+# Szöcs, Schäfer (2016). webchem: an R Package to Retrieve Chemical Information 
+#     from the Web
+# Date: 21. September 2016
 # ------------------------------------------------------------------------
 
 
@@ -50,7 +51,7 @@ pc_smiles <- smiles(pc_data)
 # ChemSpider needs a security token
 # this is a webchem specific token. 
 # Please use it only for reproduction / testing.
-token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
+token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
 # query InChiKey from ChemSpider
 csids <- get_csid(etox_cas, token = token)
 cs_data <- cs_compinfo(csids, token = token)
@@ -82,7 +83,8 @@ lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
                   ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
                          'other'))))
 lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", 
-                                          "Organo-\nphosphates", "Neo-\nnicotinoids", 
+                                          "Organo-\nphosphates", 
+                                          "Neo-\nnicotinoids", 
                                           "other"))
 
 # plot
@@ -130,8 +132,31 @@ p <- ggplot(lc50, aes(x = logp, y = value)) +
   geom_smooth(method = 'lm') +
   labs(x = 'P', y = expression(LC[50])) +
   theme_bw() +
-  scale_x_continuous(breaks = c(0, 2, 4, 6), labels = format(10^c(0, 2, 4, 6), scientific = FALSE))
+  scale_x_continuous(breaks = c(0, 2, 4, 6), labels = format(10^c(0, 2, 4, 6), 
+                                                             scientific = FALSE))
 p
+
+# Use Case IV: Regulatory information -------------------------------------
+# search EQS from ETOX using the already queried ETOX_IDs
+eqs <- etox_targets(ids$etoxid)
+# extract only MAC-EQS for the EU from the results
[…]

### 2019-10-06 revision3/article.Rnw
copied revision2 to revision3

--- 2016-09-20 revision2/article.Rnw
@@
 Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
-In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on multiple data sources that need to be joined.
+In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on the joining of multiple data sources.
 
@@
 
-\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via \citet{travis-ci} and \citet{appveyor}; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
+\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently approximately 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via \citet{travis-ci} and \citet{appveyor}; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
 
@@
 Web APIs define functions that allow accessing services and data via http and return data in a specific way.
-\pkg{webchem} uses the API of a data source provider, where this is provided.
+\pkg{webchem} uses the API of a data source provider, where available.
 For sources where an API is lacking, data is directly searched and extracted from the web pages, analogous to manual interaction with a website.
@@
 Between each request there is a time-out of 0.3 to 2 seconds depending on the data source. 
-Therefore, processing of larger data sets can take some time, but still represents a major improvement compared to manual look up.
+Therefore, processing of larger data sets can take some time, but still represents a major improvement compared to manual lookup.
 We provide a link to the \emph{Terms of Use} of data providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
@@
 Currently, data can be retrieved from 11 sources.
-These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and legal information (Figure~\ref{fig:fig1}, a detailed overview of all sources is included as supplement):
+These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and regulatory information (Figure~\ref{fig:fig1}, a detailed overview of all sources is included as supplement):
 
@@
   We use an interface to the PUG-REST web service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
-  \item[Wikidata \citep{wiki}]{Wikipedia contains information entries for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}.
-Currently \pkg{webchem} can query only chemical identifiers only.}
+  \item[Wikidata \citep{wiki}]{Wikipedia contains information for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently \pkg{webchem} can only query chemical identifiers.}
   \item[Compendium of Pesticide Common Names \citep{wood}]{The compendium provides information on pesticide common names, identifiers and classification.}
@@
 
-Though the data sources exhibit some overlap in the provided information, each has been selected because it provides unique information and we encourage the interested reader to consult the related source for details.
+Though the data sources exhibit some overlap in the provided information, each has been selected because it also provides unique information and we encourage the interested reader to consult the related source for details.
 However, we provide a brief overview in the Supporting Information.
@@
 To demonstrate the capabilities of  \pkg{webchem} we use two small publicly available real world data sets.
-The data sets are only for demonstration purposes, have been slightly preprocessed (not shown) and are available through the package.
+The data sets are only used for purpose of demonstration, have been slightly preprocessed (not shown) and are available through the package.
 
@@
 It comprises concentrations  (in $\mathrm{\mu g~/~L}$) of  34 substances  on 13 sampling occasions.
-First we load the data set and take a look at the first six rows:
+First we load the data set and inspect the first six rows:
 
@@
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
+In the same manner, we can now query other identifiers from another source using these CAS numbers (Figure~\ref{fig:fig1}), like PubChem
 <<token, echo=FALSE>>=
@@
 
-Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider web page. 
-We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other data sets, e.g.\ the \code{lc50} data set based on CAS.
+Note that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider web page. 
+Finally, we obtain a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other data sets, e.g.\ the \code{lc50} data set based on CAS.
 
@@
 Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
-Answering this question would require tedious look up of insecticide groups for each of the 124 CAS numbers in the \code{lc50} data set.
+Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the \code{lc50} data set.
 The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried using CAS numbers with \pkg{webchem}:
@@
 
-To extract the chemical group from the retrieved data set, we need to write a simple extractor function and apply this to the retrieved data:
+To extract the chemical group from the retrieved data set, we write a simple extractor function and apply this to the retrieved data:
 
@@
 
-Figure \ref{fig:fig2} displays the result after some more data cleaning (see supplement for full code).
+Figure \ref{fig:fig2} displays the result after additional data cleaning (see supplement for full code).
 Overall, it took only 5 \proglang{R} statements to retrieve, clean and plot the data using \pkg{ggplot2} \citep{ggplot2}.
@@
 \subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
-Some data sources contain also data on chemical properties that can be queried.
+Some data sources also provide data on chemical properties that can be queried.
 Here we query for the \code{lc50} data the $\mathrm{log}~P_{oct/wat}$ from the SRC PHYSPROP database to build a simple quantitative structure–activity relationship (QSAR) to predict toxicity.
@@
 We opted for this more complex approach, because the information available is very diverse and we cannot provide an extractor function for each purpose.
-Moreover, it gives users full control over how to organize their data. 
-Nevertheless, we provide in the documentation of each function examples on how to extract more complicated parts of the data.
+Moreover, it provides users with high flexibility regarding organisation of their data. 
+Nevertheless, in the documentation of each function we provide examples on how to extract more complicated parts of the data.
 The resulting data and model is displayed in Figure~\ref{fig:fig3}.
@@
 
-\subsection[Legal information]{Legal information}
-Legal information is particularly of interest if concentrations exceed national thresholds.
-In the European Union (EU) the Water Framework Directive (WFD) defines Environmental Quality Standards (EQS).
-Similarly, in the U.S. and Canadian EPA and the WHO defines Quality Standards.
+\subsection[Regulatory information]{Regulatory information}
+Regulatory information is particularly of interest if concentrations exceed national thresholds.
+In the European Union (EU) the Water Framework Directive (WFD, \citet{wfd2000directive}) defines Environmental Quality Standards (EQS).
+Similarly, the U.S. and Canadian EPA and the WHO define Quality Standards.
 Information on these standards can be queried with \pkg{webchem} from the PAN Pesticide Database (using \code{pan\_query()}) and from ETOX (using \code{etox\_targets()}).
 
-In this example we search for the minimum EQS for the EU for the compounds in the \code{jagst} data set, join these with measureds concentrations and check if there have been any exceedances.
+In this example we search for the minimum EQS for the EU for the compounds in the \code{jagst} data set, join these with measured concentrations and evaluate wether exceedances occurred..
 
-We re-use the above queried ETOX-IDs to get further information from ETOX, namely the MAC-EQS:
+We re-use the above queried ETOX-IDs to obtain further information from ETOX, namely the MAC-EQS:
 <<search_targests, message=FALSE, warning=FALSE>>=
@@
 
-Again, the returned information is humongous and we encourage the users to study the returned objects and description of the data source.
-Here the column \code{Designation} defines the type of EQS and \code{Value_Target_LR} contains the value.
-Unfortunately, we found MAC-EQS values only for 5 substances:
+Again, the returned information is humongous and we encourage users to study the returned objects and description of the data source.
+Here, the column \code{Designation} defines the type of EQS and \code{Value_Target_LR} contains the value.
+Unfortunately, we only found MAC-EQS values for 5 substances:
 
@@
 
-Finally, we can compare the measured value with the MAC, which reveals that there have been no exceedances of these 5 compounds.
+Finally, we can compare the measured value to the MAC, which reveals that there have been no exceedances of these 5 compounds.
 
@@
 \subsection[Related software]{Related software}
-Within the \proglang{R} ecosystem, there are only few similar projects:
-\pkg{rpubchem} \citep{rpubchem_2014} provides an interface PubChem.
-Similarly, \pkg{ChemmineR} \citep{chemminer_2008}, a mature chemo-informatics package, also provides an interface to Pubchem. 
+Within the \proglang{R} ecosystem, there are only a few similar projects:
+\pkg{rpubchem} \citep{rpubchem_2014} provides an interface to PubChem.
+Similarly, \pkg{ChemmineR} \citep{chemminer_2008}, a mature chemo-informatics package, provides an interface to Pubchem. 
 \pkg{webchem} does not provide any chemo-informatic functionality, but integrates access to many data sources.
@@
 The Chemical Translation Service \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also one of the sources that can be queried, allows batch conversion of chemical identifiers.
-However, it does not provide access to other data (experimental, modeled or legal data).
+However, it does not provide access to other data (experimental, modeled or regulatory data).
 

### 2019-10-06 revision3/article.tex
copied revision2 to revision3

--- 2016-09-20 revision2/article.tex
@@
 
-In the same manner we can now query other identifiers from other source using this CAS numbers (Figure~\ref{fig:fig1}), like PubChem
+In the same manner, we can now query other identifiers from another source using these CAS numbers (Figure~\ref{fig:fig1}), like PubChem
 \begin{CodeChunk}
@@
 
-\subsection[Legal information]{Legal information}
-Legal information is particularly of interest if concentrations exceed national thresholds.
-In the European Union (EU) the Water Framework Directive (WFD) defines Environmental Quality Standards (EQS).
-Similarly, in the U.S. and Canadian EPA and the WHO defines Quality Standards.
+\subsection[Regulatory information]{Regulatory information}
+Regulatory information is particularly of interest if concentrations exceed national thresholds.
+In the European Union (EU) the Water Framework Directive (WFD, \citet{wfd2000directive}) defines Environmental Quality Standards (EQS).
+Similarly, the U.S. and Canadian EPA and the WHO define Quality Standards.
 Information on these standards can be queried with \pkg{webchem} from the PAN Pesticide Database (using \code{pan\_query()}) and from ETOX (using \code{etox\_targets()}).
 
-In this example we search for the minimum EQS for the EU for the compounds in the \code{jagst} data set, join these with measureds concentrations and check if there have been any exceedances.
+In this example we search for the minimum EQS for the EU for the compounds in the \code{jagst} data set, join these with measured concentrations and evaluate wether exceedances occurred..
 
-We re-use the above queried ETOX-IDs to get further information from ETOX, namely the MAC-EQS:
+We re-use the above queried ETOX-IDs to obtain further information from ETOX, namely the MAC-EQS:
 \begin{CodeChunk}

### 2019-10-06 revision3/article.R

@@ -1,0 +1,172 @@
+# -------------------------------------------------------------------------
+# Supplemental R Code to reproduce the results of
+# Szöcs, Schäfer (2016). webchem: an R Package to Retrieve Chemical Information 
+#     from the Web
+# Date: 21. September 2016
+# ------------------------------------------------------------------------
+
+
+# Install webchem from CRAN -----------------------------------------------
+install.packages("webchem")
+
+
+# load packages -----------------------------------------------------------
+library("webchem")
+library("ggplot2")
+
+
+# load datasets -----------------------------------------------------------
+# load jagst dataset
+data("jagst")
+# print first 6 lines of data
+head(jagst)
+
+# lc50 dataset
+data("lc50")
+head(lc50)
+
+
+# Use Case 1: Query identifiers -------------------------------------------
+# unique substance names
+subs <- unique(jagst$substance)
+
+# search ETOX IDs, keeping only the best match
+ids <- get_etoxid(subs, match = 'best')
+head(ids)
+
+# use this id to query information from ETOX
+etox_data <- etox_basic(ids$etoxid)
+# extract CAS numbers from retrieved data
+etox_cas <- cas(etox_data)
+head(etox_cas)
+
+
+# query other identifiers from other resources
+
+# query SMILES from PubChem
+cids <- get_cid(etox_cas)
+pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+pc_smiles <- smiles(pc_data)
+
+# ChemSpider needs a security token
+# this is a webchem specific token. 
+# Please use it only for reproduction / testing.
+token <- '37bf5e57-9091-42f5-9274-650a64398aaf'
+# query InChiKey from ChemSpider
+csids <- get_csid(etox_cas, token = token)
+cs_data <- cs_compinfo(csids, token = token)
+cs_inchikey <- inchikey(cs_data)
+
+# combine into single data.frame
+res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+                  cid = pc_data$CID, inchikey = cs_inchikey, 
+                  csid = cs_data$csid, 
+                  stringsAsFactors = FALSE)
+head(res)
+
+
+
+# Use Case II: Toxicity of different pesticide groups ---------------------
[…]

### 2019-10-06 revision3_2581/Code/code.R
version received from editors,  must be used for edits

@@ -1,0 +1,171 @@
+# -------------------------------------------------------------------------
+# Supplemental R Code to reproduce the results of
+# Szöcs, Schäfer. webchem: An R Package to Retrieve Chemical Information 
+#   from the Web
+# ------------------------------------------------------------------------
+
+
+# Install webchem from CRAN -----------------------------------------------
+install.packages("webchem")
+
+
+# load packages -----------------------------------------------------------
+library("webchem")
+library("ggplot2")
+
+
+# load datasets -----------------------------------------------------------
+# load jagst dataset
+data("jagst", package = "webchem")
+# print first 6 lines of data
+head(jagst)
+
+# lc50 dataset
+data("lc50", package = "webchem")
+head(lc50)
+
+
+# Use Case 1: Query identifiers -------------------------------------------
+# unique substance names
+subs <- unique(jagst$substance)
+
+# search ETOX IDs, keeping only the best match
+ids <- get_etoxid(subs, match = "best")
+head(ids)
+
+# use this id to query information from ETOX
+etox_data <- etox_basic(ids$etoxid)
+# extract CAS numbers from retrieved data
+etox_cas <- cas(etox_data)
+head(etox_cas)
+
+
+# query other identifiers from other resources
+
+# query SMILES from PubChem
+cids <- get_cid(etox_cas)
+pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
+pc_smiles <- smiles(pc_data)
+
+# ChemSpider needs a security token
+# this is a webchem specific token. 
+# Please use it only for reproduction / testing.
+token <- "37bf5e57-9091-42f5-9274-650a64398aaf"
+# query InChiKey from ChemSpider
+csids <- get_csid(etox_cas, token = token)
+cs_data <- cs_compinfo(csids, token = token)
+cs_inchikey <- inchikey(cs_data)
+
+# combine into single data.frame
+res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+                  cid = pc_data$CID, inchikey = cs_inchikey, 
+                  csid = cs_data$csid, 
+                  stringsAsFactors = FALSE)
+head(res)
+
+
+
+# Use Case II: Toxicity of different pesticide groups ---------------------
[…]

### 2019-10-07 revision3_2581/paper.tex

--- 2019-10-06 revision3/article.tex
@@
-\documentclass[article, shortnames]{jss}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\usepackage{Sweave}
-
-\usepackage[utf8]{inputenc} %unicode support
-\usepackage{booktabs}
-\usepackage{fixltx2e}
+\documentclass[article]{jss}
+\usepackage{thumbpdf,lmodern} 
+\graphicspath{{Figures/}}
 
@@
 \Plainauthor{Eduard Sz\"ocs, Ralf B. Sch\"afer}
-\Plaintitle{webchem: an R Package to Retrieve Chemical Information from the Web}
+
+\Plaintitle{webchem: An R Package to Retrieve Chemical Information from the Web}
+
 \Shorttitle{\pkg{webchem}: Chemical Information from the Web}
 
-\Abstract{
-A wide range of chemical information is freely available online, including identifiers, experimental and predicted chemical properties.
-However, these data are scattered over various data sources and not easily accessible to researchers.
-Manual searching and downloading of such data is time-consuming and error-prone.  
-We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query chemical data from currently 11 web sources. 
-These cover a broad spectrum of information.
-The data are automatically imported into an \proglang{R} object and can directly be used in subsequent analyses.
-\pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web sources.
-In addition, it facilitates data cleaning, identification and reporting of substances.
-Consequently, it reduces the time researchers need to spend on chemical data compilation.
-}
+\Abstract{ A wide range of chemical information is freely available
+  online, including identifiers, experimental and predicted chemical
+  properties.  However, these data are scattered over various data
+  sources and not easily accessible to researchers.  Manual searching
+  and downloading of such data is time-consuming and error-prone.  We
+  developed the open-source \proglang{R} package \pkg{webchem} that
+  allows users to automatically query chemical data from currently 11
+  web sources.  These cover a broad spectrum of information.  The data
+  are automatically imported into an \proglang{R} object and can
+  directly be used in subsequent analyses.  \pkg{webchem} enables
+  easy, structured and reproducible data retrieval and usage from
+  publicly available web sources.  In addition, it facilitates data
+  cleaning, identification and reporting of substances.  Consequently,
+  it reduces the time researchers need to spend on chemical data
+  compilation.  }
+
 \Keywords{ecotoxicology, chemistry, data cleaning, web scraping, ropensci}
+
 \Plainkeywords{ecotoxicology, chemistry, data cleaning, web scraping, ropensci}
 
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
 %% \Volume{50}
@@
 
-%% The address of (at least) one author should be given
-%% in the following format:
 \Address{
@@
   Universit\"at Koblenz-Landau\\
-  Fortstraße 7\\
+  Fortstra{\ss}e 7\\
   76829 Landau, Germany\\
@@
 
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
 % citation alias for LUBW institution
 \defcitealias{lubw_2016}{LUBW (2016)}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \section[Introduction]{Introduction}
-Before each statistical analysis, data cleaning is often required to ensure good data quality.
-Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
-In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on the joining of multiple data sources.
+Before each statistical analysis, data cleaning is often required to
+ensure good data quality.  Data cleaning is the process of detecting
+errors and inconsistencies in data sets \citep{Chapman_2005}.  In
+practice, the data cleaning step is often more time consuming than the
+subsequent statistical analysis, particularly, when the analysis
+relies on the joining of multiple data sources.
 
-When dealing with chemical data sets (e.g.\ environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
-Several web sources provide chemical names or link them to unique codes (see also section \emph{Data sources} below).
-However, manual searching for each compound, often through a graphical web interface, is tedious, error-prone and not reproducible \citep{Peng_2009}.
+When dealing with chemical data sets (e.g., environmental monitoring
+data, toxicological data), a first step is often to validate the names
+of chemicals or to link them to unique codes that simplify subsequent
+querying and appending of compound-related physico-chemical or
+toxicological information.  Several web sources provide chemical names
+or link them to unique codes (see also
+Section~\ref{sec:data-sources}).  However, manual searching for each
+compound, often through a graphical web interface, is tedious,
+error-prone and not reproducible \citep{Peng_2009}.
 
-To simplify, robustify and automate this task, i.e.\ to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
-\proglang{R} is one of the most widely used software environments for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.
+To simplify, robustify and automate this task, i.e., to search and
+retrieve chemical information from the web, we created the
+\pkg{webchem} package \citep{cran} for the free and open source
+\proglang{R} language \citep{r_2015, Wehrens_2011}.  \proglang{R} is
+one of the most widely used software environments for data cleaning,
+analyzing and visualizing data, and supports full reproducibility of
+each step \citep{Marwick_2016}.
 
-In the following, we describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.
+In the following, we describe the basic functionality of the package
+and demonstrate with a few use cases how to clean and retrieve new
+data with \pkg{webchem}.
 
 \section[Implementation and design details]{Implementation and design details}
-The \pkg{webchem} package is written entirely in \proglang{R} and available under a MIT license.
-The development repository is hosted on \citet{github} and a stable version is released on the official \proglang{R} repository \citep{cran}.
-\pkg{webchem} is part of the rOpenSci project \citep{boettiger2015building}, which aims at fully reproducible data analysis.
+The \pkg{webchem} package is written entirely in \proglang{R} and
+available under an MIT license.  The development repository is hosted
+on \citet{github} and a stable version is released on the
+Comprehensive \proglang{R} Archive Network (CRAN) and available at
+\url{https://CRAN.R-project.org/package=webchem}.  \pkg{webchem} is
+part of the rOpenSci project \citep{boettiger2015building}, which aims
+at fully reproducible data analysis.
 
-\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently approximately 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via \citet{travis-ci} and \citet{appveyor}; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
+\pkg{webchem} follows best practices for scientific software
+\citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public
+available repository with easy collaboration and an issue tracker (via
+GitHub), (ii) a non-restrictive license, version control (git), (iii)
+an elaborate test-suite covering more than 90\% of the relevant lines
+of code (currently approximately 1500 lines, using \pkg{testthat};
+\citealt{wickham_testthat:_2011}), (iv) continuous integration (via
+\citet{travis-ci} and \citet{appveyor}; testing on Linux \& Windows
+with current and development \proglang{R} versions), (v) in-source
+documentation (using \pkg{roxygen2}; \citealt{wickham_roxygen2:_2015})
+and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
 
 \pkg{webchem} builds on top of the following \proglang{R} packages:
-\pkg{RCurl} \citep{lang_rcurl:_2015} and \pkg{httr} \citep{wickham_httr} for data transfer,
-\pkg{stringr} \citep{wickham_stringr:_2015} for string handling,
-\pkg{xml2} \citep{wickham_xml2} and \pkg{rvest} \citep{wickham_rvest} for parsing HTML and XML,
-\pkg{jsonlite} \citep{ooms_jsonlite_2014} for parsing JSON,
-\pkg{rcdk} \citep{guha_rcdk} for parsing SMILES.
-For parsing molfiles we use a lightweight implementation of \citep{Grabner_Varmuza_Dehmer_2012}.
+\pkg{RCurl} \citep{lang_rcurl:_2015} and \pkg{httr}
+\citep{wickham_httr} for data transfer, \pkg{stringr}
+\citep{wickham_stringr:_2015} for string handling, \pkg{xml2}
+\citep{wickham_xml2} and \pkg{rvest} \citep{wickham_rvest} for parsing
+HTML and XML, \pkg{jsonlite} \citep{ooms_jsonlite_2014} for parsing
+JSON, \pkg{rcdk} \citep{guha_rcdk} for parsing SMILES.  For parsing
+molfiles we use a lightweight implementation in package \pkg{RMol}
+\citep{Grabner_Varmuza_Dehmer_2012}.
 
 Some data sources provide application programming interfaces (API).
-Web APIs define functions that allow accessing services and data via http and return data in a specific way.
-\pkg{webchem} uses the API of a data source provider, where available.
-For sources where an API is lacking, data is directly searched and extracted from the web pages, analogous to manual interaction with a website.
+Web APIs define functions that allow accessing services and data via
+http and return data in a specific way.  \pkg{webchem} uses the API of
+a data source provider, where available.  For sources where an API is
+lacking, data is directly searched and extracted from the web pages,
+analogous to manual interaction with a website.
 
-Only few design decisions have been made:
-Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
-They follow the format of \code{source\_function}, e.g.\ \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
-Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
-The prefix \code{get} is used to denote these functions, e.g.\ \code{get\_csid} to retrieve the identifier used in ChemSpider.
+Only few design decisions have been made: Each function name has a
+prefix and suffix separated by an underscore
+\citep{Chamberlain_Szocs_2013}.  They follow the format of
+\code{source\_function}, e.g., \code{cs\_compinfo} uses ChemSpider as
+source (see Section~\ref{sec:data-sources}) to retrieve compound
+information.  Some functions require querying first a unique
+identifier from the data source and then use this identifier to query
+further information.  The prefix \code{get} is used to denote these
+functions, e.g., \code{get\_csid} to retrieve the identifier used in
+ChemSpider.
 
-\pkg{webchem} is friendly to the resources of data providers. 
-Between each request there is a time-out of 0.3 to 2 seconds depending on the data source. 
-Therefore, processing of larger data sets can take some time, but still represents a major improvement compared to manual lookup.
-We provide a link to the \emph{Terms of Use} of data providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
-Moreover, all functions return an URL of the source, which can be used for \mbox{(micro-)attribution}.
+\pkg{webchem} is friendly to the resources of data providers.  Between
+each request there is a time-out of 0.3 to 2 seconds depending on the
+data source.  Therefore, processing of larger data sets can take some
+time, but still represents a major improvement compared to manual
+lookup.  We provide a link to the \emph{Terms of Use} of data
+providers in the documentation of each function and we encourage the
+users to read these before using \pkg{webchem}.  Moreover, all
+functions return an URL of the source, which can be used for
+\mbox{(micro-)attribution}.
 
-\section[Data sources]{Data sources}
-The backbone of \pkg{webchem} are data sources providing their data and functionality to the public.
-Currently, data can be retrieved from 11 sources.
-These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and regulatory information (Figure~\ref{fig:fig1}, a detailed overview of all sources is included as supplement):
+\section[Data sources]{Data sources}\label{sec:data-sources}
+The backbone of \pkg{webchem} are data sources providing their data
+and functionality to the public.  Currently, data can be retrieved
+from 11 sources.  These cover a broad spectrum of available data, like
+identifiers, experimental and predicted properties and regulatory
+information (a detailed overview of all sources is included in
+Figure~\ref{fig:fig1}):
 
 \begin{description}
-  \item[NIH Chemical Identifier Resolver (CIR) \citep{cir}]{A web service that converts from and to various chemical identifiers.}
-  \item[Chemical Translation Service (CTS) \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}]{A web service that converts from and to various chemical identifiers.}
-  \item[ETOX \citep{etox}]{Information System Ecotoxicology and Environmental Quality Targets by the German Federal Environmental Agency.
-Provides basic identifiers, synonyms, ecotoxicological data and quality targets for different countries.}
-  \item[PAN Pesticide Database \citep{pan}]{Information on pesticides - provides basic identifiers, ecotoxicological data and chemical properties.}
-  \item[SRC Physprop \citep{physprop}]{Contains physical properties for over 41,000 chemicals.
-  Physical properties collected from a wide variety of sources including experimental and modeled values.}
-  \item[PubChem \citep{Kim_2016}]{PubChem is a public repository for information on chemical substances, providing identifiers, properties and synonyms.
-  We use an interface to the PUG-REST web service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
-  \item[Wikidata \citep{wiki}]{Wikipedia contains information for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently \pkg{webchem} can only query chemical identifiers.}
-  \item[Compendium of Pesticide Common Names \citep{wood}]{The compendium provides information on pesticide common names, identifiers and classification.}
-  \item[ChemID\emph{plus} \citep{Tomasulo_2002}]{is a large web-based database provided by the National Library of Medicine.
-It provides identifiers, synonyms, toxicological data and chemical properties.}
-  \item[ChemSpider \citep{pence_chemspider:_2010}]{is a free chemical structure database providing access to over 40 million structures.
-It provides identifiers, properties and can also be used to convert identifiers.}
-  \item[OPSIN  \citep{Lowe_Corbett_Murray-Rust_Glen_2011}]{The Open Parser for Systematic IUPAC nomenclature is a chemical name interpreter and provides InChI and SMILES identifiers.}
+\item[NIH Chemical Identifier Resolver (CIR)]{\citep{cir} A web
+    service that converts from and to various chemical identifiers.}
+\item[Chemical Translation Service (CTS)]{
+  \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010} A web
+    service that converts from and to various chemical identifiers.}
+\item[ETOX]{\citep{etox} Information System Ecotoxicology and
+    Environmental Quality Targets by the German Federal Environmental
+    Agency. Provides basic identifiers, synonyms, ecotoxicological
+    data and quality targets for different countries.}
+\item[PAN Pesticide Database]{ \citep{pan} Information on pesticides --
+    provides basic identifiers, ecotoxicological data and chemical
+    properties.}
+\item[SRC Physprop]{\citep{physprop} Contains physical properties for
+    over 41,000 chemicals.  Physical properties collected from a wide
+    variety of sources including experimental and modeled values.}
+\item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
+    information on chemical substances, providing identifiers,
+    properties and synonyms.  We use an interface to the PUG-REST web
+    service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
+\item[Wikidata]{\citep{wiki} Wikipedia contains information for over
+    15,000 chemicals
+    \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
+    \pkg{webchem} can only query chemical identifiers.}
+\item[Compendium of Pesticide Common Names]{\citep{wood} The
+    compendium provides information on pesticide common names,
+    identifiers and classification.}
+\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
+    database provided by the National Library of Medicine. It provides
+    identifiers, synonyms, toxicological data and chemical
+    properties.}
+\item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
+    structure database providing access to over 40 million
+    structures. It provides identifiers, properties and can also be
+    used to convert identifiers.}
+\item[OPSIN]{\citep{Lowe_Corbett_Murray-Rust_Glen_2011} The Open
+    Parser for Systematic IUPAC nomenclature is a chemical name
+    interpreter and provides InChI and SMILES identifiers.}
 \end{description}
 
-Though the data sources exhibit some overlap in the provided information, each has been selected because it also provides unique information and we encourage the interested reader to consult the related source for details.
-However, we provide a brief overview in the Supporting Information.
+Though the data sources exhibit some overlap in the provided
+information, each has been selected because it also provides unique
+information and we encourage the interested reader to consult the
+related source for details.  However, we provide a brief overview in
+the Supporting Information.
 
-\begin{figure}[ht]
+\begin{figure}[t!]
   \centering
@@
 \section[Use cases]{Use cases}
-\subsection[Install webchem]{Installation}
-\pkg{webchem} can be easily installed and loaded from CRAN:
-
+\subsection[Installation]{Installation}
+\pkg{webchem} can be easily installed from CRAN and loaded:
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
-\begin{CodeChunk}
-\end{CodeChunk}
-
-The package is under active development. The latest development version is available from GitHub and also permanently available at \citet{zenodo}.
-This document has been created using \pkg{webchem} version 0.1.
+%
+The package is under active development. The latest development
+version is available from GitHub and also permanently available at
+\citet{zenodo}.  This document has been created using \pkg{webchem}
+version 0.4.0.
 
 \subsection[Sample data sets]{Sample data sets}
-To demonstrate the capabilities of  \pkg{webchem} we use two small publicly available real world data sets.
-The data sets are only used for purpose of demonstration, have been slightly preprocessed (not shown) and are available through the package.
+To demonstrate the capabilities of \pkg{webchem} we use two small
+publicly available real world data sets.  The data sets are only used
+for purpose of demonstration, have been slightly preprocessed (not
+shown) and are available through the package.
 
@@
 First we load the data set and inspect the first six rows:
-
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> data("jagst")
+R> data("jagst", package = "webchem")
 R> head(jagst)
@@
 \end{CodeChunk}
-
+%
 This data set identifies substances only by substance names. Values below the limit of quantification (LOQ) are indicated by a qualifier column.
@@
 The data has been retrieved from the EPA ECOTOX database \citep{epa_2016}.
-
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> data("lc50")
+R> data("lc50", package = "webchem")
 R> head(lc50)
@@
 \end{CodeChunk}
-
+%
 This data set identifies the substances only by CAS numbers.
@@
 
-One possibility to resolve this, is to use different chemical identifiers allowing easy identification.
-There are several identifiers available, e.g.\  registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.
-In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to facilitate subsequent matching with other data.
+One possibility to resolve this, is to use different chemical
+identifiers allowing easy identification.  There are several
+identifiers available, e.g., registry numbers like CAS or EC, database
+identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID
+\citep{pence_chemspider:_2010}, line notations like SMILES
+\citep{Weininger_1990}, InChI and InChIKey
+\citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}.  In this
+first example we query several identifiers to create a table that can
+be used as (i) supplemental information to a research article or (ii)
+to facilitate subsequent matching with other data.
 
@@
 First we search for ETOX internal ID numbers using the substance names:
-
+%
 \begin{CodeChunk}
@@
 R> subs <- unique(jagst$substance)
-R> ids <- get_etoxid(subs, match = 'best')
+R> ids <- get_etoxid(subs, match = "best")
 R> head(ids)
@@
 \end{CodeChunk}
-
-Only three substances could not be found in ETOX. 
-Here we specify that only the \emph{`best'} match (in terms of the Levenshtein distance between query and results) is returned. 
-A manual check confirms appropriate matches. 
-Other options include: \emph{`all'} - returns all matches; \emph{`first'} - returns only the first match (not necessarily the best match); \emph{`ask'} - this enters an interactive mode, where the user is asked for a choice if multiple matches are found and \emph{`na'} which returns \code{NA} in case of multiple matches.
+%
+Only three substances could not be found in ETOX.  Here we specify
+that only the \code{"best"} match (in terms of the Levenshtein
+distance between query and results) is returned.  A manual check
+confirms appropriate matches.  Other options include: \code{"all"} --
+returns all matches; \code{"first"} -- returns only the first match
+(not necessarily the best match); \code{"ask"} -- this enters an
+interactive mode, where the user is asked for a choice if multiple
+matches are found and \code{"na"} which returns \code{NA} in case of
+multiple matches.
 
 We use these data to retrieve basic information on the substances.
-
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
+%
 \pkg{webchem} always returns a named list (one entry for each substance) and the available information content can be very voluminous.
 Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
+%
+A variety of data are available and we cannot provide extractor
+functions for each of those.  Therefore, if users need to extract
+other data, they have to write simple extractor functions (see the
+following examples).
 
-A variety of data are available and we cannot provide extractor functions for each of those.
-Therefore, if users need to extract other data, they have to write simple extractor functions (see following examples).
-
-In the same manner, we can now query other identifiers from another source using these CAS numbers (Figure~\ref{fig:fig1}), like PubChem
-\begin{CodeChunk}
-\end{CodeChunk}
-
+In the same manner, we can now query other identifiers from another
+source using these CAS numbers (see Figure~\ref{fig:fig1}), like
+PubChem
+%
 \begin{CodeChunk}
@@
 R> cids <- get_cid(etox_cas)
-R> pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
+R> pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
 R> pc_smiles <- smiles(pc_data)
@@
 \end{CodeChunk}
-
+%
 or ChemSpider
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
-Finally, we combine the queried data into one data.frame
+%
+Finally, we combine the queried data into one data frame
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
+%
+Note that in order to use the ChemSpider functions, a personal
+authentication key (\code{token}) is needed, which can be retrieved
+from the ChemSpider web page.  Finally, we obtain a compound table
+containing many different identifiers (Table~\ref{tab:comptable}),
+allowing easy identification and merging with other data sets, e.g.,
+the \code{lc50} data set based on CAS.
 
-Note that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider web page. 
-Finally, we obtain a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other data sets, e.g.\ the \code{lc50} data set based on CAS.
-
-\begin{table}[ht]
+\begin{table}[t!]
 \centering
-\begin{CodeChunk}
-% latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Tue Sep 20 16:29:57 2016
 \begin{tabular}{llllll}
-  \toprule
+  \hline
 Name & CAS & SMILES & CID & InChIKey & CSID \\ 
-  \midrule
+  \hline
 2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
   4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
-  4-para-nonylphenol & - & - & - & - & - \\ 
+  4-para-nonylphenol & -- & -- & -- & -- & -- \\ 
   Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
@@
   Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
-   \bottomrule
+   \hline
 \end{tabular}
-\end{CodeChunk}
-\caption{Identifiers for the jagst data sets as queried with \pkg{webchem}. Only the first 6 entries are shown. For SMILES and InChIKey only the first 7 characters are shown. - = not found.}
+\caption{Identifiers for the \code{jagst} data sets as queried with
+  \pkg{webchem}. Only the first 6 entries are shown. For SMILES and
+  InChIKey only the first 7 characters are shown. -- = not found.}
 \label{tab:comptable}
@@
 \subsection[Toxicity of different pesticide groups]{Toxicity of different pesticide groups}
-Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
-Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the \code{lc50} data set.
-The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried using CAS numbers with \pkg{webchem}:
-
+Another question we might ask is \emph{How does toxicity vary between
+  insecticide groups?}  Answering this question would require tedious
+lookup of insecticide groups for each of the 124 CAS numbers in the
+\code{lc50} data set.  The Compendium of Pesticide Common Names
+\citep{wood} contains such information and can be easily queried using
+CAS numbers with \pkg{webchem}:
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> aw_data <- aw_query(lc50$cas, type = 'cas')
+R> aw_data <- aw_query(lc50$cas, type = "cas")
 \end{CodeInput}
 \end{CodeChunk}
-
+%
 To extract the chemical group from the retrieved data set, we write a simple extractor function and apply this to the retrieved data:
-
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
+%
+Figure~\ref{fig:fig2} displays the result after additional data
+cleaning (see the supplementary material for the full code).  Overall, it
+took only 5 \proglang{R} statements to retrieve, clean and plot the
+data using \pkg{ggplot2} \citep{ggplot2}.
 
-Figure \ref{fig:fig2} displays the result after additional data cleaning (see supplement for full code).
-Overall, it took only 5 \proglang{R} statements to retrieve, clean and plot the data using \pkg{ggplot2} \citep{ggplot2}.
-
-\begin{figure}[ht]
-\begin{CodeChunk}
-
-{\centering \includegraphics[width=0.7\textwidth]{plot_lc50-1}
-
-}
-
-\end{CodeChunk}
-\caption{Toxicity of different pesticide groups. LC\textsubscript{50} values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=0.7\textwidth, trim = 0 10 0 5, clip]{plot_lc50-1} 
+\caption{Toxicity of different pesticide groups. LC\textsubscript{50}
+  values have been retrieved from the EPA ECOTOX database, chemical groups
+  from the Compendium of Pesticide Common Names \citep{wood}.}
 \label{fig:fig2}
@@
 \subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
-Some data sources also provide data on chemical properties that can be queried.
-Here we query for the \code{lc50} data the $\mathrm{log}~P_{oct/wat}$ from the SRC PHYSPROP database to build a simple quantitative structure–activity relationship (QSAR) to predict toxicity.
-
+Some data sources also provide data on chemical properties that can be
+queried.  Here we query for the \code{lc50} data the
+$\mathrm{log}~P_{oct/wat}$ from the SRC Physprop database to build a
+simple quantitative structure-activity relationship (QSAR) to predict
+toxicity.
+%
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
+%
 The database contains predicted and experimental values.
@@
 because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).
-
+%
 \begin{CodeChunk}
@@
 +      return(NA)
-+    y$prop$value[y$prop$variable == 'Log P (octanol-water)']
++    y$prop$value[y$prop$variable == "Log P (octanol-water)"]
 +  })
@@
 \end{CodeChunk}
+%
+We opted for this more complex approach, because the information available is very diverse and we cannot provide an extractor function for each purpose.
+Moreover, it provides users with high flexibility regarding organization of their data. 
+Nevertheless, in the documentation of each function we provide examples on how to extract more complicated parts of the data.
+The resulting data and model are displayed in Figure~\ref{fig:fig3}.
 
-We opted for this more complex approach, because the information available is very diverse and we cannot provide an extractor function for each purpose.
-Moreover, it provides users with high flexibility regarding organisation of their data. 
-Nevertheless, in the documentation of each function we provide examples on how to extract more complicated parts of the data.
-The resulting data and model is displayed in Figure~\ref{fig:fig3}.
-
-\begin{figure}[ht]
-\begin{CodeChunk}
-
-{\centering \includegraphics[width=0.7\textwidth]{plot_qsar-1}
-
-}
-
-\end{CodeChunk}
-\caption{Simple QSAR for predicting log LC\textsubscript{50} of pesticides by log P. 
-Log P values have been retrieved from SRC Physprop database (97 experimental data, 9 estimated data and 18 substances without data). 
-Blue line indicates the regression model ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log} P$, RMSE = 1.45).}
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{plot_qsar-1} 
+  \caption{Simple QSAR for predicting log LC\textsubscript{50} of
+    pesticides by $\mathrm{log}~P$.  $\mathrm{log}~P$ values have been
+    retrieved from the SRC Physprop database (97 experimental data, 9
+    estimated data and 18 substances without data).  Blue line
+    indicates the regression model
+    ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log}~P$,
+    RMSE = 1.45).}
 \label{fig:fig3}
@@
 \subsection[Regulatory information]{Regulatory information}
-Regulatory information is particularly of interest if concentrations exceed national thresholds.
-In the European Union (EU) the Water Framework Directive (WFD, \citet{wfd2000directive}) defines Environmental Quality Standards (EQS).
-Similarly, the U.S. and Canadian EPA and the WHO define Quality Standards.
-Information on these standards can be queried with \pkg{webchem} from the PAN Pesticide Database (using \code{pan\_query()}) and from ETOX (using \code{etox\_targets()}).
+Regulatory information is particularly of interest if concentrations
+exceed national thresholds.  In the European Union (EU) the Water
+Framework Directive (WFD; \citealt{wfd2000directive}) defines
+Environmental Quality Standards (EQS).  Similarly, the U.S.\ and
+Canadian EPA and the WHO define Quality Standards.  Information on
+these standards can be queried with \pkg{webchem} from the PAN
+Pesticide Database (using \code{pan\_query()}) and from ETOX (using
+\code{etox\_targets()}).
 
-In this example we search for the minimum EQS for the EU for the compounds in the \code{jagst} data set, join these with measured concentrations and evaluate wether exceedances occurred..
+In this example we search for the minimum EQS for the EU for the
+compounds in the \code{jagst} data set, join these with measured
+concentrations and evaluate whether exceedances occurred.
 
-We re-use the above queried ETOX-IDs to obtain further information from ETOX, namely the MAC-EQS:
+We re-use the above queried ETOX-IDs to obtain further information
+from ETOX, namely the MAC-EQS:
+% 
 \begin{CodeChunk}
@@
 R> eqs <- etox_targets(ids$etoxid)
-R> ids$mac <- sapply(eqs, function(y){
+R> ids$mac <- sapply(eqs, function(y) {
 +    if (length(y) == 1 && is.na(y)) {
@@
 +      res <- y$res
-+      min(res[res$Country_or_Region == 'EEC / EU' & 
-+                res$Designation == 'MAC-EQS', 'Value_Target_LR'])
++      min(res[res$Country_or_Region == "EEC / EU" & 
++        res$Designation == "MAC-EQS", "Value_Target_LR"])
 +    }
@@
 \end{CodeChunk}
-
+%
 Again, the returned information is humongous and we encourage users to study the returned objects and description of the data source.
@@
 Unfortunately, we only found MAC-EQS values for 5 substances:
-
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> (mac <- with(ids, ids[!is.na(mac) & is.finite(mac), 
-+                        c('etoxid', 'query', 'mac')]))
+R> (mac <- with(ids, ids[!is.na(mac) & is.finite(mac), c("etoxid", "query",
++    "mac")]))
 \end{CodeInput}
@@
 \end{CodeChunk}
-
-The \code{get_etoxid()} function used to search ETOX-IDs returns also the original substance name (\code{query}),
-so that we can easily join the table with MAC values with the measurements table :
+%
+The \code{get_etoxid()} function used to search ETOX-IDs returns also
+the original substance name (\code{query}), so that we can easily join
+the table with MAC values with the measurements table:
+% 
 \begin{CodeChunk}
 \begin{CodeInput}
-R> jagst_eqs <- merge(jagst, mac, by.x = 'substance', by.y = 'query')
+R> jagst_eqs <- merge(jagst, mac, by.x = "substance", by.y = "query")
 R> head(jagst_eqs)
@@
 \end{CodeChunk}
-
+%
 Finally, we can compare the measured value to the MAC, which reveals that there have been no exceedances of these 5 compounds.
@@
 The functions either use simple formatting rules,
-
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
+R> is.inchikey("BQJCRHHNABKAKU-KBQPJGBKS-AN")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> is.cas('64-17-6')
+R> is.cas("64-17-6")
 \end{CodeInput}
@@
 \end{CodeChunk}
-
+%
 or web resources like ChemSpider
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
-+    type = 'chemspider')
+R> is.inchikey("BQJCRHHNABKAKU-KBQPJGBKSA-5", type = "chemspider")
 \end{CodeInput}
@@
 \end{CodeChunk}
-
+%
 \section[Discussion]{Discussion}
 \subsection[Related software]{Related software}
-Within the \proglang{R} ecosystem, there are only a few similar projects:
-\pkg{rpubchem} \citep{rpubchem_2014} provides an interface to PubChem.
-Similarly, \pkg{ChemmineR} \citep{chemminer_2008}, a mature chemo-informatics package, provides an interface to Pubchem. 
-\pkg{webchem} does not provide any chemo-informatic functionality, but integrates access to many data sources.
-\pkg{WikidataR} \citep{wikidatar_2016} provides an interface to wikidata that could be used to retrieve chemical data from Wikipedia.
-However, it does not provide predefined methods for chemical data like \pkg{webchem}.
-Within the \proglang{Python} ecosystem the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
-\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these it provides a unique programmatic interface.
-The Chemical Translation Service \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also one of the sources that can be queried, allows batch conversion of chemical identifiers.
-However, it does not provide access to other data (experimental, modeled or regulatory data).
+Within the \proglang{R} ecosystem, there are only a few similar
+projects: \pkg{rpubchem} \citep{rpubchem_2014} provides an interface
+to PubChem.  Similarly, \pkg{ChemmineR} \citep{chemminer_2008}, a
+mature chemo-informatics package, provides an interface to PubChem.
+\pkg{webchem} does not provide any chemo-informatic functionality, but
+integrates access to many data sources.  \pkg{WikidataR}
+\citep{wikidatar_2016} provides an interface to wikidata that could be
+used to retrieve chemical data from Wikipedia.  However, it does not
+provide predefined methods for chemical data like \pkg{webchem}.
+Within the \proglang{Python} \citep{python} ecosystem the libraries
+\pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy}
+and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those
+outlined here.  \pkg{webchem} is not specialized and tries to
+integrate many data sources and for some of these it provides a unique
+programmatic interface.  The Chemical Translation Service
+\citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also
+one of the sources that can be queried, allows batch conversion of
+chemical identifiers.  However, it does not provide access to other
+data (experimental, modeled or regulatory data).
 
 \subsection[Open Science]{Open Science}
-An increasing number of scientific data is becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
-To be usable for other researchers chemical compounds should be properly identified, not only by chemical names but also with accompanying identifiers like InChIKey, SMILES and authority-assigned identifiers.
-\pkg{webchem} provides an easy way to create such meta tables as shown in Table \ref{tab:comptable} and facilitates chemical data availability to researchers.
-However, good quality of data is crucial for every analysis \citep{Stieger_2014} and additional effort and methods are needed to validate data quality.
+An increasing number of scientific data is becoming publicly available
+\citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011},
+either in public data repositories or as supplements to publications.
+To be usable for other researchers chemical compounds should be
+properly identified, not only by chemical names but also with
+accompanying identifiers like InChIKey, SMILES and authority-assigned
+identifiers.  \pkg{webchem} provides an easy way to create such meta
+tables as shown in Table~\ref{tab:comptable} and facilitates chemical
+data availability to researchers.  However, good quality of data is
+crucial for every analysis \citep{Stieger_2014} and additional effort
+and methods are needed to validate data quality.
 
 \subsection[Further development]{Further development}
-We have outlined only a few use cases that will likely be useful for many researchers.
-Given the huge amount of publicly available information, many other possibilities can be envisioned.
-\pkg{webchem} is currently under active development and several other data sources have not been implemented yet but may be in the future.
-GitHub makes contributing easy and we strongly encourage contribution to the package.
-Moreover, comments, feedback and feature requests are highly welcome.
+We have outlined only a few use cases that will likely be useful for
+many researchers.  Given the huge amount of publicly available
+information, many other possibilities can be envisioned.
+\pkg{webchem} is currently under active development and several other
+data sources have not been implemented yet but may be in the future.
+GitHub makes contributing easy and we strongly encourage contribution
+to the package.  Moreover, comments, feedback and feature requests are
+highly welcome.
 
@@
 Researchers need to have easy access to global knowledge on chemicals.
-\pkg{webchem} can save \emph{"hundreds of working hours"} gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.
+\pkg{webchem} can save \emph{hundreds of working hours} gathering
+this knowledge \citep{Munch_Galizia_2016}, so that researchers can
+focus on other tasks.
 
-\section*{Acknowledgements}
-We thank all resource maintainers for their work making their data open to the public.
-We thank Johannes Ranke and Daniel Münch for their contributions to the webchem package, as well as all users who provided feedback and feature requests. 
-We are grateful to Zacharias Steinmetz for valuable comments on the manuscript.
+\section*{Acknowledgments}
+We thank all resource maintainers for their work making their data
+open to the public.  We thank Johannes Ranke and Daniel M{\"u}nch for
+their contributions to the \pkg{webchem} package, as well as all users who
+provided feedback and feature requests.  We are grateful to Zacharias
+Steinmetz for valuable comments on the manuscript.
 
-\bibliography{references}
+\bibliography{ref}
 

### 2020-01-18 revision3_2581/paper.tex
added new data sources

@@
   developed the open-source \proglang{R} package \pkg{webchem} that
-  allows users to automatically query chemical data from currently 11
+  allows users to automatically query chemical data from currently 14
   web sources.  These cover a broad spectrum of information.  The data
@@
 with current and development \proglang{R} versions), (v) in-source
-documentation (using \pkg{roxygen2}; \citealt{wickham_roxygen2:_2015})
+documentation (using \pkg{roxygen2}; \citealt{wickham_roxygen2})
 and (vi) compliance with a style guide \citep{wickham_advanced_2015}.
@@
 \pkg{RCurl} \citep{lang_rcurl:_2015} and \pkg{httr}
-\citep{wickham_httr} for data transfer, \pkg{stringr}
-\citep{wickham_stringr:_2015} for string handling, \pkg{xml2}
+\citep{wickham_httr} for data transfer, \pkg{dplyr} \citep{wickham_dplyr} for tidying data, \pkg{stringr}
+\citep{wickham_stringr:_2015} for string handling, \pkg{purrr} \citep{wickham_purrr} for working with functions and vectors, \pkg{xml2}
 \citep{wickham_xml2} and \pkg{rvest} \citep{wickham_rvest} for parsing
@@
 and functionality to the public.  Currently, data can be retrieved
-from 11 sources.  These cover a broad spectrum of available data, like
+from 14 sources.  These cover a broad spectrum of available data, like
 identifiers, experimental and predicted properties and regulatory
@@
     interpreter and provides InChI and SMILES identifiers.}
+\item[Flavornet]{\citep{flavornet} Flavornet is a compilation of aroma compounds found in human odor space.}
+\item[NIST]{\citep{nist} The NIST Chemistry WebBook provides access to data compiled and distributed by NIST under the Standard Reference Data Program.}
+\item[ChEBI]{\citep{chebi} Chemical Entities of Biological Interest (ChEBI) is a freely available dictionary of molecular entities focused on "small" chemical compounds.}
+
 \end{description}
@@
 The package is under active development. The latest development
-version is available from GitHub and also permanently available at
-\citet{zenodo}.  This document has been created using \pkg{webchem}
-version 0.4.0.
+version is available from GitHub.  This document has been created using \pkg{webchem}
+version 0.5.0.
 
@@
   etoxid                           match distance                  query
-1   8668     2,4-Dimethylphenol ( 8668 )        0     2,4-Dimethylphenol
+1   8932     2,4-Dimethylphenol ( 8932 )        0     2,4-Dimethylphenol
 2   8494 4-Chlor-2-methylphenol ( 8494 )        0 4-Chlor-2-methylphenol

### 2020-01-20 revision3_2581/paper.tex
Restored reference to Zenodo
Added amount of data accessible

@@
 \item[NIH Chemical Identifier Resolver (CIR)]{\citep{cir} A web
-    service that converts from and to various chemical identifiers.}
+    service that converts from and to various chemical identifiers. The database holds millions of identifier combinations.}
 \item[Chemical Translation Service (CTS)]{
@@
     Agency. Provides basic identifiers, synonyms, ecotoxicological
-    data and quality targets for different countries.}
+    data for over 64,000 entries and quality targets for different countries.}
 \item[PAN Pesticide Database]{ \citep{pan} Information on pesticides --
-    provides basic identifiers, ecotoxicological data and chemical
-    properties.}
+    provides basic identifiers, ecotoxicological data, chemical
+    properties, uses and regulatory status for 6,500 pesticides.}
 \item[SRC Physprop]{\citep{physprop} Contains physical properties for
@@
 \item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
-    information on chemical substances, providing identifiers,
+    information on more than 250 million chemical substances, providing identifiers,
     properties and synonyms. We use an interface to the PUG-REST web
@@
     compendium provides information on pesticide common names,
-    identifiers and classification.}
+    identifiers and classification. The compendium contains more than 1800 active ingredients and more than 350 ester and salt derivatives. }
 \item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
-    database provided by the National Library of Medicine. It provides
+    database provided by the National Library of Medicine. It contains
     identifiers, synonyms, toxicological data and chemical
-    properties.}
+    properties for over 420,000 records.}
 \item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
-    structure database providing access to over 40 million
-    structures. It provides identifiers, properties and can also be
+    structure database providing access to over 67 million
+    structures from hundreds of data sources. It provides identifiers, properties and can also be
     used to convert identifiers.}
@@
     interpreter and provides InChI and SMILES identifiers.}
-\item[Flavornet]{\citep{flavornet} Flavornet is a compilation of aroma compounds found in human odor space.}
-\item[NIST]{\citep{nist} The NIST Chemistry WebBook provides access to data compiled and distributed by NIST under the Standard Reference Data Program.}
+\item[Flavornet]{\citep{flavornet} Flavornet is a compilation of 738 aroma compounds found in human odor space.}
+\item[NIST]{\citep{nist} The NIST Chemistry WebBook provides access to data compiled and distributed by NIST under the Standard Reference Data Program.
+The WebBook provides chemical and physical property data on over 40000 compounds.}
 \item[ChEBI]{\citep{chebi} Chemical Entities of Biological Interest (ChEBI) is a freely available dictionary of molecular entities focused on "small" chemical compounds.}
@@
 The package is under active development. The latest development
-version is available from GitHub.  This document has been created using \pkg{webchem}
+version is available from GitHub and also permanently available at \citet{zenodo}.  This document has been created using \pkg{webchem}
 version 0.5.0.

### 2020-01-23 revision3_2581/paper.tex
Added amount of data available

@@
 \item[Wikidata]{\citep{wiki} Wikipedia contains information for over
-    15,000 chemicals
+    210,000 chemicals
     \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
@@
 \item[NIST]{\citep{nist} The NIST Chemistry WebBook provides access to data compiled and distributed by NIST under the Standard Reference Data Program.
-The WebBook provides chemical and physical property data on over 40000 compounds.}
-\item[ChEBI]{\citep{chebi} Chemical Entities of Biological Interest (ChEBI) is a freely available dictionary of molecular entities focused on "small" chemical compounds.}
+The WebBook provides chemical and physical property data on over 40,000 compounds.}
+\item[ChEBI]{\citep{chebi} Chemical Entities of Biological Interest (ChEBI) is a freely available dictionary of molecular entities focused on "small" chemical compounds.
+The database contains more than 56,000 compounds.}
 

### 2020-01-30 revision3_2581/paper.tex
removed and replaced Physprop partition coef data with PubChem data

@@
     properties, uses and regulatory status for 6,500 pesticides.}
-\item[SRC Physprop]{\citep{physprop} Contains physical properties for
-    over 41,000 chemicals.  Physical properties collected from a wide
-    variety of sources including experimental and modeled values.}
 \item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
@@
 Some data sources also provide data on chemical properties that can be
-queried.  Here we query for the \code{lc50} data the
-$\mathrm{log}~P_{oct/wat}$ from the SRC Physprop database to build a
+queried.  Here we query for the \code{lc50} data predicted
+$\mathrm{log}~P_{oct/wat}$ from the PubChem database to build a
 simple quantitative structure-activity relationship (QSAR) to predict
-toxicity.
+toxicity. The resulting data and model are displayed in Figure~\ref{fig:fig3}.
 %
@@
 \begin{CodeInput}
-R> pp_data <- pp_query(lc50$cas)
+R> cid <- get_cid(lc50$cas)
+R> pc_data <- pc_prop(cid)
+R> lc50$logp <- pc_data$XLogP
 \end{CodeInput}
@@
 %
-The database contains predicted and experimental values.
-Extracting $\mathrm{log}~P_{oct/wat}$ from the data object is slightly more complicated,  
-because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).
-%
-\begin{CodeChunk}
-\begin{CodeInput}
-R> lc50$logp <- sapply(pp_data, function(y) {
-+    if (length(y) == 1 && is.na(y))
-+      return(NA)
-+    y$prop$value[y$prop$variable == "Log P (octanol-water)"]
-+  })
-\end{CodeInput}
-\end{CodeChunk}
-%
-We opted for this more complex approach, because the information available is very diverse and we cannot provide an extractor function for each purpose.
-Moreover, it provides users with high flexibility regarding organization of their data. 
-Nevertheless, in the documentation of each function we provide examples on how to extract more complicated parts of the data.
-The resulting data and model are displayed in Figure~\ref{fig:fig3}.
 
@@
     pesticides by $\mathrm{log}~P$.  $\mathrm{log}~P$ values have been
-    retrieved from the SRC Physprop database (97 experimental data, 9
-    estimated data and 18 substances without data).  Blue line
+    retrieved from the Pubchem database (118 estimated data and 6 substances without data).  Blue line
     indicates the regression model
-    ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log}~P$,
-    RMSE = 1.45).}
+    ($\mathrm{log~LC}_{50} = 2.86\ensuremath{-0.36}~\mathrm{log}~P$,
+    RMSE = 1.47).}
 \label{fig:fig3}

### 2020-01-30 revision3_2581/Code/code.R

@@ -96,43 +96,30 @@ p <- ggplot(lc50, aes(x = type, y = value)) +
   theme(text = element_text(size = 16))
 p
 
-
-
 # Use Case III: Querying partitioning coefficients ------------------------
-# query PHYSPROP DATABASE
-pp_data <- pp_query(lc50$cas)
+# query PubChem DATABASE
+cid <- get_cid(lc50$cas, first = TRUE)
+pc_data <- pc_prop(cid)
 # lookat internal structure
-str(pp_data[[1]])
+str(pc_data)
 # extract logP from properties data.frame
-lc50$logp <- sapply(pp_data, function(y){
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$value[y$prop$variable == "Log P (octanol-water)"]
-})
+lc50$logp <- pc_data$XLogP
 
 # model
 mod <- lm(log10(value) ~ logp, data = lc50)
 coef(mod) # coeficients
 summary(mod)$sigma # RMSE
 
-# check data types
-logp_type <- sapply(pp_data, function(y){
-  if (length(y) == 1 && is.na(y))
-    return(NA)
-  y$prop$type[y$prop$variable == "Log P (octanol-water)"]
-})
-table(logp_type) # experimental / predicted data
-sum(is.na(logp_type)) # NA = compounds not found
-
 # plot
 p <- ggplot(lc50, aes(x = logp, y = value)) +
   geom_point() +
-  scale_y_log10() +
   geom_smooth(method = "lm") +
+  scale_y_log10() +
+  scale_x_continuous(breaks = c(0, 2, 4, 6),
+                     labels = format(10^c(0, 2, 4, 6), 
+                                     scientific = FALSE)) +
   labs(x = "P", y = expression(LC[50])) +
-  theme_bw() +
-  scale_x_continuous(breaks = c(0, 2, 4, 6), labels = format(10^c(0, 2, 4, 6), 
-                                                             scientific = FALSE))
+  theme_bw()
 p
 
 # Use Case IV: Regulatory information -------------------------------------

### 2020-01-31 revision3_2581/paper.tex
ChEBI example

@@
 
+\subsection[Identify common names and roles of chemicals]{Identify common names and roles of chemicals}
+\pkg{webchem} can query the ChEBI database \citep{chebi} to retrieve common names and roles of chemicals on biota.
+Therefore, we first order the \code{lc50} data set by the three most toxic (i.e.
+lowest value) chemicals and then use \code{chebi_lite_entity()} to retrieve ChEBI identifiers and substance names.
+The function returns a list of tables containing the ChEBI identifier, the chemical name, a search score and entity stars for the found entry.
+The searchscore is calculated by ChEBI and represents how well the query string matches one of the database entries.
+The entitystar column reflects the quality of the ChEBI data, also provided by the database.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> lite <- chebi_lite_entity(lc50_3$cas)
+R> lite <- lapply(lite, function(x) x[ which.max(x$searchscore), ]) # highest seachscore
+R> lite <- do.call(rbind, lite) # bind to data.frmae
+R> lite$cas <- rownames(lite)
+R> lite$chebiasciiname[4] <- "tefluthrin" # erronous result
+\end{CodeInput}
+\end{CodeOutput}
+
+We can now use the ChEBI identifiers to query the complete entitiy by using \code{chebi_comp_entity()}.
+From the result, we extract the parents table and refine the data to chemical roles.
+Subsequently, we prepare the query result and merge the data with the \code{lite} object to include the CAS numbers (i.e.
+our initial query strings) to the result.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> comp <- chebi_comp_entity(lite$chebiid) # ChEBI ids
+R> comp <- do.call(rbind, lapply(comp, `[[`, "parents")) # extract parents table
+R> comp$chebiid <- gsub('\\.[0-9]+', '', rownames(comp))
+R> comp <- comp[ comp$type == "has role", ] # refine to chemical roles
+R> role <- aggregate(chebiName ~ chebiid,
+                     data = comp,
+                     FUN = function(x) paste0(x, collapse = ', '))
+R> merge(lite[ names(lite) %in% c('cas', 'chebiid', 'chebiasciiname') ],
+         role,
+         by = 'chebiid')
+\end{CodeInput}
+\end{CodeOutput}
+
+In doing so we retrieve Table~\ref{tab:chebi}, which provides information on the individual roles of the chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D.
+magna} in our table is designed to be used as an acari- and insecticide, which inhibts the acetylcholinesterase enzyme.
+
+\begin{table}[t!]
+\centering
+\begin{tabular}{p{2cm}p{2cm}p{8cm}}
+  \hline
+  cas & name & roles \\
+  \hline
+  563-12-2 & ethion & insecticide \newline environmental contaminant \newline EC 3.1.1.7 (acetylcholinesterase) inhibitor \newline acaricide \newline agrochemical \\
+  96182-53-5 & tebupirimfos & EC 3.1.1.7 (acetylcholinesterase) inhibitor \\
+  3383-96-8 & temephos & EC 3.1.1.7 (acetylcholinesterase) inhibitor \newline acaricide \newline agrochemical \newline ectoparasiticide \\
+\hline
+\end{tabular}
+\caption{CAS numbers, names and chemical roles for the three most toxic substances towards \textit{D. magna} according to ChEBI.}
+\label{tab:chebi}
+\end{table}
+
 \subsection[Querying partitioning coefficients]{Querying partitioning coefficients}

### 2020-01-31 revision3_2581/Code/code.R

@@ -63,8 +63,6 @@ res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles,
                   stringsAsFactors = FALSE)
 head(res)
 
-
-
 # Use Case II: Toxicity of different pesticide groups ---------------------
 # query the pesticide compendium using CAS-numbers
 aw_data <- aw_query(lc50$cas, type = "cas")
@@ -96,7 +94,23 @@ p <- ggplot(lc50, aes(x = type, y = value)) +
   theme(text = element_text(size = 16))
 p
 
-# Use Case III: Querying partitioning coefficients ------------------------
+# Use Case III: Name 10 most toxic chemicals ------------------------------
+lc50_3 <- lc50[ order(lc50$value), ][1:3, ]
+# query ChEBI names from CAS
+lite <- chebi_lite_entity(lc50_3$cas)
+lite <- lapply(lite, function(x) x[ which.max(x$searchscore), ]) # select entry with the highest seachscore
+lite <- do.call(rbind, lite) # bind to data.frmae
+lite$cas <- rownames(lite)
+# query ChEBI complete entity
+comp <- chebi_comp_entity(lite$chebiid) # use ChEBI ids to query the service
+comp <- do.call(rbind, lapply(comp, `[[`, "parents")) # extract parents list and bind them to a data.frame
+comp$chebiid <- gsub('\\.[0-9]+', '', rownames(comp))
+comp <- comp[ comp$type == "has role", ] # refine to chemical roles
+role <- aggregate(chebiName ~ chebiid, data = comp, FUN = function(x) paste0(x, collapse = ', '))
+role <- merge(lite[ names(lite) %in% c('cas', 'chebiid', 'chebiasciiname') ], role, by = 'chebiid')
+setNames(role, c('chebiid', 'name', 'cas', 'roles'))
+
+# Use Case IV: Querying partitioning coefficients -------------------------
 # query PubChem DATABASE
 cid <- get_cid(lc50$cas, first = TRUE)
 pc_data <- pc_prop(cid)
@@ -122,7 +136,7 @@ p <- ggplot(lc50, aes(x = logp, y = value)) +
   theme_bw()
 p
 
-# Use Case IV: Regulatory information -------------------------------------
+# Use Case V: Regulatory information --------------------------------------
 # search EQS from ETOX using the already queried ETOX_IDs
 eqs <- etox_targets(ids$etoxid)
 # extract only MAC-EQS for the EU from the results

### 2020-02-11 revision3_2581/paper.tex
Updated table outputs and added tests

@@
 source using these CAS numbers (see Figure~\ref{fig:fig1}), like
-PubChem
+PubChem.
 %
@@
 %
-or ChemSpider
+We can use SMILES to query ChemSpiderIDs and subsequently convert them to InChiKeys.
 %
@@
 \begin{CodeInput}
-R> csids <- get_csid(etox_cas, token = token)
-R> cs_data <- cs_compinfo(csids, token = token)
-R> cs_inchikey <- inchikey(cs_data)
+R> csids <- get_csid(pc_smiles, from = "smiles")$csid
+R> cs_inchikey <- cs_convert(csids, from = "csid", to = "inchikey")
 \end{CodeInput}
@@
 Note that in order to use the ChemSpider functions, a personal
-authentication key (\code{token}) is needed, which can be retrieved
+authentication key is needed, which can be retrieved
 from the ChemSpider web page.  Finally, we obtain a compound table
@@
 \end{CodeInput}
-\end{CodeOutput}
+\end{CodeChunk}
 
@@
 Here, the column \code{Designation} defines the type of EQS and \code{Value_Target_LR} contains the value.
-Unfortunately, we only found MAC-EQS values for 5 substances:
+Unfortunately, we only found MAC-EQS values for 6 substances:
 %
@@
 12   7442 Isoproturon  1.000
+38   7571     Simazin  4.000
 29   8756   Terbutryn  0.034
@@
 4   Atrazin 2013-04-23 0.0048    =   8397   2
-5   Atrazin 2013-11-05 0.0036    =   8397   2
+5   Atrazin 2013-06-18 0.0048    =   8397   2
 6   Atrazin 2013-07-16 0.0052    =   8397   2
@@
 %
-Finally, we can compare the measured value to the MAC, which reveals that there have been no exceedances of these 5 compounds.
+Finally, we can compare the measured value to the MAC, which reveals that there have been no exceedances of these 6 compounds.
 

### 2020-02-11 revision3_2581/Code/code.R

@@ -1,18 +1,18 @@
 # -------------------------------------------------------------------------
 # Supplemental R Code to reproduce the results of
-# Szöcs, Schäfer. webchem: An R Package to Retrieve Chemical Information 
+# Szöcs, Schäfer. webchem: An R Package to Retrieve Chemical Information
 #   from the Web
 # ------------------------------------------------------------------------
 
 
 # Install webchem from CRAN -----------------------------------------------
-install.packages("webchem")
+#install.packages("webchem")
 
 
 # load packages -----------------------------------------------------------
-library("webchem")
+#library("webchem")
 library("ggplot2")
-
+library("testthat")
 
 # load datasets -----------------------------------------------------------
 # load jagst dataset
@@ -47,25 +47,42 @@ cids <- get_cid(etox_cas)
 pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
 pc_smiles <- smiles(pc_data)
 
-# ChemSpider needs a security token
-# this is a webchem specific token. 
+# ChemSpider needs an API key.
+# This is a webchem specific token.
 # Please use it only for reproduction / testing.
 token <- "37bf5e57-9091-42f5-9274-650a64398aaf"
 # query InChiKey from ChemSpider
-csids <- get_csid(etox_cas, token = token)
-cs_data <- cs_compinfo(csids, token = token)
-cs_inchikey <- inchikey(cs_data)
+#csids <- get_csid(pc_smiles, from = "smiles", apikey = apikey)
+#cs_inchikey <- cs_convert(csids, from = "csid", to = "inchikey", apikey = apikey)
 
 # combine into single data.frame
-res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-                  cid = pc_data$CID, inchikey = cs_inchikey, 
-                  csid = cs_data$csid, 
-                  stringsAsFactors = FALSE)
-head(res)
+#res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles,
+#                  cid = pc_data$CID, inchikey = cs_inchikey,
+#                  csid = cs_data$csid,
+#                  stringsAsFactors = FALSE)
+#head(res)
+
+# Automatic tests to test if results are reproduced correctly
[…]

### 2020-02-11 revision3_2581/Code/code.R
Updated table outputs and added tests

@@ -1,18 +1,18 @@
 # -------------------------------------------------------------------------
 # Supplemental R Code to reproduce the results of
-# Szöcs, Schäfer. webchem: An R Package to Retrieve Chemical Information 
+# Szöcs, Schäfer. webchem: An R Package to Retrieve Chemical Information
 #   from the Web
 # ------------------------------------------------------------------------
 
 
 # Install webchem from CRAN -----------------------------------------------
-install.packages("webchem")
+#install.packages("webchem")
 
 
 # load packages -----------------------------------------------------------
-library("webchem")
+#library("webchem")
 library("ggplot2")
-
+library("testthat")
 
 # load datasets -----------------------------------------------------------
 # load jagst dataset
@@ -47,25 +47,42 @@ cids <- get_cid(etox_cas)
 pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
 pc_smiles <- smiles(pc_data)
 
-# ChemSpider needs a security token
-# this is a webchem specific token. 
+# ChemSpider needs an API key.
+# This is a webchem specific token.
 # Please use it only for reproduction / testing.
 token <- "37bf5e57-9091-42f5-9274-650a64398aaf"
 # query InChiKey from ChemSpider
-csids <- get_csid(etox_cas, token = token)
-cs_data <- cs_compinfo(csids, token = token)
-cs_inchikey <- inchikey(cs_data)
+#csids <- get_csid(pc_smiles, from = "smiles", apikey = apikey)
+#cs_inchikey <- cs_convert(csids, from = "csid", to = "inchikey", apikey = apikey)
 
 # combine into single data.frame
-res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-                  cid = pc_data$CID, inchikey = cs_inchikey, 
-                  csid = cs_data$csid, 
-                  stringsAsFactors = FALSE)
-head(res)
+#res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles,
+#                  cid = pc_data$CID, inchikey = cs_inchikey,
+#                  csid = cs_data$csid,
+#                  stringsAsFactors = FALSE)
+#head(res)
+
+# Automatic tests to test if results are reproduced correctly
[…]

### 2020-02-17 revision3_2581/paper.tex
ChEBI v0.5.0 changes

--- 2020-01-31 revision3_2581/paper.tex
@@
 Therefore, we first order the \code{lc50} data set by the three most toxic (i.e.
-lowest value) chemicals and then use \code{chebi_lite_entity()} to retrieve ChEBI identifiers and substance names.
+lowest value) chemicals and then use \code{get_chebiid()} to retrieve ChEBI identifiers and substance names.
 The function returns a list of tables containing the ChEBI identifier, the chemical name, a search score and entity stars for the found entry.
@@
 \begin{CodeInput}
-R> lite <- chebi_lite_entity(lc50_3$cas)
-R> lite <- lapply(lite, function(x) x[ which.max(x$searchscore), ]) # highest seachscore
+R> lite <- get_chebiid(lc50_3$cas)
 R> lite <- do.call(rbind, lite) # bind to data.frmae

### 2020-02-17 revision3_2581/Code/code.R

@@ -97,8 +97,7 @@ p
 # Use Case III: Name 10 most toxic chemicals ------------------------------
 lc50_3 <- lc50[ order(lc50$value), ][1:3, ]
 # query ChEBI names from CAS
-lite <- chebi_lite_entity(lc50_3$cas)
-lite <- lapply(lite, function(x) x[ which.max(x$searchscore), ]) # select entry with the highest seachscore
+lite <- get_chebiid(lc50_3$cas, match = 'best')
 lite <- do.call(rbind, lite) # bind to data.frmae
 lite$cas <- rownames(lite)
 # query ChEBI complete entity

### 2020-03-16 revision3_2581/paper.tex
Rearranged data sources to match Figure 1
Added SRS to the article
Update year for online references

@@
 \author{Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And
+	Tam\'as Stirling\\Independent Developer \And
         Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
@@
 
-\Plainauthor{Eduard Sz\"ocs, Ralf B. Sch\"afer}
+\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Ralf B. Sch\"afer}
 
@@
     service that converts from and to various chemical identifiers. The database holds millions of identifier combinations.}
+\item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
+	structure database providing access to over 67 million
+	structures from hundreds of data sources. It provides identifiers, properties and can also be
+	used to convert identifiers.}
 \item[Chemical Translation Service (CTS)]{
@@
     data for over 64,000 entries and quality targets for different countries.}
-\item[PAN Pesticide Database]{ \citep{pan} Information on pesticides --
-    provides basic identifiers, ecotoxicological data, chemical
-    properties, uses and regulatory status for 6,500 pesticides.}
+\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
+	database provided by the National Library of Medicine. It contains
+	identifiers, synonyms, toxicological data and chemical
+	properties for over 420,000 records.}
+\item[OPSIN]{\citep{Lowe_Corbett_Murray-Rust_Glen_2011} The Open
+	Parser for Systematic IUPAC nomenclature is a chemical name
+	interpreter and provides InChI and SMILES identifiers.}
+\item[Compendium of Pesticide Common Names]{\citep{wood} The
+	compendium provides information on pesticide common names,
+	identifiers and classification. The compendium contains more than 1800 active ingredients and more than 350 ester and salt derivatives. }
+\item[Wikidata]{\citep{wiki} Wikipedia contains information for over
+	210,000 chemicals
+	\citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
+	\pkg{webchem} can only query chemical identifiers.}
 \item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
@@
     service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
-\item[Wikidata]{\citep{wiki} Wikipedia contains information for over
-    210,000 chemicals
-    \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
-    \pkg{webchem} can only query chemical identifiers.}
-\item[Compendium of Pesticide Common Names]{\citep{wood} The
-    compendium provides information on pesticide common names,
-    identifiers and classification. The compendium contains more than 1800 active ingredients and more than 350 ester and salt derivatives. }
-\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
-    database provided by the National Library of Medicine. It contains
-    identifiers, synonyms, toxicological data and chemical
-    properties for over 420,000 records.}
-\item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
-    structure database providing access to over 67 million
-    structures from hundreds of data sources. It provides identifiers, properties and can also be
-    used to convert identifiers.}
-\item[OPSIN]{\citep{Lowe_Corbett_Murray-Rust_Glen_2011} The Open
-    Parser for Systematic IUPAC nomenclature is a chemical name
-    interpreter and provides InChI and SMILES identifiers.}
+\item[PAN Pesticide Database]{ \citep{pan} Information on pesticides --
+	provides basic identifiers, ecotoxicological data, chemical
+	properties, uses and regulatory status for 6,500 pesticides.}
 \item[Flavornet]{\citep{flavornet} Flavornet is a compilation of 738 aroma compounds found in human odor space.}
@@
 The database contains more than 56,000 compounds.}
+\item[SRS]{\citep{srs} Substance Registry Services is the US EPA central system for information about substances that are tracked or regulated by EPA or other sources.}
 

### 2020-03-21 revision3_2581/paper.tex
Updated code in the article

@@
 % citation alias for LUBW institution
-\defcitealias{lubw_2016}{LUBW (2016)}
+\defcitealias{lubw_2016}{LUBW (2020)}
 % \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 available under an MIT license.  The development repository is hosted
-on \citet{github} and a stable version is released on the
+on GitHub \citep{github} and a stable version is released on the
 Comprehensive \proglang{R} Archive Network (CRAN) and available at
@@
 a data source provider, where available.  For sources where an API is
-lacking, data is directly searched and extracted from the web pages,
+lacking, but policies of the service provider permit programmable access, data is directly searched and extracted from the web pages,
 analogous to manual interaction with a website.
@@
   etoxid                           match distance                  query
-1   8932     2,4-Dimethylphenol ( 8932 )        0     2,4-Dimethylphenol
+1   8668            2,4-Xylenol ( 8668 )        0     2,4-Dimethylphenol
 2   8494 4-Chlor-2-methylphenol ( 8494 )        0 4-Chlor-2-methylphenol
@@
 \begin{CodeInput}
-R> csids <- get_csid(pc_smiles, from = "smiles")$csid
-R> cs_inchikey <- cs_convert(csids, from = "csid", to = "inchikey")
+R> csids <- get_csid(pc_smiles, from = "smiles")
+R> cs_inchikey <- cs_convert(csids$csid, from = "csid", to = "inchikey")
 \end{CodeInput}
@@
 R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
-+    cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
-+    stringsAsFactors = FALSE)
++    cid = pc_data$CID, inchikey = cs_inchikey, csid = csids$csid, 
++    stringsAsFactors = FALSE, row.names = NULL)
 \end{CodeInput}
@@
 \subsection[Identify common names and roles of chemicals]{Identify common names and roles of chemicals}
+We might be interested in common names and roles of the most toxic chemicals in the \code{lc50} data set, i.e., chemicals with the lowest $LC_{50}$ values.
 \pkg{webchem} can query the ChEBI database \citep{chebi} to retrieve common names and roles of chemicals on biota.
-Therefore, we first order the \code{lc50} data set by the three most toxic (i.e.
-lowest value) chemicals and then use \code{get_chebiid()} to retrieve ChEBI identifiers and substance names.
-The function returns a list of tables containing the ChEBI identifier, the chemical name, a search score and entity stars for the found entry.
-The searchscore is calculated by ChEBI and represents how well the query string matches one of the database entries.
-The entitystar column reflects the quality of the ChEBI data, also provided by the database.
+We can extract the CAS numbers of the three most toxic chemicals, and then use \code{get_chebiid()} to retrieve ChEBI identifiers and substance names.
+We can then use the acquired ChEBI identifiers to query the complete entity by using \code{chebi_comp_entity()}.
+From the result, we can extract the parents table and refine the data to chemical roles.
 
@@
 \begin{CodeInput}
-R> lite <- get_chebiid(lc50_3$cas)
-R> lite <- do.call(rbind, lite) # bind to data.frmae
-R> lite$cas <- rownames(lite)
-R> lite$chebiasciiname[4] <- "tefluthrin" # erronous result
+R> cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
+R> chebiids <- get_chebiid(cas_rns)
+R> comp <- chebi_comp_entity(chebiids$chebiid)
+R> pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
 \end{CodeInput}
@@
 
-We can now use the ChEBI identifiers to query the complete entitiy by using \code{chebi_comp_entity()}.
-From the result, we extract the parents table and refine the data to chemical roles.
-Subsequently, we prepare the query result and merge the data with the \code{lite} object to include the CAS numbers (i.e.
-our initial query strings) to the result.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> comp <- chebi_comp_entity(lite$chebiid) # ChEBI ids
-R> comp <- do.call(rbind, lapply(comp, `[[`, "parents")) # extract parents table
-R> comp$chebiid <- gsub('\\.[0-9]+', '', rownames(comp))
-R> comp <- comp[ comp$type == "has role", ] # refine to chemical roles
-R> role <- aggregate(chebiName ~ chebiid,
-                     data = comp,
-                     FUN = function(x) paste0(x, collapse = ', '))
-R> merge(lite[ names(lite) %in% c('cas', 'chebiid', 'chebiasciiname') ],
-         role,
-         by = 'chebiid')
-\end{CodeInput}
-\end{CodeChunk}
-
-In doing so we retrieve Table~\ref{tab:chebi}, which provides information on the individual roles of the chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D.
-magna} in our table is designed to be used as an acari- and insecticide, which inhibts the acetylcholinesterase enzyme.
+Finally we can compile the data into a table.
+Table~\ref{tab:chebi} provides information on the individual roles of the three most toxic chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D.
+magna} in our table, is designed to be used as an acari- and insecticide, which inhibts the acetylcholinesterase enzyme.
 
@@
 \end{tabular}
-\caption{CAS numbers, names and chemical roles for the three most toxic substances towards \textit{D. magna} according to ChEBI.}
+\caption{CAS numbers, names and chemical roles from ChEBI for the three most toxic substances in the \code{lc50} data set towards \textit{D. magna}.}
 \label{tab:chebi}

### 2020-03-21 revision3_2581/Code/code.R

@@ -1,16 +1,14 @@
 # -------------------------------------------------------------------------
-# Supplemental R Code to reproduce the results of
-# Szöcs, Schäfer. webchem: An R Package to Retrieve Chemical Information
-#   from the Web
+# Supplemental R Code to reproduce the results of Szöcs, Stirling, Schäfer. 
+# webchem: An R Package to Retrieve Chemical Information from the Web
 # ------------------------------------------------------------------------
 
 
 # Install webchem from CRAN -----------------------------------------------
 #install.packages("webchem")
 
-
 # load packages -----------------------------------------------------------
-#library("webchem")
+library("webchem")
 library("ggplot2")
 library("testthat")
 
@@ -42,57 +40,46 @@ head(etox_cas)
 
 # query other identifiers from other resources
 
+# get_cid() returned the PubChem ID of sodium when the query was NA. This was
+# fixed in the dev version. To install the dev version:
+
+library(devtools)
+install_github("ropensci/webchem")
+
 # query SMILES from PubChem
 cids <- get_cid(etox_cas)
 pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
 pc_smiles <- smiles(pc_data)
 
-# ChemSpider needs an API key.
-# This is a webchem specific token.
-# Please use it only for reproduction / testing.
-token <- "37bf5e57-9091-42f5-9274-650a64398aaf"
-# query InChiKey from ChemSpider
-#csids <- get_csid(pc_smiles, from = "smiles", apikey = apikey)
-#cs_inchikey <- cs_convert(csids, from = "csid", to = "inchikey", apikey = apikey)
+# ChemSpider needs an API key and RSC explicitly asks users to keep the key
+# secure. In order to test these functions, please register for an API key.
+csids <- get_csid(pc_smiles, from = "smiles", apikey = apikey)
+cs_inchikey <- cs_convert(csids$csid, from = "csid", to = "inchikey",
+                          apikey = apikey)
+
+
 
 # combine into single data.frame
-#res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles,
-#                  cid = pc_data$CID, inchikey = cs_inchikey,
[…]

### 2020-03-21 revision3_2581/paper.tex
Updated affiliation

@@
   E-mail: \email{szoecs@uni-landau.de}\\
-  URL: \url{https://edild.github.io}
+  URL: \url{https://edild.github.io}\\
+  \\
+  Tam\'as Stirling\\
+  Budapest, Hungary\\
+  E-mail: \email{stirling.tamas@gmail.com}\\
+  URL: \url{https://github.com/stitam}
 }

### 2020-03-21 revision3_2581/Code/code.R
Uncomment a line in code.R

@@ -5,7 +5,7 @@
 
 
 # Install webchem from CRAN -----------------------------------------------
-#install.packages("webchem")
+install.packages("webchem")
 
 # load packages -----------------------------------------------------------
 library("webchem")

### 2020-03-21 revision3_2581/Code/code.R
Fix some typos in code.R

@@ -10,7 +10,6 @@ install.packages("webchem")
 # load packages -----------------------------------------------------------
 library("webchem")
 library("ggplot2")
-library("testthat")
 
 # load datasets -----------------------------------------------------------
 # load jagst dataset
@@ -157,7 +156,7 @@ ids$mac <- sapply(eqs, function(y){
 jagst_eqs <- merge(jagst, mac, by.x = "substance", by.y = "query")
 head(jagst_eqs)
 
-# Utilitymac functions -------------------------------------------------------
+# Utility functions -------------------------------------------------------
 # simple formatting check
 is.inchikey("BQJCRHHNABKAKU-KBQPJGBKS-AN")
 # formatting check

### 2020-03-23 revision3_2581/paper.tex
added authorship

@@
 	      Tam\'as Stirling\\Independent Developer \And
+	      Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
         Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
@@
 
-\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Ralf B. Sch\"afer}
+\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Andreas Scharm\"uller, Ralf B. Sch\"afer}
 

### 2020-03-24 revision3_2581/paper.tex
changes paper

@@
 HTML and XML, \pkg{jsonlite} \citep{ooms_jsonlite_2014} for parsing
-JSON, \pkg{rcdk} \citep{guha_rcdk} for parsing SMILES.  For parsing
+JSON, \pkg{rcdk} \citep{guha_rcdk} for parsing SMILES. To parse
 molfiles we use a lightweight implementation in package \pkg{RMol}
@@
 \begin{CodeInput}
-R> igroup <- sapply(aw_data, function(y) y$subactivity[1])
+R> igroup <- sapply(aw_data, function(y) {
+R>   if (!is.na(y)) {
+R>     y$subactivity[1]  
+R>   } else {
+R>     NA
+R>   }
+R>  })
 R> igroup[1:3]
@@
 We might be interested in common names and roles of the most toxic chemicals in the \code{lc50} data set, i.e., chemicals with the lowest $LC_{50}$ values.
-\pkg{webchem} can query the ChEBI database \citep{chebi} to retrieve common names and roles of chemicals on biota.
-We can extract the CAS numbers of the three most toxic chemicals, and then use \code{get_chebiid()} to retrieve ChEBI identifiers and substance names.
-We can then use the acquired ChEBI identifiers to query the complete entity by using \code{chebi_comp_entity()}.
-From the result, we can extract the parents table and refine the data to chemical roles.
+\pkg{webchem} can query the ChEBI database \citep{chebi} to retrieve common names and roles of chemicals.
+We extract the CAS numbers of the three most toxic chemicals, and then use \code{get_chebiid()} to retrieve ChEBI identifiers and substance names.
+Subsequently, we use the acquired ChEBI identifiers to query the complete entity by using \code{chebi_comp_entity()}.
+From the result, we extract the parents table and refine the data to chemical roles.
 
@@
 R> chebiids <- get_chebiid(cas_rns)
+R> chebiids <- do.call(rbind, chebiids)
 R> comp <- chebi_comp_entity(chebiids$chebiid)

### 2020-03-26 revision3_2581/paper.tex
update wikidata desc
Removed sentence about only querying chemical identifiers.  Seems like you can query chemical names too, but results may vary.
update description of what webchem returns
webchem does not currently "always return a named list". Based on discussion on the webchem github, I think this change is accurate and will continue to be accurate in the future.

--- 2020-03-21 revision3_2581/paper.tex
@@
 	210,000 chemicals
-	\citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
-	\pkg{webchem} can only query chemical identifiers.}
+	\citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}.}
 \item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
@@
 %
-\pkg{webchem} always returns a named list (one entry for each substance) and the available information content can be very voluminous.
+When possible, \pkg{webchem} returns a data frame with one row per substance.
+However, data from some sources can be very voluminous and not tabular.
+In these cases \pkg{webchem} always returns a named list (one entry for each substance).
 Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.

### 2020-03-27 revision3_2581/paper.tex
defined MAC

--- 2020-03-24 revision3_2581/paper.tex
@@
 We re-use the above queried ETOX-IDs to obtain further information
-from ETOX, namely the MAC-EQS:
+from ETOX, namely the Maximum Acceptible Concentration EQS (MAC-EQS):
 % 

### 2020-03-28 revision3_2581/Code/code.R
Update authors, references, etc.

@@ -1,11 +1,17 @@
 # -------------------------------------------------------------------------
-# Supplemental R Code to reproduce the results of Szöcs, Stirling, Scharmüller, Schäfer. 
-# webchem: An R Package to Retrieve Chemical Information from the Web
+# Supplemental R Code to reproduce the results of Szöcs, Stirling, Scott,
+# Scharmüller, Schäfer. webchem: An R Package to Retrieve Chemical Information
+# from the Web
 # ------------------------------------------------------------------------
 
 
 # Install webchem from CRAN -----------------------------------------------
 install.packages("webchem")
+
+# load development version -----------------------------------------------
+# fixes some small bugs
+library(devtools)
+install_github("ropensci/webchem")
 
 # load packages -----------------------------------------------------------
 library(webchem)
@@ -28,7 +34,6 @@ subs <- unique(jagst$substance)
 
 # search ETOX IDs, keeping only the best match
 ids <- get_etoxid(subs, match = "best")
-ids <- ids[!ids$etoxid %in% c(99015, 97094), ]
 head(ids)
 
 # use this id to query information from ETOX
@@ -38,11 +43,6 @@ etox_cas <- cas(etox_data)
 head(etox_cas)
 
 # query other identifiers from other resources
-
-# get_cid() returned the PubChem ID of sodium when the query was NA. This was
-# fixed in the dev version. To install the dev version:
-library(devtools)
-install_github("ropensci/webchem")
 
 # query SMILES from PubChem
 cids <- get_cid(etox_cas)
@@ -107,7 +107,6 @@ p
 # Use Case III: Name 10 most toxic chemicals ------------------------------
 cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
 chebiids <- get_chebiid(cas_rns)
-chebiids <- do.call(rbind, chebiids)
 comp <- chebi_comp_entity(chebiids$chebiid)
 pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
 
@@ -166,4 +165,3 @@ is.cas("64-17-6")
 
 # or using ChemSpider
 is.inchikey("BQJCRHHNABKAKU-KBQPJGBKSA-5", type = "chemspider")
-

### 2020-03-29 revision3_2581/Code/code.R
updated code

@@ -3,9 +3,9 @@
 # webchem: An R Package to Retrieve Chemical Information from the Web
 # ------------------------------------------------------------------------
 
-
 # Install webchem from CRAN -----------------------------------------------
-install.packages("webchem")
+library(devtools)
+install_github("ropensci/webchem")
 
 # load packages -----------------------------------------------------------
 library(webchem)
@@ -21,14 +21,12 @@ head(jagst)
 data("lc50", package = "webchem")
 head(lc50)
 
-
-# Use Case 1: Query identifiers -------------------------------------------
+# Use Case I: Query identifiers -------------------------------------------
 # unique substance names
 subs <- unique(jagst$substance)
 
 # search ETOX IDs, keeping only the best match
 ids <- get_etoxid(subs, match = "best")
-ids <- ids[!ids$etoxid %in% c(99015, 97094), ]
 head(ids)
 
 # use this id to query information from ETOX
@@ -38,12 +36,8 @@ etox_cas <- cas(etox_data)
 head(etox_cas)
 
 # query other identifiers from other resources
-
 # get_cid() returned the PubChem ID of sodium when the query was NA. This was
-# fixed in the dev version. To install the dev version:
-library(devtools)
-install_github("ropensci/webchem")
-
+# fixed in the dev version.
 # query SMILES from PubChem
 cids <- get_cid(etox_cas)
 pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
@@ -67,7 +61,6 @@ res <- data.frame(name = subs,
 head(res)
 
 # Use Case II: Toxicity of different pesticide groups ---------------------
-
 # query the pesticide compendium using CAS-numbers
 aw_data <- aw_query(lc50$cas, type = "cas")
 
@@ -107,7 +100,6 @@ p
 # Use Case III: Name 10 most toxic chemicals ------------------------------
 cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
 chebiids <- get_chebiid(cas_rns)
-chebiids <- do.call(rbind, chebiids)
 comp <- chebi_comp_entity(chebiids$chebiid)
 pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
 

### 2020-03-29 revision3_2581/paper.tex

@@
 
-\author{Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And
+\author{
+	Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And
 	Tam\'as Stirling\\Independent Developer \And
+	Eric Scott\\Tufts University \And
 	Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
@@
 
-\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Andreas Scharm\"uller, Ralf B. Sch\"afer}
+\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Eric Scott, Andreas Scharm\"uller, Ralf B. Sch\"afer}
 
@@
 %
-When possible, \pkg{webchem} returns a data frame with one row per substance.
+When possible, \pkg{webchem} returns a data frame with one or more rows per substance.
 However, data from some sources can be very voluminous and not tabular.
 In these cases \pkg{webchem} always returns a named list (one entry for each substance).
-Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
+We provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
 %
@@
 R> chebiids <- get_chebiid(cas_rns)
-R> chebiids <- do.call(rbind, chebiids)
 R> comp <- chebi_comp_entity(chebiids$chebiid)
@@
 Table~\ref{tab:chebi} provides information on the individual roles of the three most toxic chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D.
-magna} in our table, is designed to be used as an acari- and insecticide, which inhibts the acetylcholinesterase enzyme.
+magna} in our table, is designed to be used as an acari- and insecticide, which inhibits the acetylcholinesterase enzyme.
 
@@
 \begin{CodeInput}
-R> cid <- get_cid(lc50$cas)
+R> cid <- get_cid(lc50$cas, first = TRUE)
 R> pc_data <- pc_prop(cid)

### 2020-04-01 revision3_2581/paper.tex
Updated corresponding author affiliation

@@
 	Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And
-	Tam\'as Stirling\\Independent Developer \And
-	Eric Scott\\Tufts University \And
+	Tam\'as Stirling\\Independent Researcher \And
+	Eric Scott\\Tufts University \AND
 	Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
@@
 \Address{
-  Eduard Sz\"ocs\\
-  Institute for Environmental Sciences\\
-  Universit\"at Koblenz-Landau\\
-  Fortstra{\ss}e 7\\
-  76829 Landau, Germany\\
-  E-mail: \email{szoecs@uni-landau.de}\\
-  URL: \url{https://edild.github.io}\\
-  \\
   Tam\'as Stirling\\
+  Independent Researcher \\
   Budapest, Hungary\\

### 2020-04-01 revision3_2581/paper.tex
Updated acknowledgements
changes 4.6 partitioning coefficients

--- 2020-03-27 revision3_2581/paper.tex
@@
 Some data sources also provide data on chemical properties that can be
-queried.  Here we query for the \code{lc50} data predicted
-$\mathrm{log}~P_{oct/wat}$ from the PubChem database to build a
+queried. Here we query for the \code{lc50} data predicted ocatanol-water partitioning coefficients
+($\mathrm{log}~\mathit{P_{oct/wat}}$) from the PubChem database to build a
 simple quantitative structure-activity relationship (QSAR) to predict
@@
   \caption{Simple QSAR for predicting log LC\textsubscript{50} of
-    pesticides by $\mathrm{log}~P$.  $\mathrm{log}~P$ values have been
-    retrieved from the Pubchem database (118 estimated data and 6 substances without data).  Blue line
+    pesticides by the ocatanol-water partitioning coefficient $\mathrm{log}~\mathit{P_{oct/wat}}$.
+    $\mathrm{log}~\mathit{P_{oct/wat}}$ values have been retrieved from the Pubchem
+    database (118 estimated data and 6 substances without data).  Blue line
     indicates the regression model
     ($\mathrm{log~LC}_{50} = 2.86\ensuremath{-0.36}~\mathrm{log}~P$,
-    RMSE = 1.47).}
+    RMSE = 1.47). The x-axis is shown not logarithmized.}
 \label{fig:fig3}

### 2020-04-01 revision3_2581/paper.tex

--- 2020-04-01 revision3_2581/paper.tex
@@
 their contributions to the \pkg{webchem} package, as well as all users who
-provided feedback and feature requests.  We are grateful to Zacharias
-Steinmetz for valuable comments on the manuscript.
+provided feedback and feature requests.
+We thank the rOpenSci community for sharing good practices and providing a motivating development environment.
+We are grateful to Zacharias Steinmetz for valuable comments on the manuscript.
 

### 2020-04-03 revision3_2581/paper.tex
Updated Fig.3 description

@@
   \caption{Simple QSAR for predicting log LC\textsubscript{50} of
-    pesticides by the ocatanol-water partitioning coefficient $\mathrm{log}~\mathit{P_{oct/wat}}$.
+    pesticides by the ocatanol-water partitioning coefficient $\mathit{P_{oct/wat}}$.
     $\mathrm{log}~\mathit{P_{oct/wat}}$ values have been retrieved from the Pubchem
@@
     ($\mathrm{log~LC}_{50} = 2.86\ensuremath{-0.36}~\mathrm{log}~P$,
-    RMSE = 1.47). The x-axis is shown not logarithmized.}
+    RMSE = 1.47).}
 \label{fig:fig3}

### 2020-04-03 revision3_2581/paper.tex

@@
 retrieve chemical information from the web, we created the
-\pkg{webchem} package \citep{cran} for the free and open source
+\pkg{webchem} package \citep{zenodo} for the free and open source
 \proglang{R} language \citep{r_2015, Wehrens_2011}.  \proglang{R} is
@@
 available under an MIT license.  The development repository is hosted
-on GitHub \citep{github} and a stable version is released on the
+on GitHub at \url{https://github.com/ropensci/webchem/} and a stable version is released on the
 Comprehensive \proglang{R} Archive Network (CRAN) and available at
@@
 part of the rOpenSci project \citep{boettiger2015building}, which aims
-at fully reproducible data analysis.
+at fully reproducible data analysis.  \pkg{webchem} is registered on the Resource Identification Portal \citep{scicrunch}.
 

### 2020-04-06 revision3_2581/paper.tex
Minor adjustment to authors

@@
 	Tam\'as Stirling\\Independent Researcher \And
-	Eric Scott\\Tufts University \AND
+	Eric R. Scott\\Tufts University \AND
 	Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
@@
 
-\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Eric Scott, Andreas Scharm\"uller, Ralf B. Sch\"afer}
+\Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Eric R. Scott, Andreas Scharm\"uller, Ralf B. Sch\"afer}
 
@@
 information and we encourage the interested reader to consult the
-related source for details.  However, we provide a brief overview in
-the Supporting Information.
+related source for details.
 

### 2020-05-24 revision3_2581/paper.tex
Update get function examples

@@
 \begin{CodeInput}
-R> cids <- get_cid(etox_cas)
-R> pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
+R> cids <- get_cid(etox_cas, from = "xref/rn", match = "first")
+R> pc_data <- pc_prop(cids$cid, properties = "CanonicalSMILES")
 R> pc_smiles <- smiles(pc_data)
@@
 Some data sources also provide data on chemical properties that can be
-queried. Here we query for the \code{lc50} data predicted ocatanol-water partitioning coefficients
+queried. Here we query for the \code{lc50} data predicted octanol-water partitioning coefficients
 ($\mathrm{log}~\mathit{P_{oct/wat}}$) from the PubChem database to build a
@@
 \begin{CodeInput}
-R> cid <- get_cid(lc50$cas, first = TRUE)
-R> pc_data <- pc_prop(cid)
+R> cid <- get_cid(lc50$cas, from = "xref/rn", match = "first")
+R> pc_data <- pc_prop(cid$cid)
 R> lc50$logp <- pc_data$XLogP

### 2020-05-24 revision3_2581/Code/code.R

@@ -39,8 +39,8 @@ head(etox_cas)
 # query other identifiers from other resources
 
 # query SMILES from PubChem
-cids <- get_cid(etox_cas)
-pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
+cids <- get_cid(etox_cas, from = "xref/rn", match = "first")
+pc_data <- pc_prop(cids$cid, properties = "CanonicalSMILES")
 pc_smiles <- smiles(pc_data)
 
 # ChemSpider needs an API key and RSC explicitly asks users to keep the key
@@ -69,7 +69,7 @@ str(aw_data[[1]])
 # extract chemical group from list
 igroup <- sapply(aw_data, function(y) {
   if (!is.na(y)) {
-    y$subactivity[1]  
+    y$subactivity[1]
   } else {
     NA
   }

### 2020-05-25 revision3_2581/paper.tex
Update get_etoxid() output

@@
 \begin{CodeOutput}
-  etoxid                           match distance                  query
-1   8668            2,4-Xylenol ( 8668 )        0     2,4-Dimethylphenol
-2   8494 4-Chlor-2-methylphenol ( 8494 )        0 4-Chlor-2-methylphenol
-3   <NA>                            <NA>     <NA>     4-para-nonylphenol
-4   8397                Atrazin ( 8397 )        0                Atrazin
-5   7240                 Benzol ( 7240 )        0                 Benzol
-6   7331        Desethylatrazin ( 7331 )        0        Desethylatrazin
+                   query                              match etoxid
+1     2,4-Dimethylphenol               2,4-Xylenol ( 8668 )   8668
+2 4-Chlor-2-methylphenol    4-Chlor-2-methylphenol ( 8494 )   8494
+3     4-para-nonylphenol                               <NA>   <NA>
+4                Atrazin                   Atrazin ( 8397 )   8397
+5                 Benzol                    Benzol ( 7240 )   7240
+6        Desethylatrazin           Desethylatrazin ( 7331 )   7331
 \end{CodeOutput}

### 2020-05-25 revision3_2581/Code/code.R
Remove reference to webchem dev version in code examples

@@ -4,10 +4,6 @@
 # from the Web
 # ------------------------------------------------------------------------
 
-# Install webchem from GitHub -----------------------------------------------
-library(devtools)
-install_github("ropensci/webchem")
-
 # load packages -----------------------------------------------------------
 library(webchem)
 library(ggplot2)

### 2020-06-14 revision3_2581/Code/code.R
Match published content

@@ -5,8 +5,8 @@
 # ------------------------------------------------------------------------
 
 # load packages -----------------------------------------------------------
-library(webchem)
-library(ggplot2)
+library("webchem")
+library("ggplot2")
 
 # load datasets -----------------------------------------------------------
 # load jagst dataset
@@ -64,11 +64,7 @@ aw_data <- aw_query(lc50$cas, type = "cas")
 str(aw_data[[1]])
 # extract chemical group from list
 igroup <- sapply(aw_data, function(y) {
-  if (!is.na(y)) {
-    y$subactivity[1]
-  } else {
-    NA
-  }
+  if (is(y, "list")) y$subactivity[1] else NA_character_
 })
 igroup[1:3]
 
@@ -84,6 +80,7 @@ lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates",
                                           "other"))
 
 # plot
+set.seed(1234)
 p <- ggplot(lc50, aes(x = type, y = value)) +
   geom_boxplot(fill = "grey75") +
   geom_jitter(width = 0.5) +
@@ -94,15 +91,17 @@ p <- ggplot(lc50, aes(x = type, y = value)) +
 p
 
 # Use Case III: Name 3 most toxic chemicals -------------------------------
-cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
+cas_rns <- lc50[order(lc50$value)[1:3], "cas"]
 chebiids <- get_chebiid(cas_rns)
+chebiids
 comp <- chebi_comp_entity(chebiids$chebiid)
-pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
+pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role", ]))
+pars
 
 # Use Case IV: Querying partitioning coefficients -------------------------
 # query PubChem DATABASE
-cid <- get_cid(lc50$cas, first = TRUE)
-pc_data <- pc_prop(cid)
+cid <- get_cid(lc50$cas, from = "xref/rn", match = "first")
+pc_data <- pc_prop(cid$cid)
 # lookat internal structure
 str(pc_data)
 # extract logP from properties data.frame
@@ -130,14 +129,11 @@ p
 # search EQS from ETOX using the already queried ETOX_IDs
 eqs <- etox_targets(ids$etoxid)
 # extract only MAC-EQS for the EU from the results
[…]

### 2020-06-14 revision3_2581/paper.tex
Update current version to v1.0.0 in the text

@@
 \graphicspath{{Figures/}}
+\shortcites{Boyle_Guha_2011,Kim_2016,wilson_best_2014,zenodo}
 
@@
   Andreas Scharm\"uller\\Universit\"at Koblenz-Landau \And
-    Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
-\title{\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web}
-
+  Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau
+}
 \Plainauthor{Eduard Sz\"ocs, Tam\'as Stirling, Eric R. Scott, Andreas Scharm\"uller, Ralf B. Sch\"afer}
 
+\title{\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web}
 \Plaintitle{webchem: An R Package to Retrieve Chemical Information from the Web}
+\Shorttitle{\pkg{webchem}: Retrieve Chemical Information from the Web in \proglang{R}}
 
-\Shorttitle{\pkg{webchem}: Chemical Information from the Web}
-
-\Abstract{ A wide range of chemical information is freely available
+\Abstract{
+  A wide range of chemical information is freely available
   online, including identifiers, experimental and predicted chemical
@@
   it reduces the time researchers need to spend on chemical data
-  compilation.  }
+  compilation.
+}
 
-\Keywords{ecotoxicology, chemistry, data cleaning, web scraping, ropensci}
+\Keywords{ecotoxicology, chemistry, data cleaning, web scraping, rOpenSci}
 
-\Plainkeywords{ecotoxicology, chemistry, data cleaning, web scraping, ropensci}
-
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
+\Volume{93}
+\Issue{13}
+\Month{May}
+\Year{2020}
+\Submitdate{2016-03-08}
+\Acceptdate{2019-09-19}
+%\DOI{10.18637/jss.v093.i13}
 
@@
 
-% citation alias for LUBW institution
-\defcitealias{lubw_2016}{LUBW (2020)}
-% \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \section[Data sources]{Data sources}\label{sec:data-sources}
+
+\begin{figure}[b!]
+  \centering
+  \includegraphics{fig1}
+  \caption{Overview of current data sources. Input and output possibilities currently implemented in the package.}
+  \label{fig:fig1}
+\end{figure}
+
 The backbone of \pkg{webchem} are data sources providing their data
@@
 Figure~\ref{fig:fig1}):
-
+%
 \begin{description}
 \item[NIH Chemical Identifier Resolver (CIR)]{\citep{cir} A web
-    service that converts from and to various chemical identifiers. The database holds millions of identifier combinations.}
-\item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
-	structure database providing access to over 67 million
-	structures from hundreds of data sources. It provides identifiers, properties and can also be
-	used to convert identifiers.}
+    service that converts from and to various chemical
+    identifiers. The database holds millions of identifier
+    combinations.}
+\item[ChemSpider]{\citep{pence_chemspider:_2010} A free chemical
+    structure database providing access to over 67 million structures
+    from hundreds of data sources. It provides identifiers, properties
+    and can also be used to convert identifiers.}
 \item[Chemical Translation Service (CTS)]{
@@
     data for over 64,000 entries and quality targets for different countries.}
-\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
+\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} A large web-based
 	database provided by the National Library of Medicine. It contains
@@
 \item[SRS]{\citep{srs} Substance Registry Services is the US EPA central system for information about substances that are tracked or regulated by EPA or other sources.}
-
 \end{description}
-
+%
 Though the data sources exhibit some overlap in the provided
@@
 related source for details.
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics{fig1}
-  \caption{Overview of current data sources. Input and output possibilities currently implemented in the package.}
-  \label{fig:fig1}
-\end{figure}
 
@@
 version is available from GitHub and also permanently available at \citet{zenodo}.  This document has been created using \pkg{webchem}
-version 0.5.0.
+version 1.0.0.
 
@@
 (i) \code{jagst}: This data set comprises environmental monitoring data of organic substances in the river Jagst, Germany, sampled in 2013.
-The data is publicly available and can be retrieved from \citetalias{lubw_2016}.
+The data is publicly available and can be retrieved from \cite{lubw_2016}.
 It comprises concentrations  (in $\mathrm{\mu g~/~L}$) of  34 substances  on 13 sampling occasions.
@@
 \end{CodeChunk}
-%
+
+\vspace*{-0.25cm}
+
 This data set identifies the substances only by CAS numbers.
@@
 First we search for ETOX internal ID numbers using the substance names:
-%
+
+\vspace*{-0.35cm}
+
 \begin{CodeChunk}
@@
 \end{CodeInput}
+
+\vspace*{-0.5cm}
+
 \begin{CodeOutput}
+# A tibble: 6 x 3
   query                  match                           etoxid
+  <chr>                  <chr>                           <chr> 
 1 2,4-Dimethylphenol     2,4-Xylenol ( 8668 )            8668  
@@
 \end{CodeChunk}
-%
+
+\vspace*{-0.25cm}
+
 Only three substances could not be found in ETOX.  Here we specify
@@
 We use these data to retrieve basic information on the substances.
-%
+
+\vspace*{-0.35cm}
+
 \begin{CodeChunk}
@@
 following examples).
+
+\begin{table}[t!]
+\centering
+\begin{tabular}{llllll}
+  \hline
+Name & CAS & SMILES & CID & InChIKey & CSID \\ 
+  \hline
+2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
+  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
+  4-para-nonylphenol & -- & -- & -- & -- & -- \\ 
+  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
+  Benzol & 71-43-2 & C1=CC=C... & 241 & UHOVQNZ... & 236 \\ 
+  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
+   \hline
+\end{tabular}
+\caption{Identifiers for the \code{jagst} data sets as queried with
+  \pkg{webchem}. Only the first 6 entries are shown. For SMILES and
+  InChIKey only the first 7 characters are shown. -- = not found.}
+\label{tab:comptable}
+\end{table}
 
@@
 
-\begin{table}[t!]
+\subsection[Toxicity of different pesticide groups]{Toxicity of different pesticide groups}
+
+\begin{figure}[t!]
   \centering
-\begin{tabular}{llllll}
-  \hline
-Name & CAS & SMILES & CID & InChIKey & CSID \\ 
-  \hline
-2,4-Dimethylphenol & 105-67-9 & CC1=CC(... & 7771 & KUFFULV... & 13839123 \\ 
-  4-Chlor-2-methylphenol & 1570-64-5 & CC1=C(C... & 14855 & RHPUJHQ... & 14165 \\ 
-  4-para-nonylphenol & -- & -- & -- & -- & -- \\ 
-  Atrazin & 1912-24-9 & CCNC1=N... & 2256 & MXWJVTO... & 2169 \\ 
-  Benzol & 71-43-2 & C1=CC=C... & 241 & UHOVQNZ... & 236 \\ 
-  Desethylatrazin & 6190-65-4 & CC(C)NC... & 22563 & DFWFIQK... & 21157 \\ 
-   \hline
-\end{tabular}
-\caption{Identifiers for the \code{jagst} data sets as queried with
-  \pkg{webchem}. Only the first 6 entries are shown. For SMILES and
-  InChIKey only the first 7 characters are shown. -- = not found.}
-\label{tab:comptable}
-\end{table}
+  \includegraphics[width=0.7\textwidth, trim = 0 10 0 5, clip]{plot_lc50-1} 
+\caption{Toxicity of different pesticide groups. LC\textsubscript{50}
+  values have been retrieved from the EPA ECOTOX database, chemical groups
+  from the Compendium of Pesticide Common Names \citep{wood}.}
+\label{fig:fig2}
+\end{figure}
 
-\subsection[Toxicity of different pesticide groups]{Toxicity of different pesticide groups}
 Another question we might ask is \emph{How does toxicity vary between
@@
 R> igroup <- sapply(aw_data, function(y) {
-R>   if (!is.na(y)) {
-R>     y$subactivity[1]  
-R>   } else {
-R>     NA
-R>   }
-R>  })
++    if (is(y, "list")) y$subactivity[1] else NA_character_
++  })
 R> igroup[1:3]
@@
 
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=0.7\textwidth, trim = 0 10 0 5, clip]{plot_lc50-1} 
-\caption{Toxicity of different pesticide groups. LC\textsubscript{50}
-  values have been retrieved from the EPA ECOTOX database, chemical groups
-  from the Compendium of Pesticide Common Names \citep{wood}.}
-\label{fig:fig2}
-\end{figure}
-
 \subsection[Identify common names and roles of chemicals]{Identify common names and roles of chemicals}
-We might be interested in common names and roles of the most toxic chemicals in the \code{lc50} data set, i.e., chemicals with the lowest $LC_{50}$ values.
-\pkg{webchem} can query the ChEBI database \citep{chebi} to retrieve common names and roles of chemicals.
-We extract the CAS numbers of the three most toxic chemicals, and then use \code{get_chebiid()} to retrieve ChEBI identifiers and substance names.
-Subsequently, we use the acquired ChEBI identifiers to query the complete entity by using \code{chebi_comp_entity()}.
-From the result, we extract the parents table and refine the data to chemical roles.
-
+We might be interested in common names and roles of the most toxic
+chemicals in the \code{lc50} data set, i.e., chemicals with the lowest
+$LC_{50}$ values. \pkg{webchem} can query the ChEBI database
+\citep{chebi} to retrieve common names and roles of chemicals. We
+extract the CAS numbers of the three most toxic chemicals, and then
+use \code{get_chebiid()} to retrieve ChEBI identifiers and substance
+names. Subsequently, we use the acquired ChEBI identifiers to query
+the complete entity by using \code{chebi_comp_entity()}. From the
+result, we extract the parents table and refine the data to chemical
+roles.
+%
 \begin{CodeChunk}
 \begin{CodeInput}
-R> cas_rns <- lc50[order(lc50$value)[1:3],"cas"]
+R> cas_rns <- lc50[order(lc50$value)[1:3], "cas"]
 R> chebiids <- get_chebiid(cas_rns)
 R> comp <- chebi_comp_entity(chebiids$chebiid)
-R> pars <- lapply(comp, function(x) with(x, parents[parents$type == "has role",]))
+R> pars <- lapply(comp, function(x) 
++    with(x, parents[parents$type == "has role", ]))
 \end{CodeInput}
 \end{CodeChunk}
-
-Finally we can compile the data into a table.
-Table~\ref{tab:chebi} provides information on the individual roles of the three most toxic chemicals and we observe that, for example ethion, the most toxic pesticide towards \textit{D.
-magna} in our table, is designed to be used as an acari- and insecticide, which inhibits the acetylcholinesterase enzyme.
+%
+Finally we can compile the data into a table. Table~\ref{tab:chebi}
+provides information on the individual roles of the three most toxic
+chemicals and we observe that, for example ethion, the most toxic
+pesticide towards \textit{D.\ magna} in our table, is designed to be
+used as an acari- and insecticide, which inhibits the
+acetylcholinesterase enzyme.
 
@@
 \end{tabular}
-\caption{CAS numbers, names and chemical roles from ChEBI for the three most toxic substances in the \code{lc50} data set towards \textit{D. magna}.}
+\caption{CAS numbers, names and chemical roles from ChEBI for the three most toxic substances in the \code{lc50} data set towards \textit{D.\ magna}.}
 \label{tab:chebi}
@@
 R> ids$mac <- sapply(eqs, function(y) {
-+    if (length(y) == 1 && is.na(y)) {
-+      return(NA) 
-+    } else {
-+      res <- y$res
-+      min(res[res$Country_or_Region == "EEC / EU" & 
-+        res$Designation == "MAC-EQS", "Value_Target_LR"])
-+    }
++    if (is(y, "list")) {
++      min(subset(y$res, Country_or_Region == "EEC / EU" & 
++        Designation == "MAC-EQS")[["Value_Target_LR"]])
++    } else NA_real_
 +  })
@@
 \begin{CodeOutput}
+# A tibble: 6 x 3
   etoxid query           mac
-4    8397     Atrazin  2.000
-5    7240      Benzol 50.000
-11   8836     Irgarol  0.016
-12   7442 Isoproturon  1.000
-38   7571     Simazin  4.000
-29   8756   Terbutryn  0.034
+  <chr>  <chr>         <dbl>
+1 8397   Atrazin      2     
+2 7240   Benzol      50     
+3 8836   Irgarol      0.0160
+4 7442   Isoproturon  1     
+5 7571   Simazin      4     
+6 8756   Terbutryn    0.0340
 \end{CodeOutput}
@@
 
-\subsection[Open Science]{Open Science}
+\subsection[Open science]{Open science}
 An increasing number of scientific data is becoming publicly available
@@
 
-\bibliography{ref}
+\bibliography{v93i13}
 
